---
title: 'Moran effect explorations for stream metab and lake metab synchrony analysis'
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
# setwd("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis")

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB", "plotly", "MASS", "EnvStats",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)

```


### REad in PRISM climate data
```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
library(dplyr)
library(readr)

# Define the folder path
folder_path <- "/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Clim_data/PRISM_data"

# List all .csv files in the folder
file_list <- list.files(folder_path, pattern = "\\.csv$", full.names = TRUE)

# Read in all CSV files and combine them into one dataframe
PRISM_data <- file_list %>%
  lapply(function(file) read_csv(file, skip = 10)) %>%
  bind_rows() %>%
  rename(
    ppt_mm = `ppt (mm)`,
    TminC = `tmin (degrees C)`,
    TmaxC = `tmax (degrees C)`,
    TmeanC = `tmean (degrees C)`
  )
# View the first few rows
head(PRISM_data)
str(PRISM_data)


unique(PRISM_data$Name)



summary(PRISM_data)

```

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}

ppt_trends <- PRISM_data %>%
  ggplot(aes(y = ppt_mm, x = Date, color = Latitude)) +
  geom_line(linewidth=1) + theme_classic() #+ ylab(expression(Accumulated~flow~(m^3~km^-1~d^-1))) + xlab(NULL) 

ppt_trends


temp_trends <- PRISM_data %>%
  ggplot(aes(y = TmeanC, x = Date, color = Latitude)) +
  geom_line(linewidth=1) + theme_classic() #+ ylab(expression(Accumulated~flow~(m^3~km^-1~d^-1))) + xlab(NULL) 

temp_trends


```

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
### Marzolf workflow parameters 
# define constants and axis labels
metab_units_area <- (expression(paste('GPP (g C', m^-2,' ',y^-1,')', sep = ' ')))

low_p <- '#103801'
high_p <- '#43fa00'

# quotient to convert GPP from O2 to C
# g C = g O2 * (1 mol O2/32 g O2) * (2 mol O/1 mol O2) * (1 mol C/1 mol O) * (12 g C/1 mol 

rq <- (1/32)*(2/1)*(1/1)*(12/1)

## check this to see what a better assumption would be ?
rq_e <- 1

# where to save figures
save_figs <- TRUE
dpi <- 400

```

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
## Read in data 

metab_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_TS.rds")
# str(metab_ts)

dat_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/TS_SPC_pH_FNU.rds")
# str(dat_ts)
# summary(dat_ts)
# unique(dat_ts$site_name)

metab_meta_dat <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata.rds")
# str(metab_ts)


marz_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/1_site_info.csv")
# 63 different sites 

Ladwig_NTL_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/Ladwig_NTL_lake_cords.csv")%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="ladwig_NTL_lake")

Oleksy_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/lake_descriptions.csv")%>%
  dplyr::select(Site.Number ="lakeName", Latitude="lat_dd", Longitude="long_dd") %>%
  mutate(source="oleksy_lake")

Rabaey_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metadata.csv")%>%
  dplyr::select(Site.Number ="pond_name", Latitude="latitude", Longitude="longitude") %>%
  mutate(source="rabaey_lake")

corman_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/load_metab_results/corman_lake_cords.csv")%>%
  filter(country=="United States" | country== "Canada") %>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="corman_lake")

lotting_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/Lotting_lake_cords.csv")%>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="lotting_lake")


```


```{r, warning=F, size = 'small', echo=F, include=F}

dat_ts_clean <- dat_ts%>%
  filter(GPP.Rhat < 1.05 &  GPP >0 & ER <0)

# unique(dat_ts_clean$long_name)

marz_dat_names <- marz_dat%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="marzolf_stream")

dat_ts_names <-metab_meta_dat %>%
   dplyr::select(Site.Number = "site_no", Latitude = "dec_lat_va", Longitude= "dec_long_va") %>%
  mutate(source="appling_stream")

## quick site check
site_check <- rbind(marz_dat_names, 
                  #  dat_ts_names,
                    Ladwig_NTL_dat_metadf, 
                    Oleksy_dat_metadf,
                    Rabaey_dat_metadf, 
                    #corman_dat_metadf, 
                    lotting_dat_metadf)

#

# write.csv(site_check, "/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/RR_site_cords.csv")


```


#### Load Marzolf GPP data:

```{r, warning=F, size = 'small', echo=T, include=T}
river_metab <- readr::read_csv('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/5_usgs_metabolism.csv')

head(river_metab)
```

#### Gap fill

'fill_and_normalize_metab.R' is a modified function
from'code/metabolism_synthesis/Part II- Filtering and gap filling.Rmd
uses the function fillMiss3.R

**Make sure** to double check the 'summary_site-year_NAs.csv' save path
in the fxn

-   Extracts unique sites and years from the input data.

-   Tracks missing data for each site-year.

-   Pads the dataset to ensure a full-year time series.

-   Identifies missing data patterns and determines if a year is
    complete.

-   Applies a gap-filling algorithm (if \<40% of data is missing).

-   Outputs a cleaned and gap-filled dataset for further analysis.



```{r, warning=F, size = 'small', echo=F, include=T}
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/code_fxns/fill_and_normalize_metab.R')

```

```{r, warning=F, size = 'small', echo=T, include=T}
river_metab_filled <- fill_and_normalize_metab(river_metab)
head(river_metab_filled)
```

##### Check the scope of the infill

```{r, warning=F, size = 'small', echo=T, include=T}
n_obs_gpp <- length(river_metab_filled$GPP_filled)
n_obs_gpp

n_sites <- length(unique(river_metab_filled %>%
                           dplyr::filter(!is.na(site)) %>%
                           dplyr::pull(site)))
n_sites
```

```{r, warning=F, size = 'small', echo=T, include=T}
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/code_fxns/fill_and_normalize_metab_KAL.R')

river_metab_filled_ER <- fill_and_normalize_metab(river_metab)
head(river_metab_filled_ER)
```

```{r, warning=F, size = 'small', echo=T, include=T}

n_obs_er <- length(river_metab_filled_ER$ER_filled)
n_obs_er

n_sites <- length(unique(river_metab_filled_ER %>%
                           dplyr::filter(!is.na(site)) %>%
                           dplyr::pull(site)))
n_sites
```

### Calculate cumulative daily GPP and total annual GPP to get a variance dynamics:

```{r, warning=F, size = 'small', echo=F, include=F}
table_s1 <- readr::read_csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/1_site_info.csv")
```

```{r, warning=F, size = 'small', echo=F, include=F}
# calculate site-year cumulative GPP
river_metab_cumsum <- river_metab_filled %>%
  dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::mutate(jday = lubridate::yday(date),
                GPP_C = GPP_filled*rq,
                GPP_C_clean = replace(GPP_C, GPP_C < 0, 0), # replace negative values with 0 in GPP_C_cumsum
                GPP_C_cumsum = cumsum(GPP_C_clean))


# calculate annual GPP for each site-year
metab_all_annual <- river_metab_cumsum %>%
#dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::summarise(GPP_daily_mean = mean(GPP_C_clean, na.rm = TRUE),
                   GPP_ann_sum = sum(GPP_C_clean, na.rm = TRUE),
                   GPP_daily_cv = EnvStats::cv(GPP_C_clean, na.rm = TRUE, method = 'l.moments')*100,
                   Q_daily_mean = mean(discharge, na.rm = TRUE),
                   Q_daily_cv = EnvStats::cv(discharge, na.rm = TRUE, method = 'l.moments')*100,
                   temp_daily_mean = mean(temp.water, na.rm = TRUE),
                   temp_daily_cv = EnvStats::cv(temp.water,na.rm = TRUE, method = 'l.moments')*100) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`,
                            clean_name =`Name`) ,
                   'clean_site_num')
```

```{r, warning=F, size = 'small', echo=T, include=F}
##### Calculate median annual and CV for each site time series

site_mean_ann_gpp <- metab_all_annual %>%
  dplyr::group_by(site, clean_name) %>%
  dplyr::summarise(median_ann_GPP = median(GPP_ann_sum, na.rm = TRUE),
                   sd_ann_GPP = sd(GPP_ann_sum, na.rm = TRUE),
                   min_ann_GPP = min(GPP_ann_sum, na.rm = TRUE),
                   max_ann_GPP = max(GPP_ann_sum, na.rm = TRUE),
                   GPP_percent_diff = ((max_ann_GPP- min_ann_GPP)/min_ann_GPP)*100,
                   GPP_ann_cv = EnvStats::cv(GPP_ann_sum, na.rm = TRUE, method = 'l.moments')*100,
                   mean_ann_temp = mean(temp_daily_mean, na.rm = TRUE),
                   mean_sd_temp = sd(temp_daily_mean, na.rm = TRUE),
                   cv_mean_temp = EnvStats::cv(temp_daily_mean, method = 'l.moments')*100,
                   median_ann_discharge = median(Q_daily_mean, na.rm = TRUE),
                   median_sd_discharge = sd(Q_daily_mean, na.rm = TRUE),
                   cv_median_discharge = EnvStats::cv(Q_daily_mean, na.rm = TRUE, method ='l.moments')*100
                   ) %>%
  dplyr::arrange(desc(max_ann_GPP)) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`),'clean_site_num')

head(site_mean_ann_gpp)

```

##### Cumulative site-year GPP as Fig 1 in the paper

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}

river_metab_cumsum_names <- river_metab_cumsum %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`) ,'clean_site_num') %>%
  dplyr::rename(clean_name =`Name`)

river_metab_cumsum_names$clean_name <-
  suppressWarnings(forcats::fct_relevel(river_metab_cumsum_names$clean_name,site_mean_ann_gpp %>%
                                          dplyr::pull(clean_name)))

y_axis_title <- bquote(Cumulative~ .(metab_units_area))
# subset of 9 sites to show as figure 1 (for size/space concerns)
highest_4 <- site_mean_ann_gpp %>%
  arrange(desc(median_ann_GPP)) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

lowest_4 <- site_mean_ann_gpp %>%
  arrange(median_ann_GPP) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

extremes <- c(lowest_4, highest_4)

fig_1 <- ggplot(data = river_metab_cumsum_names %>%
                  dplyr::filter(site %in% extremes),
                aes(x = jday, y = GPP_C_cumsum,
                    group = interaction(year, site),
                    color = GPP_C_cumsum))+
  geom_line(show.legend = FALSE) +
  facet_wrap(forcats::fct_relevel(clean_name,
                                  site_mean_ann_gpp %>%
                                    dplyr::filter(site %in% extremes) %>%
                                    dplyr::pull(clean_name))~., scales = 'free_y',
             ncol = 4) +
  scale_color_gradient(low = low_p, high = high_p) +
  scale_y_continuous(breaks = function(y) {
    max_value <- max(y, na.rm = TRUE)*0.9
    tick_marks <- seq(0, round(max_value, digits =-1), length.out = 3)
    return(tick_marks)
  }
  ) +
  labs(x = 'Day of Year',
       y = expression(paste('Cumulative GPP (g C', m^-2, '', y^-1,')'))) +
       theme(strip.text = element_text(size = 7),
             strip.background = element_rect(fill = 'white')) +
  theme_classic()

fig_1
```

#### Same thing but for cummulative ER?

```{r, warning=F, size = 'small', echo=F, include=F}
# calculate site-year cumulative ER
river_metab_cumsum_ER <- river_metab_filled_ER %>%
  dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::mutate(jday = lubridate::yday(date),
                ER_C = (ER_filled*-1) * rq_e, # absolute value of ER
                ER_C_clean = replace(ER_C, ER_C < 0, 0), # replace negative values with 0 in ER_C_cumsum
                ER_C_cumsum = cumsum(ER_C_clean))


# calculate annual ER for each site-year
metab_all_annual_er <- river_metab_cumsum_ER %>%
#dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::summarise(ER_daily_mean = mean(ER_C_clean, na.rm = TRUE),
                   ER_ann_sum = sum(ER_C_clean, na.rm = TRUE),
                   ER_daily_cv = EnvStats::cv(ER_C_clean, na.rm = TRUE, method = 'l.moments')*100,
                   Q_daily_mean = mean(discharge, na.rm = TRUE),
                   Q_daily_cv = EnvStats::cv(discharge, na.rm = TRUE, method = 'l.moments')*100,
                   temp_daily_mean = mean(temp.water, na.rm = TRUE),
                   temp_daily_cv = EnvStats::cv(temp.water,na.rm = TRUE, method = 'l.moments')*100) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`,
                            clean_name =`Name`) ,
                   'clean_site_num')
```

```{r, warning=F, size = 'small', echo=F, include=F}
##### Calculate median annual and CV for each site time series

site_mean_ann_er <- metab_all_annual_er %>%
  dplyr::group_by(site, clean_name) %>%
  dplyr::summarise(median_ann_ER = median(ER_ann_sum, na.rm = TRUE),
                   sd_ann_ER = sd(ER_ann_sum, na.rm = TRUE),
                   min_ann_ER = min(ER_ann_sum, na.rm = TRUE),
                   max_ann_ER = max(ER_ann_sum, na.rm = TRUE),
                   ER_percent_diff = ((max_ann_ER- min_ann_ER)/min_ann_ER)*100,
                   ER_ann_cv = EnvStats::cv(ER_ann_sum, na.rm = TRUE, method = 'l.moments')*100,
                   mean_ann_temp = mean(temp_daily_mean, na.rm = TRUE),
                   mean_sd_temp = sd(temp_daily_mean, na.rm = TRUE),
                   cv_mean_temp = EnvStats::cv(temp_daily_mean, method = 'l.moments')*100,
                   median_ann_discharge = median(Q_daily_mean, na.rm = TRUE),
                   median_sd_discharge = sd(Q_daily_mean, na.rm = TRUE),
                   cv_median_discharge = EnvStats::cv(Q_daily_mean, na.rm = TRUE, method ='l.moments')*100
                   ) %>%
  dplyr::arrange(desc(max_ann_ER)) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`),'clean_site_num')

```

##### Cumulative site-year \|ER\| modfied from Fig 1 in the paper

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}

river_metab_cumsum_names_ER <- river_metab_cumsum_ER %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`) ,'clean_site_num') %>%
  dplyr::rename(clean_name =`Name`)

river_metab_cumsum_names_ER$clean_name <-
  suppressWarnings(forcats::fct_relevel(river_metab_cumsum_names_ER$clean_name,site_mean_ann_er %>%
                                          dplyr::pull(clean_name)))

y_axis_title <- bquote(Cumulative~ .(metab_units_area))
# subset of 9 sites to show as figure 1 (for size/space concerns)
highest_4 <- site_mean_ann_er %>%
  arrange(desc(median_ann_ER)) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

lowest_4 <- site_mean_ann_er %>%
  arrange(median_ann_ER) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

extremes <- c(lowest_4, highest_4)

fig_1er <- ggplot(data = river_metab_cumsum_names_ER %>%
                  dplyr::filter(site %in% extremes),
                aes(x = jday, y = ER_C_cumsum,
                    group = interaction(year, site),
                    color = ER_C_cumsum))+
  geom_line(show.legend = FALSE) +
  facet_wrap(forcats::fct_relevel(clean_name,
                                  site_mean_ann_er %>%
                                    dplyr::filter(site %in% extremes) %>%
                                    dplyr::pull(clean_name))~., scales = 'free_y',
             ncol = 4) +
  scale_color_gradient(low = low_p, high = high_p) +
  scale_y_continuous(breaks = function(y) {
    max_value <- max(y, na.rm = TRUE)*0.9
    tick_marks <- seq(0, round(max_value, digits =-1), length.out = 3)
    return(tick_marks)
  }
  ) +
  labs(x = 'Day of Year',
       y = expression(paste('Cumulative |ER| (g C', m^-2, '', y^-1,')'))) +
       theme(strip.text = element_text(size = 7),
             strip.background = element_rect(fill = 'white')) +
  theme_classic()

fig_1er
```

##### When are rivers productive?

Where - Marzolf et al - identifies when cumulative GPP (a measure of
ecosystem productivity) reaches specific percentiles (25%, 50%, 75%, and
95%) for each site and year

```{r, warning=F, size = 'small', echo=T, include=T}
df_out <- data.frame(
  site = rep(rep(unique(river_metab_cumsum_names$site), each = 14),times = 4),
  year = rep(rep(unique(river_metab_cumsum_names$year), times = 59),each = 4),
  quantile = rep(c(0.25, 0.5, 0.75, 0.95), times = 14*59))

for(i in 1:nrow(df_out)){
  deets <- df_out[i,]
  site <- deets[,'site']
  year <- deets[,'year']
  quantile <- deets[,'quantile']
  
  df_use <- river_metab_cumsum_names %>%
    dplyr::filter(site %in% !!site,
                  year %in% !!year)
  if(nrow(df_use)== 0)
    next
# Calculate the Maximum Cumulative GPP (GPP_C_cumsum) for the Year:
  cumsum <- max(df_use$GPP_C_cumsum, na.rm = TRUE)
  df_out[i,'GPP_C_cumsum'] <- cumsum

# Normalize the Cumulative GPP to Create a CDF (Cumulative Distribution Function):
  df_math <- df_use %>%
    dplyr::mutate(GPP_C_cumsum_cdf = GPP_C_cumsum/cumsum)
  
# Find the Julian Day (jday) When the GPP Reaches the Specified Quantile:
  jday <- which.min(abs(df_math$GPP_C_cumsum_cdf- quantile))
  df_out[i,'jday'] <- jday
  
  # Extract Corresponding Data for That jday:
  row <- data.frame(df_math %>%
                      dplyr::filter(jday== !!jday))
  df_out[i,'clean_name'] <- row[,'clean_name']
  df_out[i,'date'] <- row[,'date']
  df_out[i,'GPP_C_cumsum'] <- row[,'GPP_C_cumsum']
  df_out[i,'cdf'] <- row[,'GPP_C_cumsum_cdf']
  } # end for loop

# Filter and Fill Missing Data (df_quants):
df_quants <- df_out %>%
  dplyr::filter(!is.na(jday)) %>%
  dplyr::group_by(site) %>%
  tidyr::fill(clean_name, .direction = 'down')

df_quants %>%
  dplyr::group_by(quantile) %>%
  dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
                   jday_sd = sd(jday, na.rm = TRUE),
                   jday_min = min(jday, na.rm = TRUE),
                   jday_max = max(jday, na.rm = TRUE))
```

```{r, warning=F, size = 'small', echo=T, include=T}
df_quants %>%
  dplyr::group_by(quantile, clean_name) %>%
  dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
                   jday_sd = sd(jday, na.rm = TRUE),
                   jday_min = min(jday, na.rm = TRUE),
                   jday_max = max(jday, na.rm = TRUE))

```

#### Trend analysis:

trend analysis on quantiles of cumulative gross primary production
(GPP_C_cumsum) over time, using Sen's slope and the Mann-Kendall test to
determine the direction and significance of trends in the Julian day
(jday) at which specific cumulative GPP quantiles are reached.

```{r, warning=F, size = 'small', echo=T, include=T}
# trend analysis for quantiles
quantile_trends <- df_quants %>%
  dplyr::arrange(site, year) %>%
  dplyr::group_by(site, quantile) %>%
  tidyr::nest() %>%
  # Performing Sen's Slope and Mann-Kendall Trend Tests
  dplyr::mutate(sens = purrr::map(data, ~trend::sens.slope(x = .$jday)),
                mk_test = purrr::map(data, ~trend::mk.test(x = .$jday))) %>%
  # Extracting Trend Statistics
  dplyr::mutate(sens_s = purrr::map_dbl(sens,~.$estimates), 
                sens_p = purrr::map(sens,~.$p.value),
                mk_p = purrr::map_dbl(mk_test,~.$p.value),
                # mk_t = purrr::map_dbl(mk_test, ~.$estimates['tau']),
                sens_sig = ifelse(sens_p <= 0.05, 'significant', 'non-significant'),
                sens_slope = ifelse(sens_s > 0, 'increasing',  'decreasing'),
                mk_sig = ifelse(mk_p <= 0.05, 'significant', 'non-significant')
                )

quantile_sum_sigs <- quantile_trends %>%
  dplyr::ungroup() %>%
  dplyr::select(quantile, site, mk_p, #mk_t,
                mk_sig, #mk_slope,
                sens_s, sens_sig, sens_slope)

quants_n_sig <- quantile_sum_sigs %>%
  dplyr::filter(mk_p <= 0.05) %>%
  dplyr::group_by(quantile, sens_slope) %>%
  summarise(n = n()) %>%
  data.frame()

quantile_trends_sum <- df_quants %>%
  dplyr::left_join(quantile_sum_sigs) %>%
  dplyr::mutate(sens_meaning = dplyr::case_when(sens_slope== 'increasing' ~ 'Later',
                                                sens_slope== 'decreasing' ~ 'Earlier'),
# mk_meaning = case_when(mk_slope =='increasing' ~ 'Later',
# mk_slope =='decreasing' ~ 'Earlier'),
)


```

**sens.slope**: Sen’s slope (sens.slope) estimates the trend in jday
over time (whether the day at which a given quantile is reached is
shifting earlier or later).

**mk.test**: Mann-Kendall test (mk.test) assesses whether there is a
statistically significant trend in jday over time.

#### Visualize the when productivity is highest

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}
library(gghighlight)

quantile_trenders <- quantile_trends_sum %>%
  dplyr::filter(mk_p <= 0.05) %>%
  dplyr::group_by(quantile, clean_name) %>%
  dplyr::filter(row_number()== 1)

fig_3 <- ggplot(quantile_trends_sum, aes(x = year, y = jday, group = site, color = sens_meaning))+
  geom_line(linewidth = 0.8)+
  gghighlight::gghighlight(mk_sig== 'significant',
                           use_direct_label = FALSE, keep_scales = TRUE,
                           unhighlighted_params = list(alpha("grey10", 0.4),
                                                       linewidth = 0.4),
                           calculate_per_facet = TRUE)+
  facet_wrap(.~ quantile, scales = 'free_y', nrow = 2, ncol = 2)+
#scale_y_continuous(limits = c(0, 366))+
  scale_color_manual(name = element_blank(), values = c(high_p, low_p))+ 
  labs(x = 'Year', y = 'Day of Year')+
  theme(legend.title = element_blank()) + theme_bw()

fig_3

```


## NEW STUFF:

```{r, warning=F, size = 'small', echo=T, include=T}

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual_mon <- river_metab_filled %>%
  mutate(month=month(date))%>%
  group_by(site, year, month) %>%
  summarise(
    monthly_avg_GPP = mean(GPP_filled, na.rm = TRUE),
    monthly_cv_GPP = sd(GPP_filled, na.rm = TRUE) / mean(GPP_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()

# Summarize intra-annual CV for each year
data_intra_annual_yr <- data_intra_annual_mon %>%
  group_by(site,year) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_GPP, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )

# Summarize intra-annual CV for each year
data_intra_annual_overall <- data_intra_annual_mon %>%
  group_by(site) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_GPP, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )
```

```{r, warning=F, size = 'small', echo=F, include=F}

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual_mon_er <- river_metab_filled_ER %>%
  mutate(month=month(date))%>%
  group_by(site, year, month) %>%
  summarise(
    monthly_avg_ER = mean(ER_filled, na.rm = TRUE),
    monthly_cv_ER = sd(ER_filled, na.rm = TRUE) / mean(ER_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()

# Summarize intra-annual CV for each year
data_intra_annual_yr_er <- data_intra_annual_mon_er %>%
  group_by(site,year) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_ER, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_ER, na.rm = TRUE) / mean(monthly_avg_ER, na.rm = TRUE) * 100
  )

# Summarize intra-annual CV for each year
data_intra_annual_overall_er <- data_intra_annual_mon_er %>%
  group_by(site) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_ER, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_ER, na.rm = TRUE) / mean(monthly_avg_ER, na.rm = TRUE) * 100
  )
```

#### Mean GPP by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
# Create the plot
ggplot(data_intra_annual_yr, aes(x = year, y = mu_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  #  scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual GPP by stream site",
    x = "Year",
    y = "GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
#  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### Mean ER by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
# Create the plot
ggplot(data_intra_annual_yr_er, aes(x = year, y = mu_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  #  scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual ER by stream site",
    x = "Year",
    y = "ER"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
#  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### CV in GPP by site per year

CV as sd()/mean() \*100

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
ggplot(data_intra_annual_yr, aes(x = year, y = cv_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site) +
  theme_minimal() +  
  labs(
    title = "Degree of CV in GPP each year stream site",
    x = "Year",
    y = "CV of GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### CV in ER by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
ggplot(data_intra_annual_yr_er%>%filter(cv_intra_annual>-1940), aes(x = year, y = cv_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site) +
  theme_minimal() +  
  labs(
    title = "Degree of CV in GPP each year stream site",
    x = "Year",
    y = "CV of GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

```{r, warning=F, size = 'small', echo=T}

library(ggplot2)
library(ggrepel)
library(viridis)

spc_tbl <- table_s1 %>%
  mutate(site = paste0("nwis_", `Site Number`))

# Ensure `site` is correctly formatted in both dataframes
spc_tbl <- spc_tbl %>%
  mutate(site = paste0("nwis_", str_pad(`Site Number`, width = 8, pad = "0")))


#data_intra_annual_mon

site_check_with_cv <- spc_tbl %>%
  left_join(data_intra_annual_overall, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")

site_check_with_cv <- spc_tbl %>%
  left_join(data_intra_annual_overall, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")



site_check_with_cv_er <- spc_tbl %>%
  left_join(data_intra_annual_overall_er, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")

```

#### Histograms of mean GPP by location:

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=14}
site_check_with_cv$drainage_category <- cut(site_check_with_cv$drainage_area_km,
                                            breaks = quantile(site_check_with_cv$drainage_area_km, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("Small", "Medium", "Large", "Very Large"),
                                            include.lowest = TRUE)


site_check_with_cv$lat_category <- cut(site_check_with_cv$Latitude,
                                            breaks = quantile(site_check_with_cv$Latitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("south", "mid-south", "mid-north", "north"),
                                            include.lowest = TRUE)

site_check_with_cv$lon_category <- cut(site_check_with_cv$Longitude,
                                            breaks = quantile(site_check_with_cv$Longitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("west", "west-central", "east-central", "east"),
                                            include.lowest = TRUE)


hist1<- ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "Mean GPP", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "Mean GPP", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "Mean GPP", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}
site_check_with_cv_er$drainage_category <- cut(site_check_with_cv_er$drainage_area_km,
                                            breaks = quantile(site_check_with_cv_er$drainage_area_km, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("Small", "Medium", "Large", "Very Large"),
                                            include.lowest = TRUE)


site_check_with_cv_er$lat_category <- cut(site_check_with_cv_er$Latitude,
                                            breaks = quantile(site_check_with_cv_er$Latitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("south", "mid-south", "mid-north", "north"),
                                            include.lowest = TRUE)

site_check_with_cv_er$lon_category <- cut(site_check_with_cv_er$Longitude,
                                            breaks = quantile(site_check_with_cv_er$Longitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("west", "west-central", "east-central", "east"),
                                            include.lowest = TRUE)


hist1<- ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "Mean ER", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "Mean ER", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "Mean ER", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

#### Histograms of CV GPP by location:

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}

hist1<- ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "CV GPP", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "CV GPP", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "CV GPP", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}

hist1<- ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "CV ER", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "CV ER", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "CV ER", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=12}

# Create the map plot with color based on CV of GPP
coords_map <- ggmap(map) +
  geom_point(data = site_check_with_cv, aes(x = Longitude, y = Latitude, color = cv_intra_annual), size = 3) +
  geom_label_repel(data = site_check_with_cv, aes(x = Longitude, y = Latitude, label = site, color = cv_intra_annual), fontface = "bold", size = 5) +
  xlab("Longitude") + ylab("Latitude") + theme_bw() +
  scale_color_viridis_c(option = "viridis", name = "Intra-annual CV") +  # Use a continuous color scale for CV
  # Add the scale bar
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black", face = "bold"),
    panel.border = element_rect(size = 1.5, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank()
  )

# Print the map
coords_map

```

#### **Q1: Is the degree of intra-annual variation in lake or stream metabolism is a function of regional location in the US/ northern hemisphere?**

#### **A:** Possibly not in the streams.

-   Might be worth scaling GPP by catchment size/ or stream size first.

```{r, warning=F, size = 'small', echo=F, include=F}

library(dplyr)

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual <- river_metab_filled %>%
  mutate(month= month(date)) %>%
  group_by(site, year, month) %>%
  summarise(
    monthly_avg_GPP = mean(GPP_filled, na.rm = TRUE),
    monthly_cv_GPP = sd(GPP_filled, na.rm = TRUE) / mean(GPP_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()



# Summarize intra-annual CV for each year
intra_annual_variation <- data_intra_annual %>%
  group_by(year) %>%
  summarise(
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )
```


#### Load Lake productivity data:

##### Ladwig et al. NTL 2008-2021

- NEP in milligramPerMeterCubedPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Ladwig_NTL_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/model_output.csv")%>%
  mutate(date=as.Date(datetime),
         year=year(date),
         jday= yday(date)) %>%
  filter(date> as.Date("2007-12-31") & date< as.Date("2022-01-01")) %>%
  mutate(NEP_gm3d = c(NEP*0.001)) %>%
    mutate(source= "Ladwig") %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source) 

head(Ladwig_NTL_dat_metab)
```


##### Rabaey et al. 

- GPP/ER/NEP in grams O2 per m3 per day

```{r, warning=F, size = 'small', echo=T, include=T}
Rabaey_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metabolism_daily_summaries.csv")%>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>% 
  mutate(source= "Rabaey") %>%
  dplyr::select(site="name", date, year, jday, NEP_gm3d ="nep", source) 

head(Rabaey_dat_metab)
```




##### Oleksy et al. 

- GPP/ER in milligramsO2PerLiterPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Oleksy_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/daily_metabolism.csv") %>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Oleksy",
           NEP = c(GPP_mgO2L-ER_mgO2L),
           NEP_g =c(NEP * 0.001),
           NEP_gm3d =c(NEP * 0.001)) %>%
  dplyr::select(site="lakeName", date, year, jday, NEP_gm3d, source)

head(Oleksy_dat_metab)
```



##### Lotting et al. 

- Production (millimolscarbonpercubicmeterperday) estimated from high frequency dissolved oxygen time series.

-  1 mmol of carbon is associated with the production or consumption of 1 mmol of O2 so: 

    millimolscarbonpercubicmeterperday * 32.00 = milligramPerMeterCubedPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Lotting_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/lottig_etal_data.csv") %>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Lotting",
           NEP_mg =c(o2_pp_mmolcm3d * 32),
           NEP_gm3d =c(NEP_mg * 0.001)) %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source)

head(Lotting_dat_metab)
```


```{r, warning=F, size = 'small', echo=F, include=F}
lake_metab <- rbind(Ladwig_NTL_dat_metab, Rabaey_dat_metab, Oleksy_dat_metab, Lotting_dat_metab)
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=8}
ggplot(lake_metab%>%filter(source=="Ladwig"), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Ladwig data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=8}
ggplot(lake_metab%>%filter(source=="Rabaey") %>%filter(year>2007), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Rabaey data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=8}
ggplot(lake_metab%>%filter(source=="Lotting"), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Ladwig data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
ggplot(lake_metab%>%filter(source=="Oleksy") %>%filter(year>2007), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Oleksy data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```

###

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(geosphere)


time_df <- river_metab_filled%>%
  left_join(site_check_with_cv, by=c("site")) # %>%  left_join(PRISM_data)


#  Extract coordinates for the reference site (nwis_01400500)
# farthest east site
reference_site <- time_df[time_df$site == "nwis_01400500", ] 
reference_site <- reference_site[1, ]

distances <- distVincentySphere(
  p1 = c(reference_site$Longitude, reference_site$Latitude), # nwis_01400500
  p2 = cbind(locations$Longitude, locations$Latitude)
)

distance_data <- data.frame(
  site = locations$site,
  distance_to_ref = distances*0.001 # km
)


distance_data <- distance_data %>%
  mutate(distance_to_ref = ifelse(site == "nwis_01400500", 0, distance_to_ref))

# Merge distance data with river metabolism data
river_metab_f <- time_df %>%
  left_join(distance_data, by = "site")

# Define distance bands
distance_bands <- seq(0, max(distance_data$distance_to_ref, na.rm = TRUE), by = 50)

# Function to calculate the correlation for each distance band
correlation_by_distance <- function(data, bands) {
  cor_results <- data.frame(distance_band = numeric(), correlation = numeric())
  
  # Get the reference site's GPP data
  ref_data <- data %>% filter(distance_to_ref == 0)
  
  if (nrow(ref_data) == 0) {
    stop("No reference site found with distance_to_ref == 0.")
  }
  
  for (i in 1:(length(bands) - 1)) {
    # Subset the data for the current distance band
    subset_data <- data %>%
      filter(distance_to_ref >= bands[i] & distance_to_ref < bands[i+1])
    
    # Calculate the correlation for the subset (if there are enough data points)
    if (nrow(subset_data) > 1) {
      corr <- cor(subset_data$GPP, subset_data$distance_to_ref, use = "complete.obs")
    } else {
      corr <- NA
    }
    
    # Store results
    cor_results <- rbind(cor_results, data.frame(
      distance_band = (bands[i] + bands[i+1]) / 2, 
      correlation = corr
    ))
  }
  
  return(cor_results)
}

# Compute the correlations by distance band
cor_results <- correlation_by_distance(river_metab_f, distance_bands)
cor_results


```

#### streams don't appear to be synchonous with one another 

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}


ggp_spat_synch <- ggplot(cor_results, aes(x = distance_band, y = correlation)) +
  geom_smooth(se=F, method="lm")+
  geom_point(color = "red") +  # Red points for each correlation
  theme_minimal() +
  geom_hline(yintercept = 0) +
  labs(
    title = "Correlation of GPP vs Distance",
    x = "Distance (km)",
    y = "Correlation Coefficient"
  ) +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

ggp_spat_synch


```


### temp sychn


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(geosphere)

#  Extract coordinates for the reference site (nwis_01400500)
# farthest east site
reference_site <- time_df[time_df$site == "nwis_01400500", ] 
reference_site <- reference_site[1, ]

distances <- distVincentySphere(
  p1 = c(reference_site$Longitude, reference_site$Latitude), # nwis_01400500
  p2 = cbind(locations$Longitude, locations$Latitude)
)

distance_data <- data.frame(
  site = locations$site,
  distance_to_ref = distances*0.001 # km
)


distance_data <- distance_data %>%
  mutate(distance_to_ref = ifelse(site == "nwis_01400500", 0, distance_to_ref))

# Merge distance data with river metabolism data
river_metab_f <- time_df %>%
  left_join(distance_data, by = "site")

# Define distance bands
distance_bands <- seq(0, max(distance_data$distance_to_ref, na.rm = TRUE), by = 50)

# Function to calculate the correlation for each distance band
correlation_by_distance <- function(data, bands) {
  cor_results <- data.frame(distance_band = numeric(), correlation = numeric())
  
  # Get the reference site's GPP data
  ref_data <- data %>% filter(distance_to_ref == 0)
  
  if (nrow(ref_data) == 0) {
    stop("No reference site found with distance_to_ref == 0.")
  }
  
  for (i in 1:(length(bands) - 1)) {
    # Subset the data for the current distance band
    subset_data <- data %>%
      filter(distance_to_ref >= bands[i] & distance_to_ref < bands[i+1])
    
    # Calculate the correlation for the subset (if there are enough data points)
    if (nrow(subset_data) > 1) {
      corr <- cor(subset_data$discharge, subset_data$distance_to_ref, use = "complete.obs")
    } else {
      corr <- NA
    }
    
    # Store results
    cor_results <- rbind(cor_results, data.frame(
      distance_band = (bands[i] + bands[i+1]) / 2, 
      correlation = corr
    ))
  }
  
  return(cor_results)
}

# Compute the correlations by distance band
cor_results <- correlation_by_distance(river_metab_f, distance_bands)
cor_results


```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}


ggp_spat_synch <- ggplot(cor_results, aes(x = distance_band, y = correlation)) +
  geom_smooth(se=F, method="lm")+
  geom_point(color = "red") +  # Red points for each correlation
  theme_minimal() +
  geom_hline(yintercept = 0) +
  labs(
    title = "Correlation of GPP vs Distance",
    x = "Distance (km)",
    y = "Correlation Coefficient"
  ) +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

ggp_spat_synch


```



#### First exploration into Moran effect 

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
# Load required libraries
library(ncf)        # For spatial covariance function
library(ggplot2)    # For visualization
library(dplyr)      # For data manipulation
library(geosphere)  # For calculating geodesic distances
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
set.seed(1234)

x <- expand.grid(1:20, 1:5)[, 1]
y <- expand.grid(1:20, 1:5)[, 2]

# The first column (x) represents site locations, and 
# The second column (y) represents time points. 
# This creates a total of 100 site-time combinations.

Z <- cbind(rmvn.spa(x = x, y = y, p = 2, method = "gaus"), #rmvn.spa() generates random variates from a spatially autocorrelated multivariate normal distribution.
           rmvn.spa(x = x, y = y, p = 2, method = "gaus"))
# The output (Z) is a dataset of spatially correlated values, simulated for 100 sites over 2 year

fit0 <- Sncf(x = x, y = y, z = Z, resamp = 1000, quiet = T)

plot(fit0)

m0.summary <- summary(fit0)
m0.summary

# Simulate 25% Missingness
Z.m25 <- Z
Z.m25[sample(length(x), 0.25 * length(x))] <- NA


for (i in 1:nrow(Z)) {
  idx.tmp <- which(is.na(Z.m25[i, ]))
  if (length(idx.tmp > 0)) {
    Z.m25[i, idx.tmp] <- mean(Z.m25[i, ], na.rm = TRUE)
  }
}

fit25 <- Sncf(x = x, y = y, z = Z.m25, resamp = 1000, quiet = T)
plot(fit25)
m25.summary <- summary(fit25)
m25.summary


```


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
# Simulate 0% Missingness Data at 100 Sites Over 2 Years
# Load your dataset
df <- site_check_with_cv

#data_intra_annual_mon

data_intra_annual


river_metab_f <- river_metab_filled %>%
  mutate(Name = sub("^nwis_", "", site))

PRISM_dataf <- PRISM_data %>%
  mutate(Name = str_pad(Name, width = 8, side = "left", pad = "0"))

PRISM_data_month<- PRISM_data%>%
  mutate(month=month(Date),
         year= year(Date)) %>%
  group_by(Name, month, year) %>%
    summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

unique(PRISM_dataf$Name)
str(PRISM_dataf)
str(river_metab_f)


unique(river_metab_f$Name)
  
df<- river_metab_f%>%
  left_join(PRISM_data, by=c("Name","date"="Date"))

df <- na.omit(df)

unique(df$Name)
df <- as.data.frame(df)
str(df)

df <- spc_tbl %>%
  left_join(river_metab_f, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")%>%
  left
```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}

str(df)

df_clean <- df %>%
  filter(!is.na(mu_intra_annual), !is.na(Longitude), !is.na(Latitude))

# Compute Euclidean distances between site locations
distance_matrix <- distm(df_clean[, c("Longitude", "Latitude")], fun = distHaversine)/1000   # Convert to km


response_variable <- as.numeric(df_clean$monthly_avg_GPP)
df_clean$Longitude <- as.numeric(df_clean$Longitude)
df_clean$Latitude <- as.numeric(df_clean$Latitude)

df_clean<- as.data.frame(df_clean)

# Check for NAs in relevant columns
sum(is.na(df_clean$Longitude))  # Should be 0
sum(is.na(df_clean$Latitude))   # Should be 0
sum(is.na(response_variable))  # Should be 0

length(df_clean$Longitude)
length(df_clean$Latitude)
length(response_variable)

sum(is.infinite(df_clean$Longitude))   # Check for infinities in Longitude
sum(is.infinite(df_clean$Latitude))    # Check for infinities in Latitude
sum(is.infinite(response_variable))    # Check for infinities in response_variable

sum(is.nan(df_clean$Longitude))        # Check for NaNs in Longitude
sum(is.nan(df_clean$Latitude))         # Check for NaNs in Latitude
sum(is.nan(response_variable))         # Check for NaNs in response_variable


sync_result <- ncf::Sncf(
  x = distance_matrix, 
  y = df_clean$Latitude, 
  z = response_variable, 
  resamp = 50,  
  latlon = TRUE
)


sync_result_subset <- ncf::Sncf(
  x = df_clean$Longitude[1:10], 
  y = df_clean$Latitude[1:10], 
  z = response_variable[1:10], 
  resamp = 50,  
  latlon = TRUE
)

df_clean_unique <- df_clean %>% distinct(Longitude, Latitude, .keep_all = TRUE)

response_variable<- df_clean_unique$monthly_avg_GPP

sync_result_subset <- ncf::Sncf(
  x = df_clean_unique$Longitude[1:10], 
  y = df_clean_unique$Latitude[1:10], 
  z = response_variable[1:10], 
  resamp = 50,  
  latlon = TRUE
)

# Compute nonparametric spatial covariance function
synchrony_model <- Sncf(
  x = df_clean$Longitude, 
  y = df_clean$Latitude, 
  z = df_clean$mu_intra_annual
)

# Extract key metrics
initial_correlation <- synchrony_model$y[1]  # Correlation at shortest distance
mean_correlation <- mean(synchrony_model$y, na.rm = TRUE)  # Average spatial correlation
synchrony_extent <- max(synchrony_model$x[synchrony_model$y > mean_correlation])  # Distance where covariance remains above avg

# Plot spline correlogram
plot(synchrony_model, main = "Spline Correlogram of Spatial Synchrony")

# Alternative ggplot visualization
synchrony_df <- data.frame(
  Distance = synchrony_model$x,
  Correlation = synchrony_model$y
)

ggplot(synchrony_df, aes(x = Distance, y = Correlation)) +
  geom_line() +
  geom_hline(yintercept = mean_correlation, linetype = "dashed", color = "red") +
  labs(title = "Spatial Synchrony Spline Correlogram",
       x = "Euclidean Distance (km)",
       y = "Spatial Correlation") +
  theme_minimal()

```


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}

df <- site_check_with_cv

#data_intra_annual_mon

data_intra_annual


river_metab_f <- river_metab_filled %>%
  left_join(site_check_with_cv, by=c("site"))

river_metab_f<- na.omit(river_metab_f)

library(geostats)


river_metab_f<- as.data.frame(river_metab_f)
str(river_metab_f)

mantel.test(river_metab_f$Latitude, river_metab_f$Longitude,
              river_metab_f$GPP, M1 = NULL, M2 = NULL, resamp = 999, latlon = FALSE, 
            quiet = FALSE)


river_metab_f_small <- river_metab_f[sample(nrow(river_metab_f), 1000), ]
mantel.test(river_metab_f_small$Latitude, river_metab_f_small$Longitude,
            river_metab_f_small$GPP, M1 = NULL, M2 = NULL, resamp = 999, 
            latlon = FALSE, quiet = FALSE)


plot()
```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(ggplot2)
library(corrplot)
library(tidyr)
library(ggpubr)
library(geosphere)

# Assuming your data is stored in 'river_metab_f'
# 1. First, create a new column for distance based on latitude and longitude (e.g., distance from a reference point)
# Here, let's assume (lat_ref, lon_ref) is a point of reference for distance calculation

lat_ref <- 40.0  # Example reference latitude
lon_ref <- -75.0  # Example reference longitude

# Calculate the distance from the reference point for each site using Haversine formula
river_metab_f <- river_metab_f %>%
  mutate(distance = distVincentySphere(cbind(Longitude, Latitude), c(lon_ref, lat_ref)))

# 2. Separate data into two periods: 1959–1980 and 1989–2010
period_1 <- river_metab_f %>%
  filter(year >= 2008 & year <= 2014)

period_2 <- river_metab_f %>%
  filter(year >= 2015 & year <= 2021)

# 3. Compute correlations between GPP and distance for each period
cor_period_1 <- cor(period_1$GPP, period_1$distance, use = "complete.obs")
cor_period_2 <- cor(period_2$GPP, period_2$distance, use = "complete.obs")

# 4. To generate confidence bands, use bootstrapping to estimate the confidence intervals for each correlation
bootstrap_cor <- function(data, n_bootstrap = 1000) {
  boot_cor <- replicate(n_bootstrap, {
    sample_data <- data[sample(nrow(data), replace = TRUE), ]
    cor(sample_data$GPP, sample_data$distance, use = "complete.obs")
  })
  ci_lower <- quantile(boot_cor, 0.025)
  ci_upper <- quantile(boot_cor, 0.975)
  mean_cor <- mean(boot_cor)
  return(c(mean_cor, ci_lower, ci_upper))
}

# Bootstrap confidence intervals for each period
cor_ci_period_1 <- bootstrap_cor(period_1)
cor_ci_period_2 <- bootstrap_cor(period_2)

# 5. Now, create the plot with confidence bands and the correlation vs. distance plot
# Combine results into a data frame for plotting
cor_results <- data.frame(
  period = c("2008-2014", "2015-2021"),
  correlation = c(cor_period_1, cor_period_2),
  lower_ci = c(cor_ci_period_1[2], cor_ci_period_2[2]),
  upper_ci = c(cor_ci_period_1[3], cor_ci_period_2[3])
)

# Plot the results
ggplot(cor_results, aes(x = period, y = correlation, color = period)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.1) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "Correlation vs Distance for GPP",
    x = "Time Period",
    y = "Correlation Coefficient"
  )


# Load necessary libraries
library(dplyr)
library(ggplot2)
library(geosphere)

# Assuming your data is stored in 'river_metab_f'
# 1. Calculate the distance from a reference point (as before)
lat_ref <- 40.0  # Example reference latitude
lon_ref <- -75.0  # Example reference longitude

river_metab_f <- river_metab_f %>%
  mutate(distance = distVincentySphere(cbind(Longitude, Latitude), c(lon_ref, lat_ref)))

# 2. Define distance bands (e.g., 0-50 km, 50-100 km, 100-150 km, etc.)
distance_bands <- seq(0, max(river_metab_f$distance), by = 50)  # Adjust 'by' for the band width

# 3. Create a function to calculate the correlation for each distance band
correlation_by_distance <- function(data, bands) {
  cor_results <- data.frame(distance_band = numeric(), correlation = numeric())
  
  for (i in 1:(length(bands) - 1)) {
    # Subset the data for the current distance band
    subset_data <- data %>%
      filter(distance >= bands[i] & distance < bands[i+1])
    
    # Calculate the correlation for the subset (if there are enough data points)
    if (nrow(subset_data) > 1) {
      corr <- cor(subset_data$GPP, subset_data$distance, use = "complete.obs")
      cor_results <- rbind(cor_results, data.frame(distance_band = (bands[i] + bands[i+1])/2, correlation = corr))
    }
  }
  
  return(cor_results)
}

# 4. Compute the correlations by distance band
cor_results <- correlation_by_distance(river_metab_f, distance_bands)

# 5. Plot the results
ggplot(cor_results, aes(x = distance_band, y = correlation)) +
  geom_line(color = "blue") +  # Blue line for correlation
  geom_point(color = "red") +  # Red points for each correlation
  theme_minimal() +
  labs(
    title = "Correlation of GPP vs Distance",
    x = "Distance (km)",
    y = "Correlation Coefficient"
  ) +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )


```
Brute for a distnace matrix 

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
# Install required packages
install.packages("ncf")
install.packages("geosphere")

# Load libraries
library(ncf)
library(geosphere)
library(dplyr)
library(tidyverse)

locations <- river_metab_f %>% 
  select(Latitude, Longitude) %>% 
  distinct()

# Create a distance matrix using geosphere's distVincentySphere function
dist_matrix <- distVincentySphere(cbind(locations$Longitude, locations$Latitude))


# Number of distances in the vector
num_distances <- length(dist_matrix)

# Calculate the number of sites
num_sites <- nrow(locations)

# Check if the number of distances matches the expected number of pairs
expected_num_distances <- num_sites * (num_sites - 1) / 2

if (num_distances != expected_num_distances) {
  stop("The number of distances does not match the expected number based on the number of sites.")
}

# Convert the vector of distances to a square distance matrix
dist_matrix_square <- matrix(NA, nrow = num_sites, ncol = num_sites)

# Fill the upper triangle with the values from the dist_matrix vector
dist_matrix_square[upper.tri(dist_matrix_square)] <- dist_matrix

# Fill the lower triangle by symmetry
dist_matrix_square[lower.tri(dist_matrix_square)] <- t(dist_matrix_square)[lower.tri(dist_matrix_square)]

# Check the result
dist_matrix_square


# Check the distance matrix dimensions
dim(dist_matrix_square)  # This should be a square matrix

temp_data <- river_metab_f %>%
  select(site, year, GPP)

dist_vec <- dist_matrix_square[upper.tri(dist_matrix_square)]  # Distances between pairs

df <- data.frame(distance = dist_vec, gpp = temp_data$GPP)


M_b <- cor(temp_data$GPP, temp_data$GPP)

```


