---
title: 'Dataset organization for stream metab and lake metab synchrony analysis'
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
# setwd("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis")

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB", "plotly", "MASS", "EnvStats",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)

```

## **Questions**

[Draft questions]

#### **Q1:** Is the degree of intra-annual variation in lake or stream metabolism is a function of regional location in the US/ northern hemisphere?

-   ... New interest in K\~Q or K\~ residence time (for lakes)

#### **Q 2:** Is the magnitude of that intra-annual variation greater within streams or within lakes for that region?

-   Need to define regions more clearly?

-   Could use a euclidean distance metric for lat and long as a
    continuous response?

#### **Q 3:** Is the magnitude of that intra- annual variation greater within streams or within lakes for that region?

#### **Q 4:** Where is the effect of past variation in GPP or ER the strongest (in either habitat type lakes or streams and/or based on geographic position, in laces with more mild seasonal variance southeast versus inter mountain west)?

### Background:

#### (1) Stream Data:

**Marzolf et al. 2024**
[https://aslopubs.onlinelibrary.wiley.com/doi/10.1002/lol2.10407)-
Phenology of gross primary productivity in rivers displays high
variability within years but stability across
years](https://aslopubs.onlinelibrary.wiley.com/doi/10.1002/lol2.10407)

Looking at workflow from:
<https://github.com/nmarzolf91/usgs_metabolism_project>

**Appling et al. 2018** [The metabolic regimes of 356 rivers in the
United States](https://www.nature.com/articles/sdata2018292)

#### (2) Lake Data:

**Ladwig et al** [Modeling
dataset:](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-ntl.402.2)
Long-term Change in Metabolism Phenology across North-Temperate Lakes,
Wisconsin, USA 1979-2019

-   8 lakes 2008 - 2021

-   Only NEP in milligramPerMeterCubedPerDay

**Rabaey et al.** [Freshwater biogeochemical
hotspots](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1531.2):
High primary production and ecosystem respiration in shallow waterbodies

-   38 lakes 1 year

- GPP/ER/NEP in grams O2 per m3 per day

**Oleksy et al.** [Metabolism
dataset](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1541.2):
one year of high-frequency temperature, dissolved oxygen, wind,
photosynthetically active radiation observations and low-frequency
nutrient data for 58 lakes in the Global Lake Ecological Observatory
Network

-   58 lakes 1 year ... some overlap with the Ladwig time series

- GPP/ER in milligramsO2PerLiterPerDay

**Corman et al.** 

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
### Marzolf workflow parameters 
# define constants and axis labels
metab_units_area <- (expression(paste('GPP (g C', m^-2,' ',y^-1,')', sep = ' ')))

low_p <- '#103801'
high_p <- '#43fa00'

# quotient to convert GPP from O2 to C
# g C = g O2 * (1 mol O2/32 g O2) * (2 mol O/1 mol O2) * (1 mol C/1 mol O) * (12 g C/1 mol 

rq <- (1/32)*(2/1)*(1/1)*(12/1)

## check this to see what a better assumption would be ?
rq_e <- 1

# where to save figures
save_figs <- TRUE
dpi <- 400

```

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
## Read in data 

metab_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_TS.rds")
# str(metab_ts)

dat_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/TS_SPC_pH_FNU.rds")
# str(dat_ts)
# summary(dat_ts)
# unique(dat_ts$site_name)

metab_meta_dat <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata.rds")
# str(metab_ts)


marz_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/1_site_info.csv")
# 63 different sites 

Ladwig_NTL_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/Ladwig_NTL_lake_cords.csv")%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="ladwig_NTL_lake")

Oleksy_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/lake_descriptions.csv") %>%
  filter(data_request=="yes")%>%
  dplyr::select(Site.Number ="lakeName", Latitude="lat_dd", Longitude="long_dd") %>%
  mutate(source="oleksy_lake")

Oleksy_good_sites <- unique(Oleksy_dat_metadf$Site.Number)

Rabaey_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metadata.csv")%>%
  dplyr::select(Site.Number ="pond_name", Latitude="latitude", Longitude="longitude") %>%
  mutate(source="rabaey_lake")

corman_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/load_metab_results/corman_lake_cords.csv")%>%
  filter(country=="United States" | country== "Canada") %>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="corman_lake")

lotting_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/Lotting_lake_cords.csv")%>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="lotting_lake")


```


```{r, warning=F, size = 'small', echo=F, include=F}

dat_ts_clean <- dat_ts%>%
  filter(GPP.Rhat < 1.05 &  GPP >0 & ER <0)

# unique(dat_ts_clean$long_name)

marz_dat_names <- marz_dat%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="marzolf_stream")

dat_ts_names <-metab_meta_dat %>%
   dplyr::select(Site.Number = "site_no", Latitude = "dec_lat_va", Longitude= "dec_long_va") %>%
  mutate(source="appling_stream")

## quick site check
site_check <- rbind(marz_dat_names, 
                  #dat_ts_names,
                  Ladwig_NTL_dat_metadf, 
                   Oleksy_dat_metadf,
                   Rabaey_dat_metadf, #, 
                   # corman_dat_metadf, 
                    lotting_dat_metadf)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
## map packages 

library(tidyverse)
library(ggrepel)
library(ggmap)
library(ggsci)
library(scales)

library(dplyr)
library(ggplot2)
library(ggrepel)


#devtools::install_github("briatte/tidykml")
library(tidykml)
library(sf)
```


```{r, warning=F, size = 'small', echo=F, include=T, include=F}

# Define the bounding box coordinates add some wiggle room:

range(na.omit(site_check$Longitude))
range(site_check$Latitude)
# 
# Define the bounding box coordinates
bbox <- c(left = -128.300, bottom = 22.50, right = -63.800, top = 51.50)

# Retrieve terrain map layer using get_stamenmap()
map <- get_stadiamap(bbox = bbox, zoom = 5, maptype = 'stamen_terrain')
# need api run for the first time

#library(ggmap)
# Plot the terrain map
stamen_map <- ggmap(map)
```

### Map of stream data locations

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=12}

coords_map <- ggmap(map) +
  geom_point(data = site_check, aes(x = Longitude, y = Latitude, color = source, shape =source), size = 2) +
  geom_label_repel(data = site_check, aes(x = Longitude, y = Latitude, label = Site.Number,  fill = source), color= "black", fontface = "bold", size = 2, max.overlaps = 20) +
  xlab("Longitude") + ylab("Latitude") + theme_bw() +
  scale_color_viridis_d(option = "viridis",  name = "source")  +
 scale_fill_viridis_d(option = "viridis",  name = "source", alpha = 0.5)  +

  # Add the scale bar
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black", face = "bold"),
    panel.border = element_rect(size = 1.5, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank()
  )
coords_map
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig_Ch3_map.png", plot = coords_map, width = 12, height = 8, units = "in")

```

There are some site overlaps in the lake data sets for lake locations.

#### Load Marzolf GPP data:

```{r, warning=F, size = 'small', echo=T, include=T}
river_metab <- readr::read_csv('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/5_usgs_metabolism.csv')

head(river_metab)
```

#### Gap fill

'fill_and_normalize_metab.R' is a modified function
from'code/metabolism_synthesis/Part II- Filtering and gap filling.Rmd
uses the function fillMiss3.R

**Make sure** to double check the 'summary_site-year_NAs.csv' save path
in the fxn

-   Extracts unique sites and years from the input data.

-   Tracks missing data for each site-year.

-   Pads the dataset to ensure a full-year time series.

-   Identifies missing data patterns and determines if a year is
    complete.

-   Applies a gap-filling algorithm (if \<40% of data is missing).

-   Outputs a cleaned and gap-filled dataset for further analysis.



```{r, warning=F, size = 'small', echo=F, include=T}
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/code_fxns/fill_and_normalize_metab.R')

```

```{r, warning=F, size = 'small', echo=T, include=T}
river_metab_filled <- fill_and_normalize_metab(river_metab)
head(river_metab_filled)
```

##### Check the scope of the infill

```{r, warning=F, size = 'small', echo=T, include=T}
n_obs_gpp <- length(river_metab_filled$GPP_filled)
n_obs_gpp

n_sites <- length(unique(river_metab_filled %>%
                           dplyr::filter(!is.na(site)) %>%
                           dplyr::pull(site)))
n_sites
```

```{r, warning=F, size = 'small', echo=T, include=T}
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/code_fxns/fill_and_normalize_metab_KAL.R')

river_metab_filled_ER <- fill_and_normalize_metab(river_metab)
head(river_metab_filled_ER)
```

```{r, warning=F, size = 'small', echo=T, include=T}

n_obs_er <- length(river_metab_filled_ER$ER_filled)
n_obs_er

n_sites <- length(unique(river_metab_filled_ER %>%
                           dplyr::filter(!is.na(site)) %>%
                           dplyr::pull(site)))
n_sites
```

### Calculate cumulative daily GPP and total annual GPP to get a variance dynamics:

```{r, warning=F, size = 'small', echo=F, include=F}
table_s1 <- readr::read_csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/1_site_info.csv")
```

```{r, warning=F, size = 'small', echo=F, include=F}
# calculate site-year cumulative GPP
river_metab_cumsum <- river_metab_filled %>%
  dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::mutate(jday = lubridate::yday(date),
                GPP_C = GPP_filled*rq,
                GPP_C_clean = replace(GPP_C, GPP_C < 0, 0), # replace negative values with 0 in GPP_C_cumsum
                GPP_C_cumsum = cumsum(GPP_C_clean))


# calculate annual GPP for each site-year
metab_all_annual <- river_metab_cumsum %>%
#dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::summarise(GPP_daily_mean = mean(GPP_C_clean, na.rm = TRUE),
                   GPP_ann_sum = sum(GPP_C_clean, na.rm = TRUE),
                   GPP_daily_cv = EnvStats::cv(GPP_C_clean, na.rm = TRUE, method = 'l.moments')*100,
                   Q_daily_mean = mean(discharge, na.rm = TRUE),
                   Q_daily_cv = EnvStats::cv(discharge, na.rm = TRUE, method = 'l.moments')*100,
                   temp_daily_mean = mean(temp.water, na.rm = TRUE),
                   temp_daily_cv = EnvStats::cv(temp.water,na.rm = TRUE, method = 'l.moments')*100) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`,
                            clean_name =`Name`) ,
                   'clean_site_num')
```

```{r, warning=F, size = 'small', echo=T, include=F}
##### Calculate median annual and CV for each site time series

site_mean_ann_gpp <- metab_all_annual %>%
  dplyr::group_by(site, clean_name) %>%
  dplyr::summarise(median_ann_GPP = median(GPP_ann_sum, na.rm = TRUE),
                   sd_ann_GPP = sd(GPP_ann_sum, na.rm = TRUE),
                   min_ann_GPP = min(GPP_ann_sum, na.rm = TRUE),
                   max_ann_GPP = max(GPP_ann_sum, na.rm = TRUE),
                   GPP_percent_diff = ((max_ann_GPP- min_ann_GPP)/min_ann_GPP)*100,
                   GPP_ann_cv = EnvStats::cv(GPP_ann_sum, na.rm = TRUE, method = 'l.moments')*100,
                   mean_ann_temp = mean(temp_daily_mean, na.rm = TRUE),
                   mean_sd_temp = sd(temp_daily_mean, na.rm = TRUE),
                   cv_mean_temp = EnvStats::cv(temp_daily_mean, method = 'l.moments')*100,
                   median_ann_discharge = median(Q_daily_mean, na.rm = TRUE),
                   median_sd_discharge = sd(Q_daily_mean, na.rm = TRUE),
                   cv_median_discharge = EnvStats::cv(Q_daily_mean, na.rm = TRUE, method ='l.moments')*100
                   ) %>%
  dplyr::arrange(desc(max_ann_GPP)) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`),'clean_site_num')

head(site_mean_ann_gpp)

```

##### Cumulative site-year GPP as Fig 1 in the paper

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}

river_metab_cumsum_names <- river_metab_cumsum %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`) ,'clean_site_num') %>%
  dplyr::rename(clean_name =`Name`)

river_metab_cumsum_names$clean_name <-
  suppressWarnings(forcats::fct_relevel(river_metab_cumsum_names$clean_name,site_mean_ann_gpp %>%
                                          dplyr::pull(clean_name)))

y_axis_title <- bquote(Cumulative~ .(metab_units_area))
# subset of 9 sites to show as figure 1 (for size/space concerns)
highest_4 <- site_mean_ann_gpp %>%
  arrange(desc(median_ann_GPP)) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

lowest_4 <- site_mean_ann_gpp %>%
  arrange(median_ann_GPP) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

extremes <- c(lowest_4, highest_4)

fig_1 <- ggplot(data = river_metab_cumsum_names %>%
                  dplyr::filter(site %in% extremes),
                aes(x = jday, y = GPP_C_cumsum,
                    group = interaction(year, site),
                    color = GPP_C_cumsum))+
  geom_line(show.legend = FALSE) +
  facet_wrap(forcats::fct_relevel(clean_name,
                                  site_mean_ann_gpp %>%
                                    dplyr::filter(site %in% extremes) %>%
                                    dplyr::pull(clean_name))~., scales = 'free_y',
             ncol = 4) +
  scale_color_gradient(low = low_p, high = high_p) +
  scale_y_continuous(breaks = function(y) {
    max_value <- max(y, na.rm = TRUE)*0.9
    tick_marks <- seq(0, round(max_value, digits =-1), length.out = 3)
    return(tick_marks)
  }
  ) +
  labs(x = 'Day of Year',
       y = expression(paste('Cumulative GPP (g C', m^-2, '', y^-1,')'))) +
       theme(strip.text = element_text(size = 7),
             strip.background = element_rect(fill = 'white')) +
  theme_classic()

fig_1
```

#### Same thing but for cummulative ER?

```{r, warning=F, size = 'small', echo=F, include=F}
# calculate site-year cumulative ER
river_metab_cumsum_ER <- river_metab_filled_ER %>%
  dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::mutate(jday = lubridate::yday(date),
                ER_C = (ER_filled*-1) * rq_e, # absolute value of ER
                ER_C_clean = replace(ER_C, ER_C < 0, 0), # replace negative values with 0 in ER_C_cumsum
                ER_C_cumsum = cumsum(ER_C_clean))


# calculate annual ER for each site-year
metab_all_annual_er <- river_metab_cumsum_ER %>%
#dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(site, year) %>%
  dplyr::summarise(ER_daily_mean = mean(ER_C_clean, na.rm = TRUE),
                   ER_ann_sum = sum(ER_C_clean, na.rm = TRUE),
                   ER_daily_cv = EnvStats::cv(ER_C_clean, na.rm = TRUE, method = 'l.moments')*100,
                   Q_daily_mean = mean(discharge, na.rm = TRUE),
                   Q_daily_cv = EnvStats::cv(discharge, na.rm = TRUE, method = 'l.moments')*100,
                   temp_daily_mean = mean(temp.water, na.rm = TRUE),
                   temp_daily_cv = EnvStats::cv(temp.water,na.rm = TRUE, method = 'l.moments')*100) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`,
                            clean_name =`Name`) ,
                   'clean_site_num')
```

```{r, warning=F, size = 'small', echo=F, include=F}
##### Calculate median annual and CV for each site time series

site_mean_ann_er <- metab_all_annual_er %>%
  dplyr::group_by(site, clean_name) %>%
  dplyr::summarise(median_ann_ER = median(ER_ann_sum, na.rm = TRUE),
                   sd_ann_ER = sd(ER_ann_sum, na.rm = TRUE),
                   min_ann_ER = min(ER_ann_sum, na.rm = TRUE),
                   max_ann_ER = max(ER_ann_sum, na.rm = TRUE),
                   ER_percent_diff = ((max_ann_ER- min_ann_ER)/min_ann_ER)*100,
                   ER_ann_cv = EnvStats::cv(ER_ann_sum, na.rm = TRUE, method = 'l.moments')*100,
                   mean_ann_temp = mean(temp_daily_mean, na.rm = TRUE),
                   mean_sd_temp = sd(temp_daily_mean, na.rm = TRUE),
                   cv_mean_temp = EnvStats::cv(temp_daily_mean, method = 'l.moments')*100,
                   median_ann_discharge = median(Q_daily_mean, na.rm = TRUE),
                   median_sd_discharge = sd(Q_daily_mean, na.rm = TRUE),
                   cv_median_discharge = EnvStats::cv(Q_daily_mean, na.rm = TRUE, method ='l.moments')*100
                   ) %>%
  dplyr::arrange(desc(max_ann_ER)) %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`),'clean_site_num')

```

##### Cumulative site-year \|ER\| modfied from Fig 1 in the paper

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}

river_metab_cumsum_names_ER <- river_metab_cumsum_ER %>%
  dplyr::mutate(clean_site_num = as.numeric(substr(site, 6,30))) %>%
  dplyr::left_join(table_s1 %>%
                     rename(clean_site_num =`Site Number`) ,'clean_site_num') %>%
  dplyr::rename(clean_name =`Name`)

river_metab_cumsum_names_ER$clean_name <-
  suppressWarnings(forcats::fct_relevel(river_metab_cumsum_names_ER$clean_name,site_mean_ann_er %>%
                                          dplyr::pull(clean_name)))

y_axis_title <- bquote(Cumulative~ .(metab_units_area))
# subset of 9 sites to show as figure 1 (for size/space concerns)
highest_4 <- site_mean_ann_er %>%
  arrange(desc(median_ann_ER)) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

lowest_4 <- site_mean_ann_er %>%
  arrange(median_ann_ER) %>%
  head(., n = 4) %>%
  dplyr::pull(site)

extremes <- c(lowest_4, highest_4)

fig_1er <- ggplot(data = river_metab_cumsum_names_ER %>%
                  dplyr::filter(site %in% extremes),
                aes(x = jday, y = ER_C_cumsum,
                    group = interaction(year, site),
                    color = ER_C_cumsum))+
  geom_line(show.legend = FALSE) +
  facet_wrap(forcats::fct_relevel(clean_name,
                                  site_mean_ann_er %>%
                                    dplyr::filter(site %in% extremes) %>%
                                    dplyr::pull(clean_name))~., scales = 'free_y',
             ncol = 4) +
  scale_color_gradient(low = low_p, high = high_p) +
  scale_y_continuous(breaks = function(y) {
    max_value <- max(y, na.rm = TRUE)*0.9
    tick_marks <- seq(0, round(max_value, digits =-1), length.out = 3)
    return(tick_marks)
  }
  ) +
  labs(x = 'Day of Year',
       y = expression(paste('Cumulative |ER| (g C', m^-2, '', y^-1,')'))) +
       theme(strip.text = element_text(size = 7),
             strip.background = element_rect(fill = 'white')) +
  theme_classic()

fig_1er
```

##### When are rivers productive?

Where - Marzolf et al - identifies when cumulative GPP (a measure of
ecosystem productivity) reaches specific percentiles (25%, 50%, 75%, and
95%) for each site and year

```{r, warning=F, size = 'small', echo=T, include=T}
df_out <- data.frame(
  site = rep(rep(unique(river_metab_cumsum_names$site), each = 14),times = 4),
  year = rep(rep(unique(river_metab_cumsum_names$year), times = 59),each = 4),
  quantile = rep(c(0.25, 0.5, 0.75, 0.95), times = 14*59))

for(i in 1:nrow(df_out)){
  deets <- df_out[i,]
  site <- deets[,'site']
  year <- deets[,'year']
  quantile <- deets[,'quantile']
  
  df_use <- river_metab_cumsum_names %>%
    dplyr::filter(site %in% !!site,
                  year %in% !!year)
  if(nrow(df_use)== 0)
    next
# Calculate the Maximum Cumulative GPP (GPP_C_cumsum) for the Year:
  cumsum <- max(df_use$GPP_C_cumsum, na.rm = TRUE)
  df_out[i,'GPP_C_cumsum'] <- cumsum

# Normalize the Cumulative GPP to Create a CDF (Cumulative Distribution Function):
  df_math <- df_use %>%
    dplyr::mutate(GPP_C_cumsum_cdf = GPP_C_cumsum/cumsum)
  
# Find the Julian Day (jday) When the GPP Reaches the Specified Quantile:
  jday <- which.min(abs(df_math$GPP_C_cumsum_cdf- quantile))
  df_out[i,'jday'] <- jday
  
  # Extract Corresponding Data for That jday:
  row <- data.frame(df_math %>%
                      dplyr::filter(jday== !!jday))
  df_out[i,'clean_name'] <- row[,'clean_name']
  df_out[i,'date'] <- row[,'date']
  df_out[i,'GPP_C_cumsum'] <- row[,'GPP_C_cumsum']
  df_out[i,'cdf'] <- row[,'GPP_C_cumsum_cdf']
  } # end for loop

# Filter and Fill Missing Data (df_quants):
df_quants <- df_out %>%
  dplyr::filter(!is.na(jday)) %>%
  dplyr::group_by(site) %>%
  tidyr::fill(clean_name, .direction = 'down')

df_quants %>%
  dplyr::group_by(quantile) %>%
  dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
                   jday_sd = sd(jday, na.rm = TRUE),
                   jday_min = min(jday, na.rm = TRUE),
                   jday_max = max(jday, na.rm = TRUE))
```

```{r, warning=F, size = 'small', echo=T, include=T}
df_quants %>%
  dplyr::group_by(quantile, clean_name) %>%
  dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
                   jday_sd = sd(jday, na.rm = TRUE),
                   jday_min = min(jday, na.rm = TRUE),
                   jday_max = max(jday, na.rm = TRUE))

```

#### Trend analysis:

trend analysis on quantiles of cumulative gross primary production
(GPP_C_cumsum) over time, using Sen's slope and the Mann-Kendall test to
determine the direction and significance of trends in the Julian day
(jday) at which specific cumulative GPP quantiles are reached.

```{r, warning=F, size = 'small', echo=T, include=T}
# trend analysis for quantiles
quantile_trends <- df_quants %>%
  dplyr::arrange(site, year) %>%
  dplyr::group_by(site, quantile) %>%
  tidyr::nest() %>%
  # Performing Sen's Slope and Mann-Kendall Trend Tests
  dplyr::mutate(sens = purrr::map(data, ~trend::sens.slope(x = .$jday)),
                mk_test = purrr::map(data, ~trend::mk.test(x = .$jday))) %>%
  # Extracting Trend Statistics
  dplyr::mutate(sens_s = purrr::map_dbl(sens,~.$estimates), 
                sens_p = purrr::map(sens,~.$p.value),
                mk_p = purrr::map_dbl(mk_test,~.$p.value),
                # mk_t = purrr::map_dbl(mk_test, ~.$estimates['tau']),
                sens_sig = ifelse(sens_p <= 0.05, 'significant', 'non-significant'),
                sens_slope = ifelse(sens_s > 0, 'increasing',  'decreasing'),
                mk_sig = ifelse(mk_p <= 0.05, 'significant', 'non-significant')
                )

quantile_sum_sigs <- quantile_trends %>%
  dplyr::ungroup() %>%
  dplyr::select(quantile, site, mk_p, #mk_t,
                mk_sig, #mk_slope,
                sens_s, sens_sig, sens_slope)

quants_n_sig <- quantile_sum_sigs %>%
  dplyr::filter(mk_p <= 0.05) %>%
  dplyr::group_by(quantile, sens_slope) %>%
  summarise(n = n()) %>%
  data.frame()

quantile_trends_sum <- df_quants %>%
  dplyr::left_join(quantile_sum_sigs) %>%
  dplyr::mutate(sens_meaning = dplyr::case_when(sens_slope== 'increasing' ~ 'Later',
                                                sens_slope== 'decreasing' ~ 'Earlier'),
# mk_meaning = case_when(mk_slope =='increasing' ~ 'Later',
# mk_slope =='decreasing' ~ 'Earlier'),
)


```

**sens.slope**: Sen’s slope (sens.slope) estimates the trend in jday
over time (whether the day at which a given quantile is reached is
shifting earlier or later).

**mk.test**: Mann-Kendall test (mk.test) assesses whether there is a
statistically significant trend in jday over time.

#### Visualize the when productivity is highest

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=10, fig.height=6}
library(gghighlight)

quantile_trenders <- quantile_trends_sum %>%
  dplyr::filter(mk_p <= 0.05) %>%
  dplyr::group_by(quantile, clean_name) %>%
  dplyr::filter(row_number()== 1)

fig_3 <- ggplot(quantile_trends_sum, aes(x = year, y = jday, group = site, color = sens_meaning))+
  geom_line(linewidth = 0.8)+
  gghighlight::gghighlight(mk_sig== 'significant',
                           use_direct_label = FALSE, keep_scales = TRUE,
                           unhighlighted_params = list(alpha("grey10", 0.4),
                                                       linewidth = 0.4),
                           calculate_per_facet = TRUE)+
  facet_wrap(.~ quantile, scales = 'free_y', nrow = 2, ncol = 2)+
#scale_y_continuous(limits = c(0, 366))+
  scale_color_manual(name = element_blank(), values = c(high_p, low_p))+ 
  labs(x = 'Year', y = 'Day of Year')+
  theme(legend.title = element_blank()) + theme_bw()

fig_3

```

```{r, warning=F, size = 'small', echo=T, include=F}
# df_out_er <- data.frame(
#   site = rep(rep(unique(river_metab_cumsum_names_ER$site), each = 14),times = 4),
#   year = rep(rep(unique(river_metab_cumsum_names_ER$year), times = 59),each = 4),
#   quantile = rep(c(0.25, 0.5, 0.75, 0.95), times = 14*59))
# 
# for(i in 1:nrow(df_out_er)){
#   deets <- df_out_er[i,]
#   site <- deets[,'site']
#   year <- deets[,'year']
#   quantile <- deets[,'quantile']
#   
#   df_use_er <- river_metab_cumsum_names_ER %>%
#     dplyr::filter(site %in% !!site,
#                   year %in% !!year)
#   if(nrow(df_use_er)== 0)
#     next
# # Calculate the Maximum Cumulative ER (ER_C_cumsum) for the Year:
#   cumsum_er <- max(df_use_er$ER_C_cumsum, na.rm = TRUE)
#   df_out[i,'ER_C_cumsum'] <- cumsum_er
# 
# # Normalize the Cumulative ER to Create a CDF (Cumulative Distribution Function):
#   df_math_er <- df_use_er %>%
#     dplyr::mutate(ER_C_cumsum_cdf = ER_C_cumsum/cumsum_er)
#   
# # Find the Julian Day (jday) When the ER Reaches the Specified Quantile:
#   jday <- which.min(abs(df_math$ER_C_cumsum_cdf- quantile))
#   df_out[i,'jday'] <- jday
#   
#   # Extract Corresponding Data for That jday:
#   row <- data.frame(df_math_er %>%
#                       dplyr::filter(jday== !!jday))
#   df_out_er[i,'clean_name'] <- row[,'clean_name']
#   df_out_er[i,'date'] <- row[,'date']
#   df_out_er[i,'ER_C_cumsum'] <- row[,'ER_C_cumsum']
#   df_out_er[i,'cdf'] <- row[,'ER_C_cumsum_cdf']
#   } # end for loop
# 
# # Filter and Fill Missing Data (df_quants):
# df_quants_er <- df_out_er %>%
#   dplyr::filter(!is.na(jday)) %>%
#   dplyr::group_by(site) %>%
#   tidyr::fill(clean_name, .direction = 'down')
# 
# df_quants_er %>%
#   dplyr::group_by(quantile) %>%
#   dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
#                    jday_sd = sd(jday, na.rm = TRUE),
#                    jday_min = min(jday, na.rm = TRUE),
#                    jday_max = max(jday, na.rm = TRUE))
```

```{r, warning=F, size = 'small', echo=T, include=F}
# df_quants_er %>%
#   dplyr::group_by(quantile, clean_name) %>%
#   dplyr::summarise(jday_median = median(jday, na.rm = TRUE),
#                    jday_sd = sd(jday, na.rm = TRUE),
#                    jday_min = min(jday, na.rm = TRUE),
#                    jday_max = max(jday, na.rm = TRUE))

```

```{r, warning=F, size = 'small', echo=T, include=F}
# # trend analysis for quantiles
# quantile_trends_er <- df_quants_er %>%
#   dplyr::arrange(site, year) %>%
#   dplyr::group_by(site, quantile) %>%
#   tidyr::nest() %>%
#   # Performing Sen's Slope and Mann-Kendall Trend Tests
#   dplyr::mutate(sens = purrr::map(data, ~trend::sens.slope(x = .$jday)),
#                 mk_test = purrr::map(data, ~trend::mk.test(x = .$jday))) %>%
#   # Extracting Trend Statistics
#   dplyr::mutate(sens_s = purrr::map_dbl(sens,~.$estimates), 
#                 sens_p = purrr::map(sens,~.$p.value),
#                 mk_p = purrr::map_dbl(mk_test,~.$p.value),
#                 # mk_t = purrr::map_dbl(mk_test, ~.$estimates['tau']),
#                 sens_sig = ifelse(sens_p <= 0.05, 'significant', 'non-significant'),
#                 sens_slope = ifelse(sens_s > 0, 'increasing',  'decreasing'),
#                 mk_sig = ifelse(mk_p <= 0.05, 'significant', 'non-significant')
#                 )
# 
# 
# quantile_sum_sigs_er <- quantile_trends_er %>%
#   dplyr::ungroup() %>%
#   dplyr::select(quantile, site, mk_p, #mk_t,
#                 mk_sig, #mk_slope,
#                 sens_s, sens_sig, sens_slope)
# 
# quants_n_sig_er <- quantile_sum_sigs_er %>%
#   dplyr::filter(mk_p <= 0.05) %>%
#   dplyr::group_by(quantile, sens_slope) %>%
#   summarise(n = n()) %>%
#   data.frame()
# 
# quantile_trends_sum_er <- df_quants_er %>%
#   dplyr::left_join(quantile_sum_sigs_er) %>%
#   dplyr::mutate(sens_meaning = dplyr::case_when(sens_slope== 'increasing' ~ 'Later',
#                                                 sens_slope== 'decreasing' ~ 'Earlier'),
# # mk_meaning = case_when(mk_slope =='increasing' ~ 'Later',
# # mk_slope =='decreasing' ~ 'Earlier'),
# )

```

## NEW STUFF:

```{r, warning=F, size = 'small', echo=T, include=T}

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual_mon <- river_metab_filled %>%
  mutate(month=month(date))%>%
  group_by(site, year, month) %>%
  summarise(
    monthly_avg_GPP = mean(GPP_filled, na.rm = TRUE),
    monthly_cv_GPP = sd(GPP_filled, na.rm = TRUE) / mean(GPP_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()

# Summarize intra-annual CV for each year
data_intra_annual_yr <- data_intra_annual_mon %>%
  group_by(site,year) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_GPP, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )

# Summarize intra-annual CV for each year
data_intra_annual_overall <- data_intra_annual_mon %>%
  group_by(site) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_GPP, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )
```

```{r, warning=F, size = 'small', echo=F, include=F}

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual_mon_er <- river_metab_filled_ER %>%
  mutate(month=month(date))%>%
  group_by(site, year, month) %>%
  summarise(
    monthly_avg_ER = mean(ER_filled, na.rm = TRUE),
    monthly_cv_ER = sd(ER_filled, na.rm = TRUE) / mean(ER_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()

# Summarize intra-annual CV for each year
data_intra_annual_yr_er <- data_intra_annual_mon_er %>%
  group_by(site,year) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_ER, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_ER, na.rm = TRUE) / mean(monthly_avg_ER, na.rm = TRUE) * 100
  )

# Summarize intra-annual CV for each year
data_intra_annual_overall_er <- data_intra_annual_mon_er %>%
  group_by(site) %>%
  summarise(
    mu_intra_annual = mean(monthly_avg_ER, na.rm = TRUE),
    cv_intra_annual = sd(monthly_avg_ER, na.rm = TRUE) / mean(monthly_avg_ER, na.rm = TRUE) * 100
  )
```

#### Mean GPP by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
# Create the plot
ggplot(data_intra_annual_yr, aes(x = year, y = mu_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  #  scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual GPP by stream site",
    x = "Year",
    y = "GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
#  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### Mean ER by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
# Create the plot
ggplot(data_intra_annual_yr_er, aes(x = year, y = mu_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  #  scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual ER by stream site",
    x = "Year",
    y = "ER"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
#  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### CV in GPP by site per year

CV as sd()/mean() \*100

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
ggplot(data_intra_annual_yr, aes(x = year, y = cv_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site) +
  theme_minimal() +  
  labs(
    title = "Degree of CV in GPP each year stream site",
    x = "Year",
    y = "CV of GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

#### CV in ER by site per year

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=14}
ggplot(data_intra_annual_yr_er%>%filter(cv_intra_annual>-1940), aes(x = year, y = cv_intra_annual, color = mu_intra_annual, group = site)) +
  geom_line() +  
  geom_point() +  
  facet_wrap(~site) +
  theme_minimal() +  
  labs(
    title = "Degree of CV in GPP each year stream site",
    x = "Year",
    y = "CV of GPP"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  theme(legend.position = "none")+
  scale_color_viridis_c(name="GPP")  
```

```{r, warning=F, size = 'small', echo=T}

library(ggplot2)
library(ggrepel)
library(viridis)

spc_tbl <- table_s1 %>%
  mutate(site = paste0("nwis_", `Site Number`))

# Ensure `site` is correctly formatted in both dataframes
spc_tbl <- spc_tbl %>%
  mutate(site = paste0("nwis_", str_pad(`Site Number`, width = 8, pad = "0")))


site_check_with_cv <- spc_tbl %>%
  left_join(data_intra_annual_overall, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")


site_check_with_cv_er <- spc_tbl %>%
  left_join(data_intra_annual_overall_er, by = c("site")) %>%
  dplyr::rename(drainage_area_km="Drainage Area (km2)")

```

#### Histograms of mean GPP by location:

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=14}
site_check_with_cv$drainage_category <- cut(site_check_with_cv$drainage_area_km,
                                            breaks = quantile(site_check_with_cv$drainage_area_km, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("Small", "Medium", "Large", "Very Large"),
                                            include.lowest = TRUE)


site_check_with_cv$lat_category <- cut(site_check_with_cv$Latitude,
                                            breaks = quantile(site_check_with_cv$Latitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("south", "mid-south", "mid-north", "north"),
                                            include.lowest = TRUE)

site_check_with_cv$lon_category <- cut(site_check_with_cv$Longitude,
                                            breaks = quantile(site_check_with_cv$Longitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("west", "west-central", "east-central", "east"),
                                            include.lowest = TRUE)


hist1<- ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "Mean GPP", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "Mean GPP", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "Mean GPP", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}
site_check_with_cv_er$drainage_category <- cut(site_check_with_cv_er$drainage_area_km,
                                            breaks = quantile(site_check_with_cv_er$drainage_area_km, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("Small", "Medium", "Large", "Very Large"),
                                            include.lowest = TRUE)


site_check_with_cv_er$lat_category <- cut(site_check_with_cv_er$Latitude,
                                            breaks = quantile(site_check_with_cv_er$Latitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("south", "mid-south", "mid-north", "north"),
                                            include.lowest = TRUE)

site_check_with_cv_er$lon_category <- cut(site_check_with_cv_er$Longitude,
                                            breaks = quantile(site_check_with_cv_er$Longitude, probs = seq(0, 1, 0.25), na.rm = TRUE),
                                            labels = c("west", "west-central", "east-central", "east"),
                                            include.lowest = TRUE)


hist1<- ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "Mean ER", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "Mean ER", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv_er, aes(x = mu_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "Mean ER", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

#### Histograms of CV GPP by location:

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}

hist1<- ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "CV GPP", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "CV GPP", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "CV GPP", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=16, fig.height=12}

hist1<- ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = drainage_category,  color = drainage_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "drainage area") +  # Use a continuous color scale for CV
    scale_color_viridis_d(option = "viridis", name = "drainage area") +
  labs(x = "CV ER", y = "Frequency", fill = "Drainage Size") +
  theme_minimal()


hist2<- ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lat_category, color = lat_category), position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Latitude") +  
  labs(x = "CV ER", y = "Frequency", fill = "Latitude ") +
       scale_color_viridis_d(option = "viridis", name = "Latitude") +
  theme_minimal()


hist3<-ggplot(site_check_with_cv_er, aes(x = cv_intra_annual)) +
  geom_histogram(aes(fill = lon_category, color = lon_category),  position = "dodge", alpha = 0.7, bins = 10) +
   scale_fill_viridis_d(option = "viridis", name = "Longitude") + 
     scale_color_viridis_d(option = "viridis", name = "Longitude") +
  labs(x = "CV ER", y = "Frequency", fill = "Longitude ") +
  theme_minimal()

```

```{r, warning=F, size = 'small', echo=F, fig.width=12, fig.height=3}
# Arrange plots with shared legend at the bottom
grid <- ggarrange(
  hist1,
  hist2,
  hist3,
  ncol = 3, nrow = 1
)

grid
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=16, fig.height=12}

# Create the map plot with color based on CV of GPP
coords_map <- ggmap(map) +
  geom_point(data = site_check_with_cv, aes(x = Longitude, y = Latitude, color = cv_intra_annual), size = 3) +
  geom_label_repel(data = site_check_with_cv, aes(x = Longitude, y = Latitude, label = site, color = cv_intra_annual), fontface = "bold", size = 5) +
  xlab("Longitude") + ylab("Latitude") + theme_bw() +
  scale_color_viridis_c(option = "viridis", name = "Intra-annual CV") +  # Use a continuous color scale for CV
  # Add the scale bar
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black", face = "bold"),
    panel.border = element_rect(size = 1.5, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank()
  )

# Print the map
coords_map

```

#### **Q1: Is the degree of intra-annual variation in lake or stream metabolism is a function of regional location in the US/ northern hemisphere?**

#### **A:** Possibly not in the streams.

-   Might be worth scaling GPP by catchment size/ or stream size first.

```{r, warning=F, size = 'small', echo=F, include=F}

library(dplyr)

# Assuming 'data' has 'Year', 'Month', and 'GPP_filled'
data_intra_annual <- river_metab_filled %>%
  mutate(month= month(date)) %>%
  group_by(year, month) %>%
  summarise(
    monthly_avg_GPP = mean(GPP_filled, na.rm = TRUE),
    monthly_cv_GPP = sd(GPP_filled, na.rm = TRUE) / mean(GPP_filled, na.rm = TRUE) * 100
  ) %>%
  ungroup()



# Summarize intra-annual CV for each year
intra_annual_variation <- data_intra_annual %>%
  group_by(year) %>%
  summarise(
    cv_intra_annual = sd(monthly_avg_GPP, na.rm = TRUE) / mean(monthly_avg_GPP, na.rm = TRUE) * 100
  )
```


#### Load Lake productivity data:

##### Ladwig et al. NTL 2008-2021

- NEP in milligramPerMeterCubedPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Ladwig_NTL_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/model_output.csv")%>%
  mutate(date=as.Date(datetime),
         year=year(date),
         jday= yday(date)) %>%
  filter(date> as.Date("2007-12-31") & date< as.Date("2022-01-01")) %>%
  mutate(NEP_gm3d = c(NEP*0.001)) %>%
    mutate(source= "Ladwig") %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source) 

head(Ladwig_NTL_dat_metab)
```


##### Rabaey et al. 

- GPP/ER/NEP in grams O2 per m3 per day

```{r, warning=F, size = 'small', echo=T, include=T}
Rabaey_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metabolism_daily_summaries.csv")%>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>% 
  mutate(source= "Rabaey") %>%
  dplyr::select(site="name", date, year, jday, NEP_gm3d ="nep", source) 

head(Rabaey_dat_metab)
```




##### Oleksy et al. 

- GPP/ER in milligramsO2PerLiterPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Oleksy_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/daily_metabolism.csv") %>%
  filter(lakeName %in% Oleksy_good_sites)%>% ## Oleksy_good_sites
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Oleksy",
           NEP = c(GPP_mgO2L-ER_mgO2L),
           NEP_g =c(NEP * 0.001),
           NEP_gm3d =c(NEP * 0.001)) %>%
  dplyr::select(site="lakeName", date, year, jday, NEP_gm3d, source)

head(Oleksy_dat_metab)
```



##### Lotting et al. 

- Production (millimolscarbonpercubicmeterperday) estimated from high frequency dissolved oxygen time series.

-  1 mmol of carbon is associated with the production or consumption of 1 mmol of O2 so: 

    millimolscarbonpercubicmeterperday * 32.00 = milligramPerMeterCubedPerDay

```{r, warning=F, size = 'small', echo=T, include=T}
Lotting_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/lottig_etal_data.csv") %>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Lotting",
           NEP_mg =c(o2_pp_mmolcm3d * 32),
           NEP_gm3d =c(NEP_mg * 0.001)) %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source)

head(Lotting_dat_metab)
```


```{r, warning=F, size = 'small', echo=F, include=F}
lake_metab <- rbind(Ladwig_NTL_dat_metab, Rabaey_dat_metab, Oleksy_dat_metab, Lotting_dat_metab)
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=7}
ggplot(lake_metab%>%filter(source=="Ladwig"), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Ladwig data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
ggplot(lake_metab%>%filter(source=="Rabaey") %>%filter(year>2007), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Rabaey data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```
We want lakes from 025-D Hoxit2 (poor resolution), 205, susquereba , mahaney 
We have P1 and P8 in Bella's


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=6}
ggplot(lake_metab%>%filter(source=="Lotting"), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Lotting data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```

Looks like we really only want Castle and Acton from this data set.



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=12, fig.height=9}
ggplot(lake_metab%>%filter(source=="Oleksy") %>%filter(year>2007), aes(x = jday, y = NEP_gm3d, color = factor(year), group = interaction(site, year), shape=source)) +
  geom_point() +  
  facet_wrap(~site, scales = 'free_y') +  # scales = 'free_y'
  theme_minimal() +  
  labs(
    title = "Average annual NEP by Lake in Oleksy data",
    x = "Day of Year",
    y = "NEP gO m3 d"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    panel.grid.major = element_line(size = 0.5, color = "gray", linetype = "dashed"),
    panel.grid.minor = element_blank()  # Remove minor grid lines
  ) +
  scale_color_viridis_d(name="Year")  # Use discrete color scale for 'year'
```

