---
title: "Synchrony in aquatic ecosystem productivity"
author: "K. Loria"
date: "`r Sys.Date()`"
format: 
  revealjs:   # << this is the important part!
    theme: default
    slide-number: true
    toc: false
    transition: fade
editor: visual
---


## Moran effect

**Moran effect:** Spatial Synchrony is expected to be the strongest between nearby ecosystem with spatially auto correlated abiotic conditions.

Aquatic ecosystems integrate localized and larger scale phenomena (climatic) and the degree of variation in local catchment characteristics (forest cover, land use, topography) is likely to effect how sensitive a given lake or stream is to the Moran effect.

## Synchrony as means of testing the Moran effect

Synchrony can be used to evaluate how sensitive streams and lake are broad- and fine-scale ecosystem drivers.

We think this has been done in lakes (Kling et al. 2000, Vogt et al. 2011, Walter et al. 2021, broadly the landscape limnology concepts by Soranno) but not in as much in streams (Dunkle et al. 2025 -fish, Valentine et al. 2024 -also fish).

We expect GPP and ER streams and lakes to vary based on size (strahler stream order), dominant land cover, topography, geology, and water temperature.

## Research Questions

1\. How does the scale and strength of spatial Synchrony of aquatic metabolism vary among streams and how does that compare to the spatial Synchrony of climate variables?

2\. For streams and lakes within comparable ecoregion, how synchronous is aquatic metabolism and what explains why some areas are more synchronous across lake-stream pairs?

## Stream Data:

**Marzolf et al. 2024** [https://aslopubs.onlinelibrary.wiley.com/doi/10.1002/lol2.10407)- Phenology of gross primary productivity in rivers displays high variability within years but stability across years](https://aslopubs.onlinelibrary.wiley.com/doi/10.1002/lol2.10407)

-   Roughly 63 streams
-   Metabolism from 2008 to 2020

## Stream Data QA'Qc:

-   Model convergence (rhat\>1.05)
-   logical flow - k600
    -   Normality of the distribution of k600 across different flows
-   logical GPP (GPP \> 0) and ER (ER \< 0) estimates

## Lake Data:

Paired case study approach to stream locations.

<p><small>**Ladwig et al.** [Modeling dataset:](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-ntl.402.2) Long-term Change in Metabolism Phenology across North-Temperate Lakes, Wisconsin, USA 1979-2019 - 8 lakes 2008 - 2021 - Only NEP in milligramPerMeterCubedPerDay</small></p>

<p><small>**Rabaey et al.** [Freshwater biogeochemical hotspots](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1531.2): High primary production and ecosystem respiration in shallow waterbodies (38 lakes 1 year - GPP/ER/NEP in grams O2 per m3 per day)</small></p>

<p><small>**Oleksy et al.** [Metabolism dataset](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1541.2): one year of high-frequency temperature, dissolved oxygen, wind, photo synthetically active radiation observations and low-frequency nutrient data for 58 lakes in the Global Lake Ecological Observatory Network. (58 lakes 1 year- GPP/ER in milligramsO2PerLiterPerDay) </small></p>

<p><small>**Lotting et al.** Added castle and and midwest lake </small></p>

```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables

```

```{r, echo=FALSE, include=FALSE}
RR_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/RR_dat.csv")%>%
  mutate(date= as.Date(date)) %>%
  dplyr::select(Site_ID, Lat, Lon, date, year, GPP_filled, discharge, temp.water, 
                ppt_mm, TmeanC) 
```

## Geographic coverage of stream and lake metabolism data:

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
## Read in data 

metab_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_TS.rds")
# str(metab_ts)

dat_ts <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/TS_SPC_pH_FNU.rds")
# str(dat_ts)
# summary(dat_ts)
# unique(dat_ts$site_name)

metab_meta_dat <- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata.rds")
# str(metab_ts)


marz_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/1_site_info.csv")
# 63 different sites 

Ladwig_NTL_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/Ladwig_NTL_lake_cords.csv")%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="ladwig_NTL_lake")

Oleksy_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/lake_descriptions.csv") %>%
  filter(data_request=="yes")%>%
  dplyr::select(Site.Number ="lakeName", Latitude="lat_dd", Longitude="long_dd") %>%
  mutate(source="oleksy_lake")

Oleksy_good_sites <- unique(Oleksy_dat_metadf$Site.Number)

Rabaey_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metadata.csv")%>%
  dplyr::select(Site.Number ="pond_name", Latitude="latitude", Longitude="longitude") %>%
  mutate(source="rabaey_lake")

corman_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/load_metab_results/corman_lake_cords.csv")%>%
  filter(country=="United States" | country== "Canada") %>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="corman_lake")

lotting_dat_metadf <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/Lotting_lake_cords.csv")%>%
  dplyr::select(Site.Number ="Lake", Latitude="Latitude", Longitude="Longitude") %>%
  mutate(source="lotting_lake")


```

```{r, warning=F, size = 'small', echo=F, include=F}

dat_ts_clean <- dat_ts%>%
  filter(GPP.Rhat < 1.05 &  GPP >0 & ER <0)

# unique(dat_ts_clean$long_name)

marz_dat_names <- marz_dat%>%
  dplyr::select(Site.Number, Latitude, Longitude) %>%
  mutate(source="marzolf_stream")

dat_ts_names <-metab_meta_dat %>%
   dplyr::select(Site.Number = "site_no", Latitude = "dec_lat_va", Longitude= "dec_long_va") %>%
  mutate(source="appling_stream")

## quick site check
site_check <- rbind(marz_dat_names, 
                  #dat_ts_names,
                  Ladwig_NTL_dat_metadf, 
                   Oleksy_dat_metadf,
                   Rabaey_dat_metadf, #, 
                   # corman_dat_metadf, 
                    lotting_dat_metadf)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
## map packages 

library(tidyverse)
library(ggrepel)
library(ggmap)
library(ggsci)
library(scales)

library(dplyr)
library(ggplot2)
library(ggrepel)


#devtools::install_github("briatte/tidykml")
library(tidykml)
library(sf)
```

```{r, warning=F, size = 'small', echo=F, include=T, include=F}

# Define the bounding box coordinates add some wiggle room:

range(na.omit(site_check$Longitude))
range(site_check$Latitude)
# 
# Define the bounding box coordinates
bbox <- c(left = -128.300, bottom = 22.50, right = -63.800, top = 51.50)

# Retrieve terrain map layer using get_stamenmap()
map <- get_stadiamap(bbox = bbox, zoom = 5, maptype = 'stamen_terrain')
# need api run for the first time

#library(ggmap)
# Plot the terrain map
stamen_map <- ggmap(map)
```

```{r, size='large', warning=F, echo=F, include=T}
coords_map <- ggmap(map) +
  geom_point(data = site_check, aes(x = Longitude, y = Latitude, color = source, shape =source), size = 2) +
  geom_label_repel(data = site_check, aes(x = Longitude, y = Latitude, label = Site.Number,  fill = source), color= "black", fontface = "bold", size = 2, max.overlaps = 20) +
  xlab("Longitude") + ylab("Latitude") + theme_bw() +
  scale_color_viridis_d(option = "viridis",  name = "source")  +
 scale_fill_viridis_d(option = "viridis",  name = "source", alpha = 0.5)  +

  # Add the scale bar
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black", face = "bold"),
    panel.border = element_rect(size = 1.5, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank()
  )
coords_map
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
Ladwig_NTL_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/knb-lter-ntl.402.2/model_output.csv")%>%
  mutate(date=as.Date(datetime),
         year=year(date),
         jday= yday(date)) %>%
  filter(date> as.Date("2007-12-31") & date< as.Date("2022-01-01")) %>%
  mutate(NEP_gm3d = c(NEP*0.001)) %>%
    mutate(source= "Ladwig") %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source) 

head(Ladwig_NTL_dat_metab)

Rabaey_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1531.2/metabolism_daily_summaries.csv")%>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>% 
  mutate(source= "Rabaey") %>%
  dplyr::select(site="name", date, year, jday, NEP_gm3d ="nep", source) 

head(Rabaey_dat_metab)


Oleksy_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/edi.1541.2/daily_metabolism.csv") %>%
  filter(lakeName %in% Oleksy_good_sites)%>% ## Oleksy_good_sites
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Oleksy",
           NEP = c(GPP_mgO2L-ER_mgO2L),
           NEP_g =c(NEP * 0.001),
           NEP_gm3d =c(NEP * 0.001)) %>%
  dplyr::select(site="lakeName", date, year, jday, NEP_gm3d, source)

head(Oleksy_dat_metab)


Lotting_dat_metab <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Lakes_dat/Facundo_Rec/knb-lter-ntl.397.1/lottig_etal_data.csv") %>%
  mutate(date=as.Date(date),
         year=year(date),
         jday= yday(date)) %>%
    mutate(source= "Lotting",
           NEP_mg =c(o2_pp_mmolcm3d * 32),
           NEP_gm3d =c(NEP_mg * 0.001)) %>%
  dplyr::select(site="lake", date, year, jday, NEP_gm3d, source)

head(Lotting_dat_metab)

lake_metab <- rbind(Ladwig_NTL_dat_metab, Rabaey_dat_metab, Oleksy_dat_metab, Lotting_dat_metab)

```

## Temporal resolution of stream metabolism data:

```{r, echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

plot_Sample_Temporal_Coverage <- RR_dat %>% 
  group_by(Site_ID, date) %>% 
  summarise(`log(N)` = log(n())) %>% 
  ggplot() + ylab("Site ID") +
  geom_point(aes(x = date, y = Site_ID, color = `log(N)`)) +
  theme_bw() +
  theme(text = element_text(size=10), legend.position = "none")

plot_Sample_Temporal_Coverage
```

## Temporal resolution of lake metabolism data:

```{r, echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
plot_Sample_Temporal_Coverage <- lake_metab %>% 
  filter(year>2007 & year<=2020)%>%
  group_by(site, date, source) %>% 
  summarise(`log(N)` = log(n())) %>% 
  ggplot() + ylab("Site ID") +
  geom_point(aes(x = date, y = site, color = source)) +
  theme_bw() +
  theme(text = element_text(size=10))

plot_Sample_Temporal_Coverage
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

# Load Packages
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables



RR_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/RR_dat.csv")%>%
  mutate(date= as.Date(date)) %>%
  dplyr::select(Site_ID, Lat, Lon, date, year, GPP_filled, discharge, temp.water, 
                ppt_mm, TmeanC) 


```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

RR_dat_sum <-  RR_dat%>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(GPP_an= mean(GPP_filled, na.rm=T),
            GPP_max= max(GPP_filled, na.rm=T),
        Q_anl= mean(discharge, na.rm=T),
        wtemp_min= min(temp.water, na.rm=T),
        wtemp_max= max(temp.water, na.rm=T),
        wtemp_an= mean(temp.water, na.rm=T),
        ppt_ann= mean(ppt_mm, na.rm=T),
        Atemp_ann= max(TmeanC, na.rm=T))
  
RR_dat_TS_gpp <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, GPP_an) %>%
  pivot_wider(names_from = c(year), values_from = GPP_an) 


RR_dat_TS_gpp <-na.omit(RR_dat_TS_gpp)

 
# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr <- Sncf(x=RR_dat_TS_gpp$Lon,
                   y=RR_dat_TS_gpp$Lat,
                   z=RR_dat_TS_gpp[,4:ncol(RR_dat_TS_gpp)],
                               resamp = 1000, 
                               latlon = T)

plot(RR_df.corr)


RR_dat_TS_gpp1 <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, GPP_max) %>%
  pivot_wider(names_from = c(year), values_from = GPP_max) 


RR_dat_TS_gpp1 <-na.omit(RR_dat_TS_gpp1)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr1 <- Sncf(x=RR_dat_TS_gpp1$Lon,
                   y=RR_dat_TS_gpp1$Lat,
                   z=RR_dat_TS_gpp1[,4:ncol(RR_dat_TS_gpp1)],
                   resamp = 1000, 
                   latlon = T)


plot(RR_df.corr1)


GPP_corrgram_summary.table <- summary(RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df <- data.frame(x = matrix(unlist(RR_df.corr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr$real$predicted$y)))


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df <- data.frame(x = matrix(unlist(RR_df.corr$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`)

```

## Correlogam of annual GPP Synchrony

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

GPP.synch.plot <- ggplot() +
  geom_ribbon(data = GPP.corr.boot.df,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
              fill = "#4f8c4f", alpha=0.1) +
  geom_line(data = GPP.corr.pred.df,
            aes(x = x, y = y), color="#4f8c4f") +
  geom_hline(yintercept = 0) +
 # geom_hline(yintercept = RR_df.corr$real$cbar, linetype = "dashed") +
  theme_classic() +
  labs(title = "Annual mean stream GPP", 
       y= "Correlation", 
       x="Distance (km)") +
  theme(text = element_text(size=12))
GPP.synch.plot
```

## Annual Synchrony of hydro/climate/environmental parameters

Mean spring flow example, where "spring" is month Feb-May

```{r, size='small', echo=T, include=T, warning=FALSE}
### look at an element of environmental variation -- lets look at spring flow
RR_dat_sum <-  RR_dat%>%
  mutate(month=month(date)) %>%
  filter(month>1 & month<=6) %>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(Q_max= max(discharge, na.rm=T),
            Q_anl= mean(discharge, na.rm=T))

# choosing max flow:
RR_dat_TS_Q <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, Q_max) %>%
  pivot_wider(names_from = c(year), values_from = Q_max) 


RR_dat_TS_Q <-na.omit(RR_dat_TS_Q)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
Q_RR_df.corr <- Sncf(x=RR_dat_TS_Q$Lon,
                   y=RR_dat_TS_Q$Lat,
                   z=RR_dat_TS_Q[,4:ncol(RR_dat_TS_Q)],
                   resamp = 1000, 
                   latlon = T)

```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
Q_corrgram_summary.table <- summary(Q_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Q.corr.pred.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$real$predicted$x)),
                               y = matrix(unlist(Q_RR_df.corr$real$predicted$y)))


# ...and a dataframe of the y values for the bootstrap prediction
Q.corr.bootValues.df <- as.data.frame(t(Q_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Q.corr.boot.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$boot$boot.summary$predicted$x)),
                               yminQ = Q.corr.bootValues.df$`0.025`,
                               ymaxQ = Q.corr.bootValues.df$`0.975`)



RR_dat_sum <-  RR_dat%>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(GPP_an= mean(GPP_filled, na.rm=T),
            GPP_max= max(GPP_filled, na.rm=T),
        Q_anl= mean(discharge, na.rm=T),
        wtemp_min= min(temp.water, na.rm=T),
        wtemp_max= max(temp.water, na.rm=T),
        wtemp_an= mean(temp.water, na.rm=T),
        ppt_ann= mean(ppt_mm, na.rm=T),
        Atemp_ann= max(TmeanC, na.rm=T))


### Alright lets look at Flow
RR_dat_TS_temp <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, wtemp_an) %>%
  pivot_wider(names_from = c(year), values_from = wtemp_an) 


RR_dat_TS_temp <-na.omit(RR_dat_TS_temp)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
temp_RR_df.corr <- Sncf(x=RR_dat_TS_temp$Lon,
                     y=RR_dat_TS_temp$Lat,
                     z=RR_dat_TS_temp[,4:ncol(RR_dat_TS_temp)],
                     resamp = 1000, 
                     latlon = T)

plot(temp_RR_df.corr)

temp_corrgram_summary.table <- summary(temp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
temp.corr.pred.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$real$predicted$x)),
                             y = matrix(unlist(temp_RR_df.corr$real$predicted$y)))


# ...and a dataframe of the y values for the bootstrap prediction
temp.corr.bootValues.df <- as.data.frame(t(temp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
temp.corr.boot.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$boot$boot.summary$predicted$x)),
                             yminWtemp = temp.corr.bootValues.df$`0.025`,
                             ymaxWtemp = temp.corr.bootValues.df$`0.975`)



RR_dat_sum <-  RR_dat%>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(GPP_an= mean(GPP_filled, na.rm=T),
            GPP_max= max(GPP_filled, na.rm=T),
        Q_anl= mean(discharge, na.rm=T),
        wtemp_min= min(temp.water, na.rm=T),
        wtemp_max= max(temp.water, na.rm=T),
        wtemp_an= mean(temp.water, na.rm=T),
        ppt_ann= mean(ppt_mm, na.rm=T),
        Atemp_ann= max(TmeanC, na.rm=T))


RR_dat_TS_ppt <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, ppt_ann) %>%
  pivot_wider(names_from = c(year), values_from = ppt_ann) 


RR_dat_TS_ppt <-na.omit(RR_dat_TS_ppt)



# Now, calculate spatial Synchrony in the time series using the Sncf() function
ppt_RR_df.corr <- Sncf(x=RR_dat_TS_ppt$Lon,
                        y=RR_dat_TS_ppt$Lat,
                        z=RR_dat_TS_ppt[,4:ncol(RR_dat_TS_ppt)],
                        resamp = 1000, 
                        latlon = T)

plot(ppt_RR_df.corr)

ppt_corrgram_summary.table <- summary(ppt_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
ppt.corr.pred.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$real$predicted$x)),
                                y = matrix(unlist(ppt_RR_df.corr$real$predicted$y)))


# ...and a dataframe of the y values for the bootstrap prediction
ppt.corr.bootValues.df <- as.data.frame(t(ppt_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ppt.corr.boot.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$boot$boot.summary$predicted$x)),
                                yminPPT = ppt.corr.bootValues.df$`0.025`,
                                ymaxPPT = ppt.corr.bootValues.df$`0.975`)

### Alright lets look max summer temp
RR_dat_sum <-  RR_dat%>%
  mutate(month =month(date)) %>%
  filter(month>5 & month<10) %>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(wtemp_max= max(temp.water, na.rm=T),
            wtemp_an= mean(temp.water, na.rm=T))



RR_dat_sum <-  RR_dat%>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(GPP_an= mean(GPP_filled, na.rm=T),
            GPP_max= max(GPP_filled, na.rm=T),
        Q_anl= mean(discharge, na.rm=T),
        wtemp_min= min(temp.water, na.rm=T),
        wtemp_max= max(temp.water, na.rm=T),
        wtemp_an= mean(temp.water, na.rm=T),
        ppt_ann= mean(ppt_mm, na.rm=T),
        Atemp_ann= max(TmeanC, na.rm=T))


RR_dat_TS_Atemp <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, Atemp_ann) %>%
  pivot_wider(names_from = c(year), values_from = Atemp_ann) 


RR_dat_TS_Atemp <-na.omit(RR_dat_TS_Atemp)





# Now, calculate spatial Synchrony in the time series using the Sncf() function
Atemp_RR_df.corr <- Sncf(x=RR_dat_TS_Atemp$Lon,
                       y=RR_dat_TS_Atemp$Lat,
                       z=RR_dat_TS_Atemp[,4:ncol(RR_dat_TS_Atemp)],
                       resamp = 1000, 
                       latlon = T)

plot(Atemp_RR_df.corr)

Atemp_corrgram_summary.table <- summary(Atemp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Atemp.corr.pred.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$real$predicted$x)),
                               y = matrix(unlist(Atemp_RR_df.corr$real$predicted$y)))


# ...and a dataframe of the y values for the bootstrap prediction
Atemp.corr.bootValues.df <- as.data.frame(t(Atemp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Atemp.corr.boot.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$boot$boot.summary$predicted$x)),
                               yminAtemp = Atemp.corr.bootValues.df$`0.025`,
                               ymaxAtemp = Atemp.corr.bootValues.df$`0.975`)



############
big_boot_df<- GPP.corr.boot.df %>%
  left_join(Q.corr.boot.df, by=c("x")) %>%
  left_join(temp.corr.boot.df, by=c("x")) %>%
  left_join(ppt.corr.boot.df, by=c("x")) %>%
  left_join(Atemp.corr.boot.df, by=c("x")) 
  
```

## Correlogam of Synchrony of GPP + other hydro/climate/environmental parameters near streams

```{r, size = 'large', echo=FALSE, include=T, warning=FALSE}

pred_df <- bind_rows(
  Q.corr.pred.df %>% mutate(group = "Flow"),
  GPP.corr.pred.df %>% mutate(group = "GPP"),
  temp.corr.pred.df %>% mutate(group = "Water Temp"),
  ppt.corr.pred.df %>% mutate(group = "PPT"),
  Atemp.corr.pred.df %>% mutate(group = "Air Temp")
)

ribbon_df <- big_boot_df %>%
  transmute(
    x = x,
    Flow_ymin = yminQ, Flow_ymax = ymaxQ,
    GPP_ymin = yminGPP, GPP_ymax = ymaxGPP,
    WaterTemp_ymin = yminWtemp, WaterTemp_ymax = ymaxWtemp,
    PPT_ymin = yminPPT, PPT_ymax = ymaxPPT,
    AirTemp_ymin = yminAtemp, AirTemp_ymax = ymaxAtemp
  ) %>%
  pivot_longer(cols = -x, names_to = c("group", ".value"), names_sep = "_")

synch.plot <- ggplot() +
  geom_line(data = pred_df,
            aes(x = x, y = y, color = group)) +
  geom_hline(yintercept = 0) +
  theme_classic() +
  labs(title = "Synchrony across GPP, flow, temp, and ppt",
              y= "Correlation", 
       x="Distance (km)",
       color = "Variable", fill = "Variable") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size=14),
        legend.position = "bottom") +
  scale_color_manual(values = c(
    "Flow" = "#6293c4",
    "GPP" = "#5a8f5e",
    "Water Temp" = "#DAA520",
    "PPT" = "#654cad",
    "Air Temp" = "#b82c02"
  )) +
   geom_ribbon(data = big_boot_df,
              aes(x = x, ymin = yminQ, ymax = ymaxQ),
              fill = "#6293c4", alpha=0.1) +
  geom_ribbon(data = big_boot_df,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
              fill = "#5a8f5e", alpha=0.1) +
  geom_ribbon(data = big_boot_df,
              aes(x = x, ymin = yminWtemp, ymax = ymaxWtemp),
              fill = "#DAA520", alpha=0.1) +
  geom_ribbon(data = big_boot_df,
              aes(x = x, ymin = yminPPT, ymax = ymaxPPT),
              fill = "#654cad", alpha=0.1) +
  geom_ribbon(data = big_boot_df,
              aes(x = x, ymin = yminAtemp, ymax = ymaxAtemp),
              fill = "#b82c02", alpha=0.1)


synch.plot


```

## Next steps for streams analyses:

-   Data org:

    -   Identifying ecologically relevant aggregations of time series data to evaluate Synchrony? (e.g. annual spring max flow, average summer temperature)

    -   Land cover data organization - [NHD tools](https://www.sciencebase.gov/catalog/item/5669a79ee4b08895842a1d47)

    -   Light availability (given shading) - MODIS

## Next steps for streams analyses :

-   Compare based how Synchrony (possibly at finer than annual time sieries) in GPP and ER varies based on

    -   Stream size (small, med, large)

    -   Dominate land cover in the catchment/ or along the corridor

    -   Catchment slope

    -   Water temperature

## Next steps cont.

Lakes:

-   Data aggregation -\> QA'QC (GPP, ER, residence time?)

-   Appropriate temporal comparisons for the case study?

-   Lake synchony based on location or size to

    -   Lake size (small, med, large)

    -   Dominate land cover in the catchment/ or along the corridor

    -   Water temperature

## That's it so far!

Advice? / comments ?

Thanks!!


