---
title: "Data cleaning and infill for synchrony in aquatic ecosystem productivity"
description: "Final stream GPP and ER with gap ID"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```


```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables

```

```{r, echo=FALSE, include=FALSE}

source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/fillMiss3.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/cor_plot_fxn.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/depth_stability_fxn.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/ER_K_covar_fxn.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/fill_and_normalize_ER_metab.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/fill_and_normalize_metab.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/K_Q_covar_fxn.R')
source('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/RiverRecoveryDynamics/data_structure_fxn/K_stability_fxn.R')

```

```{r, echo=FALSE, include=FALSE}
RR_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/RR_dat.csv")%>%
  mutate(date= as.Date(date)) %>%
  dplyr::select(Site_ID, Lat, Lon, date, year, GPP, ER, GPP_filled, discharge, temp.water, 
                ppt_mm, TmeanC) 

unique(RR_dat$Site_ID)

```

#### Load Marzolf GPP data:
Starting concerns: 

- High rhats on all parameters..., supposed to be less than 1.2 
- Flow does go to zero?
- Depth goes to 39 ?!?
- Flow goes to 3213.608...


```{r, warning=F, size = 'small', echo=T, include=T}
river_metab <- readr::read_csv('/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/5_usgs_metabolism.csv')

summary(river_metab)
```

### Remove poorly converged days
Still some questionable values of K, ER, and GPP

```{r, warning=F, size = 'small', echo=T, include=T}

river_metab <-  river_metab%>%filter(GPP.Rhat<=1.05 & ER.Rhat<=1.05 & K600.Rhat <=1.05)
summary(river_metab)

```

```{r, warning=F, size = 'small', echo=T, include=T}

qualified_site_years <- river_metab %>%
  mutate(year = year(date)) %>%
  group_by(site, year) %>%
  summarise(n_days = n_distinct(date), .groups = "drop") %>%
  filter(n_days >= 288)

unique(qualified_site_years$site)

filtered_data <- river_metab %>%
  mutate(year = year(date)) %>%
  inner_join(qualified_site_years, by = c("site", "year"))

```

## Data quality 

1. What is the ER~ K?

2. What is the k variance 

3. What are GPP and ER ranges 


```{r, warning=F, size = 'small', echo=F, include=T}

ER_K_df <- ER_K_covar_metab(filtered_data)
summary(ER_K_df)

```



```{r, warning=F, size = 'small', echo=F, include=T}

plot <- ggplot() +
  geom_point(data = ER_K_df,
            aes(y = site, x = cov_ER_K, color=year)) +
  geom_vline(xintercept = 0) +
 # geom_hline(yintercept = RR_df.corr$real$cbar, linetype = "dashed") +
  theme_classic() +
  labs(title = "Annual GPP to ER correlation") +
  theme(text = element_text(size=12))

plot

```


```{r, warning=F, size = 'small', echo=F, include=T}

K600_stab <- K600_stability(filtered_data)
summary(K600_stab)
```

```{r, warning=F, size = 'small', echo=F, include=T}

plot <- ggplot() +
  geom_point(data = K600_stab,
            aes(y = site, x = slope_K600, color=year)) +
  geom_vline(xintercept = 0) +
 # geom_hline(yintercept = RR_df.corr$real$cbar, linetype = "dashed") +
  theme_classic() +
  labs(title = "Annual variation k600") +
  theme(text = element_text(size=12))


plot

```



```{r, warning=F, size = 'small', echo=F, include=T}

depth_stab <- depth_stability(filtered_data)
summary(depth_stab)
```

```{r, warning=F, size = 'small', echo=F, include=T}

plot <- ggplot() +
  geom_point(data = depth_stab,
            aes(y = site, x = slope_z, color=year)) +
  geom_vline(xintercept = 0) +
  theme_classic() +
  labs(title = "Annual variation depth") +
  theme(text = element_text(size=12))

plot

```





```{r, warning=F, size = 'small', echo=F, include=T}

KQ_cor <- K_Q_covar_metab(filtered_data)
summary(KQ_cor)

```

```{r, warning=F, size = 'small', echo=F, include=T}

plot <- ggplot() +
  geom_point(data = KQ_cor,
            aes(y = site, x = cov_K_Q, color=year)) +
  geom_vline(xintercept = 0) +
  theme_classic() +
  labs(title = "Annual K600 and flow") +
  theme(text = element_text(size=12))


plot

```


```{r, warning=F, size = 'small', echo=F, include=T}

filtered_data_flagged <- filtered_data %>%
  group_by(site, year) %>%
  mutate(K600_mean = mean(K600, na.rm = TRUE),
         K600_sd   = sd(K600, na.rm = TRUE),
         K600_flag = if_else(K600 > K600_mean + 3 * K600_sd, TRUE, FALSE)) %>%
  ungroup()

filtered_data_clean <- filtered_data_flagged %>%
  filter(!K600_flag)

# results <- cor_plot_K600_ER_discharge(filtered_data)

```



#### Gap fill

'fill_and_normalize_metab.R' is a modified function
from'code/metabolism_synthesis/Part II- Filtering and gap filling.Rmd
uses the function fillMiss3.R

**Make sure** to double check the 'summary_site-year_NAs.csv' save path
in the fxn

-   Extracts unique sites and years from the input data.

-   Tracks missing data for each site-year.

-   Pads the dataset to ensure a full-year time series.

-   Identifies missing data patterns and determines if a year is
    complete.

-   Applies a gap-filling algorithm (if \<40% of data is missing).

-   Outputs a cleaned and gap-filled dataset for further analysis.



#### Histogram of infilled GPP 

```{r, warning=F, size = 'small', echo=F, include=T}
new_dat <- filtered_data %>%
  dplyr::select(site:shortwave)

river_metab_filled <- fill_and_normalize_metab1(new_dat)
head(river_metab_filled)

ggplot(data = river_metab_filled) +
  geom_histogram(aes(x = GPP), bins = 30, fill = "green4", color = "black") +
    geom_histogram(aes(x = GPP_filled), bins = 30, fill = "#227066") +
  facet_wrap(. ~ site) +
  theme_minimal()

```

#### Histogram of infilled ER 


```{r, warning=F, size = 'small', echo=F, include=T}

river_metab_fill2 <- fill_and_normalize_metab_ER(river_metab_filled)
head(river_metab_fill2)

ggplot(data = river_metab_fill2) +
  geom_histogram(aes(x = ER), bins = 30, fill = "goldenrod", color = "black") +
    geom_histogram(aes(x = ER_filled), bins = 30, fill = "goldenrod1") +
  facet_wrap(. ~ site) +
  theme_minimal()

```


```{r, warning=F, size = 'small', echo=F, include=F}

library(dplyr)

# 
date_site_counts <- river_metab_fill2 %>%
  distinct(site, date) %>%
  count(date, name = "n_sites")

# 
n_total_sites <- river_metab_fill2 %>%
  distinct(site) %>%
  nrow()

# 
complete_dates <- date_site_counts %>%
  pull(date)

# 
river_metab_matched <- river_metab_fill2 %>%
  filter(date %in% complete_dates)

names(river_metab_filled)

```





#### Gap fill

'fill_and_normalize_metab.R' is a modified function
from'code/metabolism_synthesis/Part II- Filtering and gap filling.Rmd
uses the function fillMiss3.R

**Make sure** to double check the 'summary_site-year_NAs.csv' save path
in the fxn

-   Extracts unique sites and years from the input data.

-   Tracks missing data for each site-year.

-   Pads the dataset to ensure a full-year time series.

-   Identifies missing data patterns and determines if a year is
    complete.

-   Applies a gap-filling algorithm (if \<40% of data is missing).

-   Outputs a cleaned and gap-filled dataset for further analysis.



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

# Load Packages
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables

# load data package with lat/long

RR_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/RR_dat.csv")%>%
  mutate(date= as.Date(date)) %>%
  dplyr::select(Site_ID, Lat, Lon, date, year, GPP_filled, discharge, temp.water, 
                ppt_mm, TmeanC)  

unique(RR_dat$Site_ID)

RR_dat <- RR_dat %>%
  mutate(Site_ID = str_remove(Site_ID, "^nwis_"))

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

unique(river_metab_matched$site)

river_metab_matched <- river_metab_matched %>%
  mutate(Site_ID = str_remove(site, "^nwis_"))

river_metab_matched <- river_metab_matched %>%
  mutate(Site_ID = str_remove(Site_ID, "^0+"))

stream_cords <- RR_dat%>%
  dplyr::select(Site_ID, Lon, Lat) %>%
  group_by(Site_ID, Lon, Lat)%>%
  summarise(Lon=mean(Lon),
            Lat=mean(Lat))
   
stream_cords <- as.data.frame(stream_cords)
str(stream_cords)

# Data with correct values
fix_data <- data.frame(
  Site_ID = c("10129900", "10133650", "10133800", "11044000", "14206950", "14211010"),
  Lat = c(40.73522634, 40.7238375, 40.75966979, 33.4739175, 45.403452, 45.3792874),
  Lon = c(-111.4760313, -111.5196438, -111.5640912, -117.1422536, -122.7548185, -122.5773134)
)

# Join and replace missing values
df_fixed <- stream_cords %>%
  left_join(fix_data, by = "Site_ID", suffix = c("", ".fix")) %>%
  mutate(
    Lat = if_else(is.na(Lat), Lat.fix, Lat),
    Lon = if_else(is.na(Lon), Lon.fix, Lon)
  ) %>%
  dplyr::select(-Lat.fix, -Lon.fix)

river_metab_dat <- river_metab_matched %>%
  left_join(df_fixed, by = c("Site_ID" = "Site_ID"))

river_metab_dat <- as.data.frame(river_metab_dat)

river_metab_dat <-  river_metab_dat %>%
  mutate(week=week(date),
         year = year(date))

# RDS save 
# 
# med_fill <-  river_metab_dat %>%
#   mutate(week=week(date))%>%
#   group_by(Site_ID, week, year, Lat, Lon) %>%
#   summarise(weekly_med = median(GPP_filled, na.rm=T),
#             weekly_er = median(ER_filled, na.rm=T))
# 
# river_metab_dat_filled <- river_metab_dat %>%
#   left_join(med_fill, by = c("Site_ID", "week", "year", "Lat", "Lon")) %>%
#   mutate(GPP_filled = ifelse(is.na(GPP_filled), weekly_med, GPP_filled),
#          ER_filled = ifelse(is.na(ER_filled), weekly_er, ER_filled)) %>%
#   dplyr::select(-weekly_med)

```
