---
title: "Spline Correlogram analysis for synchrony in aquatic ecosystem productivity"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```


```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables
library(broom)
library(grid)

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

river_metab_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/stream_spline_cor_df.rds")

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
### read in prism data for PPT

PPT_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Clim_data/PRISM_stream.rds")%>%
  mutate(Site_ID=as.character(Name))

river_metab_dat <- river_metab_dat%>%
  left_join(PPT_dat, by=c("date"="Date", "Site_ID")) %>%
  dplyr::select(site:TmaxC)

```

## Part 1: Synchrony in metabolism at differnt timescales

Compare synchrony for GPP and ER from 56 different sites in Marzolf dataset at different timescales:

- Daily

- Weekly (mean)

- Monthly (mean)

- Quarterly (3 month mean)

- Yearly (mean)


### A. GPP synchrony at differnt timescales:


```{r, warning=F, size = 'small', echo=T, include=T}

#### Daily synchrony

river_metab_dat<-as.data.frame(river_metab_dat)

GPP_long <- river_metab_dat %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day <- GPP_long %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_day <- Sncf(x=GPP_wide_day$Lon,
                   y=GPP_wide_day$Lat,
                   z=GPP_wide_day[,4:ncol(GPP_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year

RR_df.corr_day

plot(RR_df.corr_day)


GPP_corrgram_summary.table <- summary(RR_df.corr_day)

GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP")

## Sample average line
GPP.corr.pred.df.day$mean_synch <- mean(GPP.corr.pred.df.day$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP")

```



```{r, warning=F, size = 'small', echo=F, include=F}
#### Weekly synchrony

GPP_week <-  river_metab_dat %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))

GPP_wide_week <- GPP_week %>%
  dplyr::select(Site_ID, Lon, Lat, week, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }

GPP_wide_week_1 <-na.omit(GPP_wide_week)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=GPP_wide_week$Lon,
                   y=GPP_wide_week$Lat,
                   z=GPP_wide_week[,4:ncol(GPP_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)


GPP_corrgram_summary.table <- summary(RR_df.corr_wk)
GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly",
                                     parameter = "GPP")

## Sample average line
GPP.corr.pred.df.wk$mean_synch <- mean(GPP.corr.pred.df.wk$y, na.rm = TRUE)


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly",
                                     parameter = "GPP")
```



```{r, warning=F, size = 'small', echo=F, include=F}
#### Monthly synchrony

GPP_month <-  river_metab_dat %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))


GPP_wide_month <- GPP_month %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=GPP_wide_month$Lon,
                   y=GPP_wide_month$Lat,
                   z=GPP_wide_month[,4:ncol(GPP_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


GPP_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly",
                                   parameter = "GPP")

## Sample average line
GPP.corr.pred.df.mon$mean_synch <- mean(GPP.corr.pred.df.mon$y, na.rm = TRUE)

GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                time_scale = "monthly",
                                   parameter = "GPP")
```



```{r, warning=F, size = 'small', echo=F, include=F}
#### Yearly synchrony

GPP_yr <- river_metab_dat %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_yr <- GPP_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, GPP_filled) %>%
  pivot_wider(names_from = c(year), values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=GPP_wide_yr$Lon,
                   y=GPP_wide_yr$Lat,
                   z=GPP_wide_yr[,4:ncol(GPP_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

GPP_corrgram_summary.table <- summary(RR_df.corr_yr)
GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly",
                                  parameter = "GPP")


## Sample average line
GPP.corr.pred.df.yr$mean_synch <- mean(GPP.corr.pred.df.yr$y, na.rm = TRUE)


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly",
                                     parameter = "GPP")

```




```{r, warning=F, size = 'small', echo=F, include=F}
#### Quarterly synch

GPP_qrt <- river_metab_dat %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_qrt <- GPP_qrt %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=GPP_wide_qrt$Lon,
                   y=GPP_wide_qrt$Lat,
                   z=GPP_wide_qrt[,4:ncol(GPP_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)

GPP_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly",   
                                parameter = "GPP")
## Sample average line
GPP.corr.pred.df.qrtly$mean_synch <- mean(GPP.corr.pred.df.qrtly$y, na.rm = TRUE)

GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly",    
                               parameter = "GPP")


```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
##### combined synchonry df 
pred_df_all <- bind_rows(GPP.corr.pred.df.day, GPP.corr.pred.df.wk, GPP.corr.pred.df.mon, GPP.corr.pred.df.qrtly, GPP.corr.pred.df.yr)
ci_df_all   <- bind_rows(GPP.corr.boot.df.day, GPP.corr.boot.df.wk, GPP.corr.boot.df.mon, GPP.corr.boot.df.qrtly, GPP.corr.boot.df.yr)

```

##### Calculate the percent difference between time steps

```{r, echo=T, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each time_scale becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x, y, time_scale) %>%
  pivot_wider(names_from = time_scale, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )

pred_diff
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
scale_colors <- c(
  "daily" = "#E69F00",     
  "quarterly" = "#56B4E9",    
  "monthly" = "#009E73",   
  "weekly" = "#F0E442", 
  "yearly" = "#0072B2"    
)
```


##### Figure of precent difference in snych signal 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Pivot longer
pred_long_diff <- pred_diff %>%
  dplyr::select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )

# Plot histogram
ggplot(pred_long_diff, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```

### Figure 2 (?): Spline Correlogram of GPP at different timescales

Points represent site median and 95% quantile of GPP (for the whole record). 
They are included to get a sense of geographic distance of predicted values.

** GPP is arbitrarily transformed (/ 25) to scale along the slpine.  

```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)

### add point for mean GPP to plot 
library(geosphere)
# Get site centroids
site_coords <- GPP_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (the first one in the dataset)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

GPP_long_dist <- GPP_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")


GPP_long_dist_plt <- GPP_long_dist %>%
  group_by(Site_ID, distance_km) %>%
  summarise(
    GPP_mean = median(GPP_filled, na.rm = TRUE), 
    GPP_low = quantile(GPP_filled, probs = 0.25, na.rm = TRUE), 
    GPP_up = quantile(GPP_filled, probs = 0.75, na.rm = TRUE),
    .groups = "drop"
  )

###
GPP_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%filter(x<3250),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%filter(x<3250),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all%>%filter(x<3100),
             aes(yintercept = mean_synch, color = time_scale),
             linetype = "dashed") +
  # Add GPP points (mean)
  geom_point(data = GPP_long_dist_plt,
             aes(x = distance_km, y = GPP_mean/25),
             shape = 1, size = 2,
             alpha=0.5) +
    ylim(-0.2, 0.75) +
  # Add GPP error bars
  geom_errorbar(data = GPP_long_dist_plt,
                aes(x = distance_km, ymin = GPP_low/25, ymax = GPP_up/25), 
                width = 2, color = "black", 
                alpha=0.25) +
  # Labels and theme
  labs(title = "Synchrony in GPP across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff%>%filter(x<3250), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(5, 5, 5, 5),
        legend.position="none")

# Combine using patchwork (put rotated density plot on the right)
combined_plot <- GPP_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```


```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/GPP_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```


#### GPP: Look when synchrony dips below the sample average for each timescale:

This gives an idea for at what distances stream GPP is no longer synchronous 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df_all %>%
  mutate(
    mean_synch = round(mean_synch, 2),
    y = round(y, 2),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(time_scale) %>%
  # Filter only rows where y is approximately equal to mean_synch
 dplyr:: filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```

Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

max_synch <- pred_df_all %>%
  group_by(time_scale) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

max_synch
```


### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}
library(dplyr)
library(purrr)
library(broom)

fit_models <- pred_df_all %>%
  group_by(time_scale) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(time_scale, term,rounded_estimate: rounded_p.value)

fit_models_tab


```




### B. ER synchrony at differnt timescales:


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
#### Daily synchrony:


river_metab_dat<-as.data.frame(river_metab_dat)

ER_long <- river_metab_dat %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day <- ER_long %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
ER_df.corr_day <- Sncf(x=ER_wide_day$Lon,
                   y=ER_wide_day$Lat,
                   z=ER_wide_day[,4:ncol(ER_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(ER_df.corr_day)


ER_corrgram_summary.table <- summary(ER_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(ER_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                                  parameter = "ER")


## Sample average line
ER.corr.pred.df.day$mean_synch <- mean(ER.corr.pred.df.day$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(ER_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "daily",     
                               parameter = "ER")


```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
#### Weekly synchrony


ER_week <-  river_metab_dat %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

ER_wide_week <- ER_week %>%
  dplyr::select(Site_ID,Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled)

river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_week <- ER_week %>%
 # filter(year>2011) %>%
 dplyr::select(Site_ID, Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=ER_wide_week$Lon,
                   y=ER_wide_week$Lat,
                   z=ER_wide_week[,4:ncol(ER_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)



ER_corrgram_summary.table <- summary(RR_df.corr_wk)
ER_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly",     
                                 parameter = "ER")
## Sample average line
ER.corr.pred.df.wk$mean_synch <- mean(ER.corr.pred.df.wk$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly",
                                     parameter = "ER")
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

#### Monthly synchrony


ER_month <-  river_metab_dat %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_month <- ER_month %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, ER_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=ER_wide_month$Lon,
                   y=ER_wide_month$Lat,
                   z=ER_wide_month[,4:ncol(ER_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


ER_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly",
                                    parameter = "ER")

## Sample average line
ER.corr.pred.df.mon$mean_synch <- mean(ER.corr.pred.df.mon$y, na.rm = TRUE)

ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                time_scale = "monthly",
                                    parameter = "ER")

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
#### Quarterly synchrony


ER_qrt <- river_metab_dat %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_qrt <- ER_qrt %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, ER_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=ER_wide_qrt$Lon,
                   y=ER_wide_qrt$Lat,
                   z=ER_wide_qrt[,4:ncol(ER_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)


ER_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly",     
                                parameter = "ER")

## Sample average line
ER.corr.pred.df.qrtly$mean_synch <- mean(ER.corr.pred.df.qrtly$y, na.rm = TRUE)


ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly",     
                               parameter = "ER")


```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

#### Yearly synchrony


ER_yr <- river_metab_dat %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


ER_wide_yr <- ER_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, ER_filled) %>%
  pivot_wider(names_from = c(year), values_from = ER_filled)



# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=ER_wide_yr$Lon,
                   y=ER_wide_yr$Lat,
                   z=ER_wide_yr[,4:ncol(ER_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

ER_corrgram_summary.table <- summary(RR_df.corr_yr)
ER_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly",
                                   parameter = "ER")

## Sample average line
ER.corr.pred.df.yr$mean_synch <- mean(ER.corr.pred.df.yr$y, na.rm = TRUE)


ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly",    
                                 parameter = "ER")

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
##### combined synch 


pred_df_all_ER <- bind_rows(ER.corr.pred.df.day, ER.corr.pred.df.wk, ER.corr.pred.df.mon, ER.corr.pred.df.qrtly, ER.corr.pred.df.yr)
ci_df_all_ER   <- bind_rows(ER.corr.boot.df.day, ER.corr.boot.df.wk, ER.corr.boot.df.mon, ER.corr.boot.df.qrtly, ER.corr.boot.df.yr)

```

#### ER: Look when synchrony dips below the sample average

This gives an idea for at what distances stream ER is no longer synchronous 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df_all_ER %>%
  mutate(
    mean_synch = round(mean_synch, 2),
    y = round(y, 2),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(time_scale) %>%
  # Filter only rows where y is approximately equal to mean_synch
  dplyr::filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```

Looks like ER snychrony breaks down at 1105 km for daily dynamics and 2429 km for yearly (which is bascall the whole length of the US) 



### Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

max_synch <- pred_df_all_ER %>%
  group_by(time_scale) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

max_synch
```


### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}
library(dplyr)
library(purrr)
library(broom)

fit_models <- pred_df_all_ER %>%
  group_by(time_scale) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(time_scale, term,rounded_estimate: rounded_p.value)

fit_models_tab


```


```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

ggplot() +
  geom_ribbon(data = ci_df_all_ER,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  geom_line(data = pred_df_all_ER,
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  ##
    labs(title = "Synchrony in ER across different time scales", 
       y= "Correlation", 
       x="Distance (km)") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic()


```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=4}


# Spread the data so each time_scale becomes a column
pred_wide_ER <- pred_df_all_ER %>%
  dplyr::select(x,y,time_scale)%>%
  pivot_wider(names_from = time_scale, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_ER <- pred_wide_ER %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )



# Pivot longer
pred_long_diff_ER <- pred_diff_ER %>%
 dplyr:: select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )


# Plot histogram
ggplot(pred_long_diff_ER, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```


### Figure 2/3 (?): Spline Correlogram of ER at different timescales

Points represent site median and 95% quantile of |ER| (for the whole record). 
They are included to get a sense of geographic distance of predicted values.

** ER is arbitrarily transformed (/ 25) to scale along the slpine.  


```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)

### add point for mean ER to plot 
library(geosphere)

site_coords <- ER_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (e.g., the upstream-most one)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

ER_long_dist <- ER_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")


ER_long_dist_plt <- ER_long_dist %>%
  group_by(Site_ID, distance_km) %>%
  summarise(
    ER_mean = (median(ER_filled, na.rm = TRUE)*-1), 
    ER_low = (quantile(ER_filled, probs = 0.25, na.rm = TRUE)*-1), 
    ER_up = (quantile(ER_filled, probs = 0.75, na.rm = TRUE)*-1),
    .groups = "drop"
  )

###
ER_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_ER%>%filter(x<3250),
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_ER%>%filter(x<3250),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all_ER%>%filter(x<3100),
             aes(yintercept = mean_synch, color = time_scale),
             linetype = "dashed") +
  # Add ER points (mean)
  geom_point(data = ER_long_dist_plt,
             aes(x = distance_km, y = ER_mean/25),
             shape = 1, size = 2,
             alpha=0.5) +
  # Add ER error bars
  geom_errorbar(data = ER_long_dist_plt,
                aes(x = distance_km, ymin = ER_low/25, ymax = ER_up/25),
                width = 2, color = "black", 
                alpha=0.25) +
  ylim(-0.2, 0.75) +
  # Labels and theme
  labs(title = "Synchrony in ER across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff_ER%>%filter(x<3250), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(5, 5, 5, 5),
        legend.position="none")

# Combine using patchwork (put rotated density plot on the right)
combined_plot <- ER_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```


```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/ER_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```



## Part 2 - Moran Effect

- MODIS data for leaf area index (LAI)

- Daily streamflow (Q)

- Daily water temperature (water_temp)  

- Daily air temp (air_temp)

- Daily precipitation (PPT)


#### Bring in MODIS data for LAI to get at 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

LAI_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Modis_data/StreamDat_MOD_combined.rds")%>%
  mutate(date=as.Date(Date))

```

#### LAI Daily synch (filled)

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}

str(LAI_dat)

stream_LAI_dat <- LAI_dat %>%
  left_join(river_metab_dat, by = c("Site_ID", "date"))

### 

LAI_long <- stream_LAI_dat %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(Lai = mean(Lai, na.rm = TRUE), .groups = "drop")


LAI_filled_coords <- LAI_long %>%
  group_by(Site_ID) %>%
  fill(Lat, Lon, .direction = "downup") %>%
  ungroup()


LAI_filled_coords1 <- LAI_filled_coords %>%
  group_by(Site_ID) %>%
  fill(Lai, .direction = "downup") %>%
  ungroup()

LAI_wide_day <- LAI_filled_coords1 %>%
  pivot_wider(
    names_from = date,
    values_from = Lai)

LAI_wide_day<- na.omit(LAI_wide_day)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
LAI_df.corr_day <- Sncf(x=LAI_wide_day$Lon,
                   y=LAI_wide_day$Lat,
                   z=LAI_wide_day[,4:ncol(LAI_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(LAI_df.corr_day)


LAI_corrgram_summary.table <- summary(LAI_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
LAI.corr.pred.df.day <- data.frame(x = matrix(unlist(LAI_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(LAI_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter="LAI")
## Sample average line
LAI.corr.pred.df.day$mean_synch <- mean(LAI.corr.pred.df.day$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
LAI.corr.bootValues.df <- as.data.frame(t(LAI_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
LAI.corr.boot.df.day <- data.frame(x = matrix(unlist(LAI_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminLAI = LAI.corr.bootValues.df$`0.025`,
                                             ymaxLAI = LAI.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                parameter="LAI")

```


#### Daily flow:

```{r, size='small', echo=F, include=T, warning=FALSE}
### look at an element of environmental variation -- lets look at spring flow
# RR_dat_sum <-  river_metab_dat %>%
#   mutate(month=month(date)) %>%
#   dplyr::filter(month>1 & month<=6) %>%
#   dplyr::group_by(Site_ID, year, Lon, Lat) %>%
#   summarise(Q_max= max(discharge, na.rm=T),
#             Q_anl= mean(discharge, na.rm=T))

# choosing max flow:
RR_dat_TS_Q <- river_metab_dat %>%
  dplyr::select(Site_ID,Lon, Lat, date, discharge) %>%
  pivot_wider(names_from = c(date), values_from = discharge)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
Q_RR_df.corr <- Sncf(x=RR_dat_TS_Q$Lon,
                   y=RR_dat_TS_Q$Lat,
                   z=RR_dat_TS_Q[,4:ncol(RR_dat_TS_Q)],
                   resamp = 1000,
                   latlon = T,
                     na.rm = T)

plot(Q_RR_df.corr)

Q_corrgram_summary.table <- summary(Q_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Q.corr.pred.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$real$predicted$x)),
                               y = matrix(unlist(Q_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="Q")

## Sample average line
Q.corr.pred.df$mean_synch <- mean(Q.corr.pred.df$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
Q.corr.bootValues.df <- as.data.frame(t(Q_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Q.corr.boot.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$boot$boot.summary$predicted$x)),
                               yminQ = Q.corr.bootValues.df$`0.025`,
                               ymaxQ = Q.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="Q")

```


#### Daily water temperature:


```{r, size='small', echo=F, include=T, warning=FALSE}

### Water temperature
RR_dat_TS_temp <- river_metab_dat %>%
  dplyr::select(Site_ID, Lon, Lat, date, temp.water) %>%
  pivot_wider(names_from = c(date), values_from = temp.water)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
temp_RR_df.corr <- Sncf(x=RR_dat_TS_temp$Lon,
                     y=RR_dat_TS_temp$Lat,
                     z=RR_dat_TS_temp[,4:ncol(RR_dat_TS_temp)],
                     resamp = 1000,
                     latlon = T,
                     na.rm = T)

plot(temp_RR_df.corr)
temp_corrgram_summary.table <- summary(temp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
temp.corr.pred.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$real$predicted$x)),
                             y = matrix(unlist(temp_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="water_temp")

## Sample average line
temp.corr.pred.df$mean_synch <- mean(temp.corr.pred.df$y, na.rm = TRUE)


# ...and a dataframe of the y values for the bootstrap prediction
temp.corr.bootValues.df <- as.data.frame(t(temp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
temp.corr.boot.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$boot$boot.summary$predicted$x)),
                             yminWtemp = temp.corr.bootValues.df$`0.025`,
                             ymaxWtemp = temp.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="water_temp")
```


#### Daily air temperature:


```{r, size='small', echo=T, include=T, warning=FALSE}

### Water temperature
RR_dat_TS_Atemp <- river_metab_dat %>%
  dplyr::select(Site_ID, Lon, Lat, date, TmeanC) %>%
  pivot_wider(names_from = c(date), values_from = TmeanC)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
Atemp_RR_df.corr <- Sncf(x=RR_dat_TS_Atemp$Lon,
                     y=RR_dat_TS_Atemp$Lat,
                     z=RR_dat_TS_Atemp[,4:ncol(RR_dat_TS_Atemp)],
                     resamp = 1000,
                     latlon = T,
                     na.rm = T)

plot(Atemp_RR_df.corr)
Atemp_corrgram_summary.table <- summary(Atemp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Atemp.corr.pred.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$real$predicted$x)),
                             y = matrix(unlist(Atemp_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="air_temp")

## Sample average line
Atemp.corr.pred.df$mean_synch <- mean(Atemp.corr.pred.df$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
Atemp.corr.bootValues.df <- as.data.frame(t(Atemp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Atemp.corr.boot.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$boot$boot.summary$predicted$x)),
                             yminAtemp = Atemp.corr.bootValues.df$`0.025`,
                             ymaxAtemp = Atemp.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="air_temp")
```


#### Daily PPT:


```{r, size='small', echo=T, include=T, warning=FALSE}

RR_dat_TS_ppt <- river_metab_dat %>%
  dplyr::select(Site_ID, Lon, Lat, date, ppt_mm) %>%
  pivot_wider(names_from = c(date), values_from = ppt_mm)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
ppt_RR_df.corr <- Sncf(x=RR_dat_TS_ppt$Lon,
                        y=RR_dat_TS_ppt$Lat,
                        z=RR_dat_TS_ppt[,4:ncol(RR_dat_TS_ppt)],
                        resamp = 1000,
                        latlon = T,
                       na.rm = T)

plot(ppt_RR_df.corr)

ppt_corrgram_summary.table <- summary(ppt_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
ppt.corr.pred.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$real$predicted$x)),
                                y = matrix(unlist(ppt_RR_df.corr$real$predicted$y)),
                            time_scale = "daily",
                             parameter="PPT")

## Sample average line
ppt.corr.pred.df$mean_synch <- mean(ppt.corr.pred.df$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
ppt.corr.bootValues.df <- as.data.frame(t(ppt_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ppt.corr.boot.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$boot$boot.summary$predicted$x)),
                                yminPPT = ppt.corr.bootValues.df$`0.025`,
                                ymaxPPT = ppt.corr.bootValues.df$`0.975`,
                                    time_scale = "daily",
                             parameter="PPT")

```



```{r, size='small', echo=F, include=F, warning=FALSE}

big_boot_df<- GPP.corr.boot.df.day %>%
  left_join(Q.corr.boot.df, by=c("x")) %>%
  left_join(temp.corr.boot.df, by=c("x")) %>%
  left_join(ppt.corr.boot.df, by=c("x")) %>%
  left_join(Atemp.corr.boot.df, by=c("x")) 



# Create a vector of unique x values from big_boot_df
big_x_vals <- unique(big_boot_df$x)

# Add a column to LAI.corr.boot.df.day with the nearest x from big_boot_df
LAI.corr.boot.df.day <- LAI.corr.boot.df.day %>%
  mutate(x_nearest = map_dbl(x, ~big_x_vals[which.min(abs(big_x_vals - .x))]))

# Now join with big_boot_df using the nearest x value
joined_df <- LAI.corr.boot.df.day %>%
  left_join(big_boot_df, by = c("x_nearest" = "x"))

```

## correlogram of Synchrony of GPP + other hydro/climate/environmental parameters near streams

```{r, size = 'large', echo=FALSE, include=F, warning=FALSE}

pred_df <- bind_rows(
  Q.corr.pred.df %>% mutate(group = "Flow"),
  GPP.corr.pred.df.day %>% mutate(group = "GPP"),
  temp.corr.pred.df %>% mutate(group = "Water Temp"),
  ppt.corr.pred.df %>% mutate(group = "PPT"),
  Atemp.corr.pred.df %>% mutate(group = "Air Temp"),
 LAI.corr.pred.df.day %>% mutate(group = "LAI")
)

```



```{r, size = 'large', echo=FALSE, include=T, warning=FALSE}
ribbon_df <- joined_df %>%
  transmute(
    x = x,
    Flow_ymin = yminQ, Flow_ymax = ymaxQ,
    GPP_ymin = yminGPP, GPP_ymax = ymaxGPP,
    WaterTemp_ymin = yminWtemp, WaterTemp_ymax = ymaxWtemp,
    PPT_ymin = yminPPT, PPT_ymax = ymaxPPT,
    AirTemp_ymin = yminAtemp, AirTemp_ymax = ymaxAtemp,
    yminLAI = yminLAI, ymaxLAI = ymaxLAI
  ) %>%
  pivot_longer(cols = -x, names_to = c("group", ".value"), names_sep = "_")

synch.plot <- ggplot() +
      geom_line(data = pred_df%>%dplyr::filter(x<3200),
            aes(x = x, y = mean_synch, color = group), lty=2, alpha=0.75) +
  geom_line(data = pred_df%>%dplyr::filter(x<3200),
            aes(x = x, y = y, color = group)) +
  geom_hline(yintercept = 0) +
  theme_classic() +
  labs(title = "Synchrony of environmental conditions",
              y= "Correlation",
       x="Distance (km)",
       color = "Variable", fill = "Variable") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size=14),
        legend.position = "bottom") +
  scale_color_manual(values = c(
    "Flow" = "#0072B2",
    "GPP" = "#009E73",
    "Water Temp" = "#D55E00",
    "PPT" = "#56B4E9",
    "Air Temp" = "#CC79A7",
    "LAI" = "#E2D523"
  )) +
   geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminQ, ymax = ymaxQ),
              fill = "#0072B2", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
              fill = "#009E73", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminWtemp, ymax = ymaxWtemp),
              fill = "#D55E00", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminPPT, ymax = ymaxPPT),
              fill = "#56B4E9", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminAtemp, ymax = ymaxAtemp),
              fill = "#CC79A7", alpha=0.2)+
    geom_ribbon(data = joined_df%>%dplyr::filter(x<3200),
              aes(x = x, ymin = yminLAI, ymax = ymaxLAI),
              fill = "#E2D523", alpha=0.2)

synch.plot

```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Moran_spline_correlogram.png", plot = synch.plot, width =5.5, height = 4.5, units = "in")

```


### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}
library(dplyr)
library(purrr)
library(broom)

fit_models <- pred_df %>%
  group_by(parameter) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params)

fit_models

```



| Variable         | a (initial synch) | b (decay rate) | Interpretation                                                                                               |
| ---------------- | ----------------: | -------------: | -------------------------------------------------------------------------------------------------------------|
| **Q (Flow)**     |            0.6937 |        0.00223 | Starts moderately synchronized, but decays rapidly over space.                                               |
| **GPP**          |            0.2597 |       0.000037 | Low synchrony, and it barely decays — very stable across space, suggesting local factors                     |
| **Water Temp**   |            0.9361 |      0.0000285 | High synchrony,  — very stable across space                                                                  |
| **PPT (Precip)** |            0.9009 |        0.00277 | High synchrony, but decays quickly —  similar to flow. Tight confidence intervals.                           |
| **Air Temp**     |            0.9853 |       0.000103 | The most synchronous slowly decaying — regional coherence in air temp. Tight confidence intervals            |
| **LAI**          |            0.6879 |       0.000153 | Moderate synchrony and mild decay — stable across space, suggesting local factors                            |


### Half-distance

Calculate the distance at which synchrony drops to half of its max value:


```{r, warning=F, size = 'small', echo=T, include=T}

half_distance <- pred_df %>%
  dplyr::group_by(parameter) %>%
  summarise(
    #max_synch = max(y, na.rm = TRUE),
    mean_synch = mean(y, na.rm = TRUE),
    half_synch = mean_synch / 2
  ) %>%
  left_join(pred_df, by = "parameter") %>%
  dplyr::group_by(parameter) %>%
  dplyr::filter(y <= half_synch) %>%
  slice_min(x) %>%
  dplyr::select(parameter, x, y) %>%
  rename(half_distance = x)

half_distance
```

#### Pull boot-strapped max synch.


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

max_synch <- pred_df %>%
  group_by(parameter) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

```


#### Look at to see when each spline dips below the sample average. 

This gives an idea for at what distances stream GPP is no longer synchronous 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df %>%
  mutate(
    mean_synch = round(mean_synch, 2),
    y = round(y, 2),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(time_scale) %>%
  # Filter only rows where y is approximately equal to mean_synch
# dplyr::filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```



```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df %>%
  mutate(
    mean_synch = round(mean_synch, 2),
    y = round(y, 2),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(parameter) %>%
  # Filter only rows where y is approximately equal to mean_synch
  dplyr::filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```


LAI looks a like it 1/2 decay rate might be biased by longer distance bump. 
That sample mean might be better 


## Part 3. Landscape level factors as explaining synchrony:

A. left join time series dat with landscape parameters 

B. ER / GPP explained by stream size (via regression)
  
  i. Dominant land cover explained by stream size 
  
  ii. Geology PCA
  
  iii. Slope/ Catchment size/ basin characteristics
  
  iv. Water temperature/ flow 
    
C. Create landscape level groups and then run synchronous
  
  i. landcover (PCA? developed to forest?) - possibly cropland with moderate development
  
  ii. Geology (PCA?) - possibly metamorphic
  
  iii. Slope? or Catchment size 
  
  iv. Water temperature/ depth/flow? 
    
5. When landscape character explains some synchronous, visualize with synch splines (Q1-Q3 for i to iv splines) 


```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

dat_NHDinfo<- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata_25v2.rds")%>%
  mutate(Site_ID = str_remove(site_no, "^0+"))

river_metab_dat1 <- river_metab_dat%>%
  left_join(dat_NHDinfo, by=c("Site_ID")) 

```

#### A. Random forest to get at what should describe National Land Cover Database

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}
library(randomForest)

# Prepare data for RF
rf_LC_data <- river_metab_dat1[, c("Developed_OpenSpace_pct", "Developed_LowIntensity_pct",
                       "Developed_MedIntensity_pct", "Developed_HiIntensity_pct",
                       "Barren_Land_pct", "Forest_Deciduous_pct",
                       "Water_Open_pct", "Ice_Snow_Perennial_pct",
                       "Shrub_Scrub_pct", "Grass_landHerbaceous_pct",
                       "Forest_Evergreen_pct", "Forest_mixed_pct",
                       "Grass_PastureHay_pct", "Cultivated_Crops_pct",
                       "Wetlands_Woody_pct","Wetlands_EmergentHerb_pct",
                       "road_Rural_den", "road_private_den", "road_Total_den")]


rf_LC_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_LC_data)
rf_LC_input_clean <- na.omit(rf_LC_input)

# Now run the model
rf_LC_model <- randomForest(GPP_synch ~ ., data = rf_LC_input_clean,
                         importance = TRUE, ntree = 200)

# Plot importance
varImpPlot(rf_LC_model)

# Get importance metrics from your existing model
importance_vals <- importance(rf_LC_model, type = 1)  # type = 1 for %IncMSE
importance_df <- data.frame(variable = rownames(importance_vals),
                            IncMSE = importance_vals[, 1])

# Sort by importance and keep top 15
top_vars_LC <- importance_df %>%
  arrange(desc(IncMSE)) %>%
  slice(1:10) %>%
  pull(variable)

top_vars_LC
# 
# rf_LC_input_reduced <- rf_LC_input_clean %>%
#   dplyr::select(GPP_synch, all_of(top_vars_LC))
# 
# rf_LC_model_reduced <- randomForest(GPP_synch ~ ., data = rf_LC_input_reduced,
#                                  importance = TRUE, ntree = 500)
# 
# varImpPlot(rf_LC_model_reduced)
# 
# # Sort by importance and keep top 10
# top_vars_LC <- importance_df %>%
#   arrange(desc(IncMSE)) %>%
#   slice(1:5) %>%
#   pull(variable)
# 
# top_vars_LC

```

### Create the composite variables for land cover/geology/basin characteristics

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Categorize summary columns based on their initial patterns
col_patterns <- c("alt_va" = "reach morphology", "drain_area_va" = "drainage_area", "Water_Open_pct" = "water",
                  "Ice_Snow_Perennial_pct" = "ice_snow", "Developed_OpenSpace_pct" = "open_space",
                  "Developed_LowIntensity_pct" = "open_space", "Developed_MedIntensity_pct" = "developed",
                  "Developed_HiIntensity_pct" = "developed", "Barren_Land_pct" = "open_space",
                  "Forest_Deciduous_pct" = "forest", "Forest_Evergreen_pct" = "forest", 
                  "Forest_mixed_pct" = "forest", "Shrub_Scrub_pct" = "shrub", 
                  "Grass_landHerbaceous_pct" = "shrub", "Grass_PastureHay_pct" = "crops", 
                  "Cultivated_Crops_pct" = "crops", "Wetlands_Woody_pct" = "wetlands", 
                  "Wetlands_EmergentHerb_pct" = "wetlands", "geol_gneiss" = "geology_metamorphic", 
                  "geol_granitic" = "geology_igneous", "geol_ultramafic" = "geology_metamorphic", 
                  "geol_quarternary" = "geology_sedimentary", "geol_sedimentary" = "geology_sedimentary", 
                  "geol_volcanic" = "geology_igneous", "geol_intermediate_pultonic" = "geology_igneous", 
                  "CAT_BASIN_AREA" = "drainage_area", "CAT_BASIN_SLOPE" = "slope", 
                  "CAT_ELEV_MEAN" = "elevation", "CAT_ELEV_MIN" = "elevation", 
                  "CAT_ELEV_MAX" = "elevation", "CAT_STREAM_SLOPE" = "reach morphology", 
                  "CAT_STREAM_LENGTH" = "stream_length", "road_Rural_den" = "road_density", 
                  "road_private_den" = "road_density", "road_Total_den" = "road_density", 
                  "Tree_Canopyin100mRip" = "shading", 
                  "temp.water" = "water quality", "DO.obs" = "water quality",
                  "depth" = "reach morphology", "discharge" = "discharge",
                  "pathlength" = "hydrologic hierarchy", "lengthkm" = "geometry",
                 "shape_length"=  "geometry", "streamleve"= "hydrologic hierarchy",
                 "streamorde"=  "hydrologic hierarchy", 
                 "streamcalc" = "hydrologic hierarchy", 
                 "totdasqkm" = "drainage_area", "slope" = "slope", "pathlength" = "hydrologic hierarchy" )

```


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=6}
### start the PCA 

LC_pca_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID, all_of(top_vars_LC))

LC_pca_data1 <- LC_pca_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
LC_pca_data <- LC_pca_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

LC_pca_input_clean <- na.omit(LC_pca_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(LC_pca_data), function(col) {
  col_patterns[col]
}))

pca_LC <- prcomp(LC_pca_input_clean, center = TRUE, scale. = TRUE)

library(factoextra)

# Visualize PCA eigenvalues
scree_plt_LC <- fviz_eig(pca_LC)
scree_plt_LC

# ggsave(plot = scree_plt_LC, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_LC_25.png",sep=""),width=3,height=2.5,dpi=300)


biplot(prcomp(LC_pca_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_LC <- fviz_pca_biplot(pca_LC, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories)

watershed_loadplot_LC
# ggsave(plot = watershed_loadplot_LC, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_LC_25.png",sep=""),width=5,height=3.5,dpi=300)


```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)
LC_pca_scores <- as.data.frame(pca_LC$x)

LC_pca_input_clean_sites <- na.omit(LC_pca_data1)

# Add sites
LC_pca_with_site <- cbind(Site_ID=LC_pca_input_clean_sites$Site_ID, LC_pca_scores)

# Add GPP_synch back in (at the annual scale)
LC_pca_with_response <- river_metab_dat %>%
  left_join(LC_pca_with_site, by=c("Site_ID"))

```

### A. How much variance in GPP is explained by landscape? 

GLM: 
- PC1: slope = +0.271
- PC2: slope = +0.298

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_LC <- lm(GPP_filled ~ PC1 + PC2, data = LC_pca_with_response)
summary(lm_model_LC)

hist(residuals(lm_model_LC)) # meh
```

#### BRMs 
- PC1: slope + 0.27
- PC3: slope + 0.30

Very similar to glm 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
library(brms)

# Bayesian linear model
brm_model <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = LC_pca_with_response,
  family = gaussian(),  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)


summary(brm_model)
plot(brm_model)  # Diagnostic plots

# bayes_R2(brm_model)

```

##### So it seems like daily GPP is positively associated with *PC1* and *PC2*

PC1: (-) open space to crops (+) 
PC2: (-) forest space to open space (+) 

-  Postive values of PC 1 correspond to land cover as *crops* "Cultivated_Crops_pct.Cultivated_Crops_pct"
"Grass_PastureHay_pct.Grass_PastureHay_pct" as opposed to *open space* "Developed_OpenSpace_pct.Developed_OpenSpace_pct" or "Barren_Land_pct.Barren_Land_pct" 

- Positive values of PC2 represent forested areas. 


#### Quantile of different land cover type: 

```{r, size='large', echo=F, include=F, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
df_LC <- LC_pca_with_response %>%
  mutate(PC1_quantile = ntile(PC1, 3)) %>%
  mutate(PC1_category = case_when(
    PC1_quantile == 1 ~ "Lower",
    PC1_quantile == 2 ~ "Middle",
    PC1_quantile == 3 ~ "Upper"
  ))

```


LC is fairly different across the 3 different groups 


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Lower" = "#00BFC4",   
  "Middle" = "#7CAE00",
  "Upper" = "#F8766D"
)
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "LC_Q1" = "#00BFC4",   
  "LC_Q2" = "#7CAE00",
  "LC_Q3" = "#F8766D"
)
```

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_LC, aes(x = PC1, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by LC quantile",
       x = "PC1",
       y = "Count")
```


However daily GPP looks fairly similar across all land cover types - despite the postive regression resuls with cropland 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(df_LC, aes(x = GPP_filled, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "GPP by LC quantile",
       x = "GPP",
       y = "Count")
```

##### i. Land cover Q1 
All open space or forested 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q1 <- df_LC %>%
dplyr::filter(PC1_category=="Lower") 

unique(land_cover_q1$Site_ID)
unique(land_cover_q1$Latitude)

hist(land_cover_q1$GPP_filled,  breaks = 30)

```

##### ii. Land cover Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q2 <- df_LC %>%
dplyr::filter(PC1_category=="Middle") 


unique(land_cover_q2$Site_ID)
unique(land_cover_q2$Latitude)

hist(land_cover_q2$GPP_filled)

```

##### iii. Land cover Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q3 <- df_LC %>%
dplyr::filter(PC1_quantile==3)
unique(land_cover_q3$Site_ID)
unique(land_cover_q3$Latitude)

hist(land_cover_q2$GPP_filled)

```



#### Daily synchrony land cover Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- land_cover_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_LC_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_LC_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_LC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_LC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "LC_Q1")

## Sample average line
GPP.corr.pred.LC.Q1$mean_synch <- mean(GPP.corr.pred.LC.Q1$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q1 <- as.data.frame(t(RR_df.corr_LC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "LC_Q1")

```

#### Daily synchrony land cover Q2


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_Q2 <- land_cover_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_Q2 <- GPP_long_Q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_LC_Q2 <- Sncf(x=GPP_wide_day_Q2$Lon,
                   y=GPP_wide_day_Q2$Lat,
                   z=GPP_wide_day_Q2[,4:ncol(GPP_wide_day_Q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_LC_Q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_LC_Q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_LC_Q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_LC_Q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "LC_Q2")

## Sample average line
GPP.corr.pred.LC.Q2$mean_synch <- mean(GPP.corr.pred.LC.Q2$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q2 <- as.data.frame(t(RR_df.corr_LC_Q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_LC_Q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "LC_Q2")

```




#### Daily synchrony land cover Q3


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_Q3 <- land_cover_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_Q3 <- GPP_long_Q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_LC_Q3 <- Sncf(x=GPP_wide_day_Q3$Lon,
                   y=GPP_wide_day_Q3$Lat,
                   z=GPP_wide_day_Q3[,4:ncol(GPP_wide_day_Q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_LC_Q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_LC_Q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_LC_Q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_LC_Q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "LC_Q3")

## Sample average line
GPP.corr.pred.LC.Q3$mean_synch <- mean(GPP.corr.pred.LC.Q3$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q3 <- as.data.frame(t(RR_df.corr_LC_Q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_LC_Q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "LC_Q3")
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all <- bind_rows(GPP.corr.pred.LC.Q1, GPP.corr.pred.LC.Q2, GPP.corr.pred.LC.Q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)


ci_df_all   <- bind_rows(GPP.corr.boot.LC.Q1, GPP.corr.boot.LC.Q2, GPP.corr.boot.LC.Q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)

```

##### Calculate the percent difference between different land cover types 

```{r, echo=T, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    LQ1_diff = (abs(LC_Q1 - LC_Q1) / LC_Q1) * 100,
    LQ2_diff = (abs(LC_Q1 - LC_Q2) / LC_Q1) * 100,
    LQ3_diff = (abs(LC_Q1 - LC_Q3) / LC_Q1) * 100)

pred_diff
```


### Figure 4a (?): Spline Correlogram of GPP at different landcover types


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_LC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%filter(x<3250),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%filter(x<3250),
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all%>%filter(x<3100),
             aes(yintercept = mean_synch, color = strat),
             linetype = "dashed") +
  labs(title = "Synchrony in GPP across different land cover types",
       y = "Correlation",
       x = "Distance (km)",
       fill = "LC Type",
       color = "LC Type") +
  scale_color_manual(values = strat_colors) +
  scale_fill_manual(values = strat_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_LC

```


Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
max_synch <- pred_df_all %>%
  group_by(strat) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 2))

max_synch
```

Agricultural and forest influences sites may have the most synchronous GPP.

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}

fit_models <- pred_df_all %>%
  group_by(strat) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```


#### Sediment/ geologic land cover 

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}
library(randomForest)

# Prepare data for RF
rf_geol_data <- river_metab_dat1[, c("geol_gneiss", "geol_granitic",
                       "geol_ultramafic", "geol_quarternary",
                       "geol_sedimentary", "geol_volcanic",
                       "geol_ultramafic", "geol_anorthositic",
                       "geol_intermediate_pultonic")]


rf_geol_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_geol_data)
rf_geol_input_clean <- na.omit(rf_geol_input)

# Now run the model
rf_geol_model <- randomForest(GPP_synch ~ ., data = rf_geol_input_clean,
                         importance = TRUE, ntree = 200)

# Plot importance
varImpPlot(rf_geol_model)

# Get importance metrics from your existing model
importance_vals <- importance(rf_geol_model, type = 1)  # type = 1 for %IncMSE
importance_df <- data.frame(variable = rownames(importance_vals),
                            IncMSE = importance_vals[, 1])

# Sort by importance and keep top 10
top_vars_geol <- importance_df %>%
  arrange(desc(IncMSE)) %>%
  slice(1:10) %>%
  pull(variable)

# 
# rf_geol_input_reduced <- rf_geol_input_clean %>%
#   dplyr::select(GPP_synch, all_of(top_vars_geol))
# 
# rf_geol_model_reduced <- randomForest(GPP_synch ~ ., data = rf_geol_input_reduced,
#                                  importance = TRUE, ntree = 500)
# 
# varImpPlot(rf_geol_model_reduced)
# 
# # Sort by importance and keep top 10
# top_vars_geol <- importance_df %>%
#   arrange(desc(IncMSE)) %>%
#   slice(1:5) %>%
#   pull(variable)
# 
# top_vars_geol

```


#### Create the composite variable for geology

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
### start the PCA 
pca_geol_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID, geol_gneiss, geol_granitic, geol_sedimentary, geol_volcanic, geol_ultramafic, geol_quarternary, geol_anorthositic)

pca_geol_data1 <- pca_geol_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
pca_geol_data <- pca_geol_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

pca_geol_input_clean <- na.omit(pca_geol_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(pca_geol_data), function(col) {
  col_patterns[col]
}))


pca_geol <- prcomp(pca_geol_input_clean, center = TRUE, scale. = TRUE)

# Visualize PCA eigenvalues
scree_plt_geol <- fviz_eig(pca_geol)
scree_plt_geol

# ggsave(plot = scree_plt_geol, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_geol_25.png",sep=""),width=4,height=3.5,dpi=300)


biplot(prcomp(pca_geol_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_geol<- fviz_pca_biplot(pca_geol, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories)

watershed_loadplot_geol

# ggsave(plot = watershed_loadplot_geol, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_geol_25.png",sep=""),width=5,height=3.5,dpi=300)

```


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)

Geol_pca_scores <- as.data.frame(pca_geol$x)

Geol_pca_input_clean_sites <- na.omit(pca_geol_data1)

# Add sites
Geol_pca_with_site <- cbind(Site_ID=Geol_pca_input_clean_sites$Site_ID, Geol_pca_scores)

# Add GPP_synch back in (at the annual scale)
Geol_pca_with_response <- river_metab_dat %>%
  left_join(Geol_pca_with_site, by=c("Site_ID"))

```

### B. How much variance in GPP is explained by geology?

GLM: 
- PC1: slope = +0.271
- PC2: slope = +0.298

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_geol <- lm(GPP_filled ~ PC1 + PC2, data = Geol_pca_with_response)
summary(lm_model_geol)

hist(residuals(lm_model_geol)) # meh
```

#### BRMs slope + 0.25

Very similar to glm again 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
# Bayesian linear model
brm_model_geol <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = Geol_pca_with_response,
  family = gaussian(),  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)

summary(brm_model_geol)
plot(brm_model_geol)  # Diagnostic plots

# bayes_R2(brm_model)

```

Positive slope PC1 represents increased GPP with metamorphic or geology "geol_gneiss.geol_gneiss" or "geol_ultramafic.geol_ultramafic"
PC2 Positive slope represents more geology_sedimentary


#### C. Quantile of different geologies: 

```{r, size='large', echo=F, include=F, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
df_geol <- Geol_pca_with_response %>%
  filter(!is.na(PC1))%>%
  mutate(PC1_quantile = ntile(PC1, 3)) %>%
  mutate(PC1_category = case_when(
    PC1_quantile == 1 ~ "Lower",
    PC1_quantile == 2 ~ "Middle",
    PC1_quantile == 3 ~ "Upper"
  ))

```


Geology groups look relatively more similar


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Lower" = "#00BFC4",   
  "Middle" = "#7CAE00",
  "Upper" = "#F8766D"
)
```

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_geol, aes(x = PC1, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by geology quantile",
       x = "PC1",
       y = "Count")
```


However daily GPP looks fairly similar across all land cover types - despite the postive regression resuls with cropland 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(df_geol, aes(x = GPP_filled, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "GPP by geology quantile",
       x = "GPP",
       y = "Count")
```

##### i. Geology Q1 
All open space or forested 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

geol_q1 <- df_geol %>%
dplyr::filter(PC1_category=="Lower") 

unique(geol_q1$Site_ID)
unique(geol_q1$Latitude)

hist(geol_q1$GPP_filled,  breaks = 30)

```

##### ii. Geology Q2

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

geol_q2 <- df_geol %>%
dplyr::filter(PC1_category=="Middle") 


unique(geol_q2$Site_ID)
unique(geol_q2$Latitude)

hist(geol_q2$GPP_filled)

```

##### iii. Geology Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

geol_q3 <- df_geol %>%
dplyr::filter(PC1_quantile==3)
unique(geol_q3$Site_ID)
unique(geol_q3$Latitude)

hist(geol_q2$GPP_filled)

```



#### Daily synchrony Geology Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- geol_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_GL_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_GL_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_GL_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.GL.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_GL_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "GL_Q1")

## Sample average line
GPP.corr.pred.GL.Q1$mean_synch <- mean(GPP.corr.pred.GL.Q1$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.GL.Q1 <- as.data.frame(t(RR_df.corr_GL_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.GL.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.GL.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.GL.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "GL_Q1")

```

#### Daily synchrony Geology Q2


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- geol_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_GL_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_GL_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_GL_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.GL.q2 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_GL_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "GL_Q2")

## Sample average line
GPP.corr.pred.GL.q2$mean_synch <- mean(GPP.corr.pred.GL.q2$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.GL.q2 <- as.data.frame(t(RR_df.corr_GL_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.GL.q2 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.GL.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.GL.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "GL_Q2")
```



#### Daily synchrony Geology Q3


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- geol_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_GL_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_GL_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_GL_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.GL.q3 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_GL_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "GL_Q3")

## Sample average line
GPP.corr.pred.GL.q3$mean_synch <- mean(GPP.corr.pred.GL.q3$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.GL.q3 <- as.data.frame(t(RR_df.corr_GL_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.GL.q3 <- data.frame(x = matrix(unlist(RR_df.corr_GL_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.GL.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.GL.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "GL_Q3")

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_geol <- bind_rows(GPP.corr.pred.GL.Q1, GPP.corr.pred.GL.q2, GPP.corr.pred.GL.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)


ci_df_all_geol   <- bind_rows(GPP.corr.boot.GL.Q1, GPP.corr.boot.GL.q2, GPP.corr.boot.GL.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)

```

##### Calculate the percent difference between different Geology types 

```{r, echo=T, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_geol <- pred_df_all_geol %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_geol <- pred_wide_geol %>%
  mutate(
    LQ1_diff = (abs(GL_Q1 - GL_Q1) / GL_Q1) * 100,
    LQ2_diff = (abs(GL_Q1 - GL_Q2) / GL_Q1) * 100,
    LQ3_diff = (abs(GL_Q1 - GL_Q3) / GL_Q1) * 100)

pred_diff_geol
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "GL_Q1" = "#00BFC4",   
  "GL_Q2" = "#7CAE00",
  "GL_Q3" = "#F8766D"
)
```


### Figure 4b (?): Spline Correlogram of GPP at different geology types


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_GL <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_geol,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_geol,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all_geol,
             aes(yintercept = mean_synch, color = strat),
             linetype = "dashed") +
  labs(title = "Synchrony in GPP across different geology types",
       y = "Correlation",
       x = "Distance (km)",
       fill = "GL Type",
       color = "GL Type") +
  scale_color_manual(values = strat_colors) +
  scale_fill_manual(values = strat_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_GL

```

Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
max_synch <- pred_df_all_geol %>%
  group_by(strat) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

max_synch
```

Catchments with underlying metamorphic rock are more similar than igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}

fit_models <- pred_df_all_geol %>%
  group_by(strat) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```



### C. Basin characteristics  

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}

rf_geol_data <- river_metab_dat1[, c("geol_gneiss", "geol_granitic",
                       "geol_ultramafic", "geol_quarternary",
                       "geol_sedimentary", "geol_volcanic",
                       "geol_ultramafic", "geol_anorthositic",
                       "geol_intermediate_pultonic")]

# Prepare data for RF
rf_basin_data <- river_metab_dat1[,c("lengthkm", "shape_length", "streamleve", 
                                     "streamorde", "streamcalc", "totdasqkm", "slope",
                                     "pathlength")]


rf_basin_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_basin_data)
rf_basin_input_clean <- na.omit(rf_basin_input)

# Now run the model
rf_basin_model <- randomForest(GPP_synch ~ ., data = rf_basin_input_clean,
                         importance = TRUE, ntree = 200)

# Plot importance
varImpPlot(rf_basin_model)

# Get importance metrics from your existing model
importance_vals <- importance(rf_basin_model, type = 1)  # type = 1 for %IncMSE
importance_df <- data.frame(variable = rownames(importance_vals),
                            IncMSE = importance_vals[, 1])

# Sort by importance and keep top 10
top_vars_basin <- importance_df %>%
  arrange(desc(IncMSE)) %>%
  slice(1:10) %>%
  pull(variable)
```


#### Create the composite variable for basin characteristics

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
### start the PCA 
pca_basin_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID,  all_of(top_vars_basin))

pca_basin_data1 <- pca_basin_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
pca_basin_data <- pca_basin_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

pca_basin_input_clean <- na.omit(pca_basin_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(pca_basin_data), function(col) {
  col_patterns[col]
}))


pca_basin <- prcomp(pca_basin_input_clean, center = TRUE, scale. = TRUE)

# Visualize PCA eigenvalues
scree_plt_basin <- fviz_eig(pca_basin)
scree_plt_basin

# ggsave(plot = scree_plt_basin, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_basin_25.png",sep=""),width=4,height=3.5,dpi=300)


biplot(prcomp(pca_basin_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_basin<- fviz_pca_biplot(pca_basin, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories)

watershed_loadplot_basin

# ggsave(plot = watershed_loadplot_basin, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_basin_25.png",sep=""),width=5,height=3.5,dpi=300)

```

*PC1:*

- low PC1 score site is one with large, steep, high-elevation basins.

- A high PC1 score site is one with smaller, flatter, low-elevation basins

*PC2:*

- Low PC2 score = Large, long basins with large drainage networks

- High PC2 score = Smaller, steeper, higher-elevation basins



```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)

basin_pca_scores <- as.data.frame(pca_basin$x)

basin_pca_input_clean_sites <- na.omit(pca_basin_data1)

# Add sites
basin_pca_with_site <- cbind(Site_ID=basin_pca_input_clean_sites$Site_ID, basin_pca_scores)

# Add GPP_synch back in (at the annual scale)
basin_pca_with_response <- river_metab_dat %>%
  left_join(basin_pca_with_site, by=c("Site_ID"))

```

#### B. How much variance in GPP is explained by geology?

GLM: 
- PC1: slope = -0.223
- PC2: slope = +0.298

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_basin <- lm(GPP_filled ~ PC1 + PC2, data = basin_pca_with_response)
summary(lm_model_basin)

hist(residuals(lm_model_basin)) # meh
```

Slope PC1: Showing lower GPP in large, steep, high-elevation basins


#### BRMs slope + 0.25

Very similar to glm again 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
# Bayesian linear model
brm_model_basin <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = basin_pca_with_response,
  family = gaussian(),  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)

summary(brm_model_basin)
plot(brm_model_basin)  # Diagnostic plots

# bayes_R2(brm_model)

```




#### C. Quantile of different basinogies: 

```{r, size='large', echo=F, include=F, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
df_basin <- basin_pca_with_response %>%
  mutate(PC1_quantile = ntile(PC1, 3)) %>%
  mutate(PC1_category = case_when(
    PC1_quantile == 1 ~ "Lower",
    PC1_quantile == 2 ~ "Middle",
    PC1_quantile == 3 ~ "Upper"
  ))

```


basinogy groups look relatively more similar


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Lower" = "#00BFC4",   
  "Middle" = "#7CAE00",
  "Upper" = "#F8766D"
)
```

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_basin, aes(x = PC1, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by character quantile",
       x = "PC1",
       y = "Count")
```


However daily GPP looks fairly similar across all land cover types - despite the postive regression resuls with cropland 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(df_basin, aes(x = GPP_filled, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "GPP by basin character quantile",
       x = "GPP",
       y = "Count")
```

##### i. basinogy Q1 
All open space or forested 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q1 <- df_basin %>%
dplyr::filter(PC1_category=="Lower") 

unique(basin_q1$Site_ID)
unique(basin_q1$Latitude)

hist(basin_q1$GPP_filled,  breaks = 30)

```

##### ii. basinogy Q2

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q2 <- df_basin %>%
dplyr::filter(PC1_category=="Middle") 


unique(basin_q2$Site_ID)
unique(basin_q2$Latitude)

hist(basin_q2$GPP_filled)

```

##### iii. basinogy Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q3 <- df_basin %>%
dplyr::filter(PC1_quantile==3)
unique(basin_q3$Site_ID)
unique(basin_q3$Latitude)

hist(basin_q2$GPP_filled)

```




#### Daily synchrony basin Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- basin_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "BC_Q1")

## Sample average line
GPP.corr.pred.BC.Q1$mean_synch <- mean(GPP.corr.pred.BC.Q1$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.Q1 <- as.data.frame(t(RR_df.corr_BC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "BC_Q1")

```

#### Daily synchrony basinogy Q2


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- basin_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "BC_Q2")

## Sample average line
GPP.corr.pred.BC.q2$mean_synch <- mean(GPP.corr.pred.BC.q2$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.q2 <- as.data.frame(t(RR_df.corr_BC_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "BC_Q2")
```



#### Daily synchrony basinogy Q3


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- basin_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "BC_Q3")

## Sample average line
GPP.corr.pred.BC.q3$mean_synch <- mean(GPP.corr.pred.BC.q3$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.q3 <- as.data.frame(t(RR_df.corr_BC_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "BC_Q3")

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_basin <- bind_rows(GPP.corr.pred.BC.Q1, GPP.corr.pred.BC.q2, GPP.corr.pred.BC.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)


ci_df_all_basin   <- bind_rows(GPP.corr.boot.BC.Q1, GPP.corr.boot.BC.q2, GPP.corr.boot.BC.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<1200)

```

##### Calculate the percent difference between different basinogy types 

```{r, echo=T, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_basin <- pred_df_all_basin %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_basin <- pred_wide_basin %>%
  mutate(
    BC_Q1_diff = (abs(BC_Q1 - BC_Q1) / BC_Q1) * 100,
    BC_Q2_diff = (abs(BC_Q1 - BC_Q2) / BC_Q1) * 100,
    BC_Q3_diff = (abs(BC_Q1 - BC_Q3) / BC_Q1) * 100)

pred_diff_basin
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "BC_Q1" = "#00BFC4",   
  "BC_Q2" = "#7CAE00",
  "BC_Q3" = "#F8766D"
)
```


### Figure 4c (?): Spline Correlogram of GPP at different basin characteristics


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_BC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_basin,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_basin,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all_basin,
             aes(yintercept = mean_synch, color = strat),
             linetype = "dashed") +
  labs(title = "Synchrony in GPP across different basin characteristics",
       y = "Correlation",
       x = "Distance (km)",
       fill = "BC Type",
       color = "BC Type") +
  scale_color_manual(values = strat_colors) +
  scale_fill_manual(values = strat_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_BC

```

Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
max_synch <- pred_df_all_basin %>%
  group_by(strat) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

max_synch
```

Catchments with underlying metamorphic rock are more similar than igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate:
higher b = faster loss of synchrony with distance.

```{r, warning=F, size = 'small', echo=T, include=T}

fit_models <- pred_df_all_basin %>%
  group_by(strat) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```




###

### D. streamflow characteristics  

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}
# library(randomForest)
# 
# # Prepare data for RF
# rf_flow_data <- river_metab_dat1[, c("discharge", "depth",
#                        "temp.water")]
# 
# 
# rf_flow_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_flow_data)
# rf_flow_input_clean <- na.omit(rf_flow_input)
# 
# # Now run the model
# rf_flow_model <- randomForest(GPP_synch ~ ., data = rf_flow_input_clean,
#                          importance = TRUE, ntree = 500)
# 
# # Plot importance
# varImpPlot(rf_flow_model)
# 
# # Get importance metrics from your existing model
# importance_vals <- importance(rf_flow_model, type = 1)  # type = 1 for %IncMSE
# importance_df <- data.frame(variable = rownames(importance_vals),
#                             IncMSE = importance_vals[, 1])
# 
# # Sort by importance and keep top 10
# top_vars_flow <- importance_df %>%
#   arrange(desc(IncMSE)) %>%
#   slice(1:10) %>%
#   pull(variable)
```


#### Create the composite variable for flow characteristics

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
### start the PCA 
pca_flow_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID, discharge, depth, temp.water, DO.obs, Tree_Canopyin100mRip)

pca_flow_data1 <- pca_flow_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
pca_flow_data <- pca_flow_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

pca_flow_input_clean <- na.omit(pca_flow_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(pca_flow_data), function(col) {
  col_patterns[col]
}))


pca_flow <- prcomp(pca_flow_input_clean, center = TRUE, scale. = TRUE)

# Visualize PCA eigenvalues
scree_plt_flow <- fviz_eig(pca_flow)
scree_plt_flow

# ggsave(plot = scree_plt_flow, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_flow_25.png",sep=""),width=4,height=3.5,dpi=300)


biplot(prcomp(pca_flow_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_flow<- fviz_pca_biplot(pca_flow, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories)

watershed_loadplot_flow

# ggsave(plot = watershed_loadplot_flow, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_flow_25.png",sep=""),width=5,height=3.5,dpi=300)

```

*PC1:*

- Higher: Larger, deeper, colder more oxygenated streams 

- Lower: smaller, warmer, more canopy-covered streams



```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)

flow_pca_scores <- as.data.frame(pca_flow$x)

flow_pca_input_clean_sites <- na.omit(pca_flow_data1)

# Add sites
flow_pca_with_site <- cbind(Site_ID=flow_pca_input_clean_sites$Site_ID, flow_pca_scores)

# Add GPP_synch back in (at the annual scale)
flow_pca_with_response <- river_metab_dat %>%
  left_join(flow_pca_with_site, by=c("Site_ID"))

```

#### B. How much variance in GPP is explained by geology?

GLM: 
- PC1: slope = 0.69

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_flow <- lm(GPP_filled ~ PC1 + PC2, data = flow_pca_with_response)
summary(lm_model_flow)

hist(residuals(lm_model_flow)) # meh
```

Slope PC1: Showing lower GPP in large, steep, high-elevation flows


#### BRMs slope + 0.25

Very similar to glm again 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
# Bayesian linear model
brm_model_flow <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = flow_pca_with_response,
  family = gaussian(),  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)

summary(brm_model_flow)
plot(brm_model_flow)  # Diagnostic plots

# bayes_R2(brm_model)

```




#### C. Quantile of different flow: 

```{r, size='large', echo=F, include=F, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
df_flow <- flow_pca_with_response %>%
  mutate(PC1_quantile = ntile(PC1, 3)) %>%
  mutate(PC1_category = case_when(
    PC1_quantile == 1 ~ "Lower",
    PC1_quantile == 2 ~ "Middle",
    PC1_quantile == 3 ~ "Upper"
  ))

```


flowogy groups look relatively more similar


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Lower" = "#00BFC4",   
  "Middle" = "#7CAE00",
  "Upper" = "#F8766D"
)
```

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_flow, aes(x = PC1, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by flow quantile",
       x = "PC1",
       y = "Count")
```


However daily GPP looks fairly similar across all land cover types - despite the postive regression resuls with cropland 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(df_flow, aes(x = GPP_filled, fill = factor(PC1_category))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Lower", "Middle", "Upper")
  ) +
  theme_minimal() +
  labs(title = "GPP by flow character quantile",
       x = "GPP",
       y = "Count")
```

##### i. flow Q1 
All open space or forested 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

flow_q1 <- df_flow %>%
dplyr::filter(PC1_category=="Lower") 

unique(flow_q1$Site_ID)
unique(flow_q1$Latitude)

hist(flow_q1$GPP_filled,  breaks = 30)

```

##### ii. flow Q2

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

flow_q2 <- df_flow %>%
dplyr::filter(PC1_category=="Middle") 


unique(flow_q2$Site_ID)
unique(flow_q2$Latitude)

hist(flow_q2$GPP_filled)

```

##### iii. flow Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

flow_q3 <- df_flow %>%
dplyr::filter(PC1_quantile==3)
unique(flow_q3$Site_ID)
unique(flow_q3$Latitude)

hist(flow_q2$GPP_filled)

```




#### Daily synchrony flow Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- flow_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q1")

## Sample average line
GPP.corr.pred.flow.Q1$mean_synch <- mean(GPP.corr.pred.flow.Q1$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.Q1 <- as.data.frame(t(RR_df.corr_flow_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q1")

```

#### Daily synchrony flowogy Q2


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- flow_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q2")

## Sample average line
GPP.corr.pred.flow.q2$mean_synch <- mean(GPP.corr.pred.flow.q2$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.q2 <- as.data.frame(t(RR_df.corr_flow_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q2")
```



#### Daily synchrony flow Q3


```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- flow_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q3")

## Sample average line
GPP.corr.pred.flow.q3$mean_synch <- mean(GPP.corr.pred.flow.q3$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.q3 <- as.data.frame(t(RR_df.corr_flow_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q3")

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_flow <- bind_rows(GPP.corr.pred.flow.Q1, GPP.corr.pred.flow.q2, GPP.corr.pred.flow.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<1200)


ci_df_all_flow   <- bind_rows(GPP.corr.boot.flow.Q1, GPP.corr.boot.flow.q2, GPP.corr.boot.flow.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<1200)

```

##### Calculate the percent difference between different flowogy types 

```{r, echo=T, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_flow <- pred_df_all_flow %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_flow <- pred_wide_flow %>%
  mutate(
    flow_Q1_diff = (abs(flow_Q1 - flow_Q1) / flow_Q1) * 100,
    flow_Q2_diff = (abs(flow_Q1 - flow_Q2) / flow_Q1) * 100,
    flow_Q3_diff = (abs(flow_Q1 - flow_Q3) / flow_Q1) * 100)

pred_diff_flow
```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "flow_Q1" = "#00BFC4",   
  "flow_Q2" = "#7CAE00",
  "flow_Q3" = "#F8766D"
)
```


### Figure 4d (?): Spline Correlogram of GPP at different flow characteristics


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_flow <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_flow,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_flow,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all_flow,
             aes(yintercept = mean_synch, color = strat),
             linetype = "dashed") +
  labs(title = "Synchrony in GPP across different flow",
       y = "Correlation",
       x = "Distance (km)",
       fill = "BC Type",
       color = "BC Type") +
  scale_color_manual(values = strat_colors) +
  scale_fill_manual(values = strat_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_flow

```

Max synch? 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
max_synch <- pred_df_all_flow %>%
  group_by(strat) %>%
  summarise(
    max_synch = round((max(y, na.rm=T)), 3))

max_synch
```




```{r, warning=F, size = 'small', echo=F,include=TRUE, fig.width=12, fig.height=7}
# Arrange plots with shared legend at the bottom
library(ggpubr)

spline_grid <- ggarrange(
  GPP_synch_spline_LC,
  GPP_synch_spline_GL,
GPP_synch_spline_BC,
#GPP_synch_spline_flow,
  ncol = 3, nrow = 1,
  common.legend = F, 
  legend = "bottom"
)

spline_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Sup_figure_EGRET.png", plot = KD_coeff_grid, width = 12, height = 10, units = "in")

```


## might be nicer to know the individual influence of just flow, temp, stream light, in GPP and ER synch 




## To DOS:
- LC/Q labels to great - for ecological context 
-  RUN individual quantiles of Temp, Flow, and Light on GPP and ER
