---
title: "Spline Correlogram analysis for synchrony in aquatic ecosystem productivity"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```


```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

river_metab_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/stream_spline_cor_df.rds")

```

### Daily synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
river_metab_dat<-as.data.frame(river_metab_dat)

GPP_long <- river_metab_dat_filled %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day <- GPP_long %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_day <- Sncf(x=GPP_wide_day$Lon,
                   y=GPP_wide_day$Lat,
                   z=GPP_wide_day[,4:ncol(GPP_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(RR_df.corr_day)


GPP_corrgram_summary.table <- summary(RR_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_day$real$predicted$y)),
                               time_scale = "daily")

## Sample average line
GPP.corr.pred.df.day$mean_synch <- mean(GPP.corr.pred.df.day$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "daily")

```


## Weekly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}


GPP_week <-  river_metab_dat %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))

GPP_wide_week <- GPP_week %>%
  dplyr::select(Site_ID, Lon, Lat, week, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }

GPP_wide_week_1 <-na.omit(GPP_wide_week)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=GPP_wide_week$Lon,
                   y=GPP_wide_week$Lat,
                   z=GPP_wide_week[,4:ncol(GPP_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)


GPP_corrgram_summary.table <- summary(RR_df.corr_wk)
GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly")

## Sample average line
GPP.corr.pred.df.wk$mean_synch <- mean(GPP.corr.pred.df.wk$y, na.rm = TRUE)


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly")
```

### Correlogam of annual GPP synchrony

## Monthly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

GPP_month <-  river_metab_dat %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))

river_metab_dat_filled<- as.data.frame(river_metab_dat_filled)

GPP_wide_month <- GPP_month %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=GPP_wide_month$Lon,
                   y=GPP_wide_month$Lat,
                   z=GPP_wide_month[,4:ncol(GPP_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


GPP_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly")

## Sample average line
GPP.corr.pred.df.mon$mean_synch <- mean(GPP.corr.pred.df.mon$y, na.rm = TRUE)

GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                time_scale = "monthly")
```


## year synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

GPP_yr <- river_metab_dat %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_yr <- GPP_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, GPP_filled) %>%
  pivot_wider(names_from = c(year), values_from = GPP_filled)


str(GPP_wide_qrt)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=GPP_wide_yr$Lon,
                   y=GPP_wide_yr$Lat,
                   z=GPP_wide_yr[,4:ncol(GPP_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

GPP_corrgram_summary.table <- summary(RR_df.corr_yr)
GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly")


## Sample average line
GPP.corr.pred.df.yr$mean_synch <- mean(GPP.corr.pred.df.yr$y, na.rm = TRUE)


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly")

```




## Quarterly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

GPP_qrt <- river_metab_dat %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_qrt <- GPP_qrt %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=GPP_wide_qrt$Lon,
                   y=GPP_wide_qrt$Lat,
                   z=GPP_wide_qrt[,4:ncol(GPP_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)

GPP_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly")
## Sample average line
GPP.corr.pred.df.qrtly$mean_synch <- mean(GPP.corr.pred.df.qrtly$y, na.rm = TRUE)

GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly")


```


### combined synch 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all <- bind_rows(GPP.corr.pred.df.day, GPP.corr.pred.df.wk, GPP.corr.pred.df.mon, GPP.corr.pred.df.qrtly, GPP.corr.pred.df.yr)
ci_df_all   <- bind_rows(GPP.corr.boot.df.day, GPP.corr.boot.df.wk, GPP.corr.boot.df.mon, GPP.corr.boot.df.qrtly, GPP.corr.boot.df.yr)

```

### calculate the percent difference between time steps

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

# Step 1: Spread the data so each time_scale becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x, y, time_scale) %>%
  pivot_wider(names_from = time_scale, values_from = y)

# Step 2: Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )

```


### Figure of precent difference in snych signal 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}

# Step 1: Pivot longer
pred_long_diff <- pred_diff %>%
  dplyr::select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )

# Step 2: Set a color-blind friendly palette
scale_colors <- c(
  "daily" = "#E69F00",     # orange
  "quarterly" = "#56B4E9",    # sky blue
  "monthly" = "#009E73",   # bluish green
  "weekly" = "#F0E442", # yellow
  "yearly" = "#0072B2"     # blue
)

# Step 3: Plot histogram
ggplot(pred_long_diff, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```

###

Add points for each stream to get a sense of geographic distance of predicted values.


```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)

### add point for mean GPP to plot 
library(geosphere)
# Get site centroids (optional, or pick one reference site)
site_coords <- GPP_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (e.g., the upstream-most one)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

GPP_long_dist <- GPP_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")


GPP_long_dist_plt <- GPP_long_dist %>%
  group_by(Site_ID, distance_km) %>%
  summarise(
    GPP_mean = median(GPP_filled, na.rm = TRUE), 
    GPP_low = quantile(GPP_filled, probs = 0.25, na.rm = TRUE), 
    GPP_up = quantile(GPP_filled, probs = 0.75, na.rm = TRUE),
    .groups = "drop"
  )

###
GPP_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%filter(x<3250),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%filter(x<3250),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all%>%filter(x<3100),
             aes(yintercept = mean_synch, color = time_scale),
             linetype = "dashed") +
  # Add GPP points (mean)
  geom_point(data = GPP_long_dist_plt,
             aes(x = distance_km, y = GPP_mean/25),
             shape = 1, size = 2,
             alpha=0.5) +
    ylim(-0.2, 0.75) +
  # Add GPP error bars
  geom_errorbar(data = GPP_long_dist_plt,
                aes(x = distance_km, ymin = GPP_low/25, ymax = GPP_up/25),
                width = 2, color = "black", 
                alpha=0.25) +
  # Labels and theme
  labs(title = "Synchrony in GPP across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff%>%filter(x<3250), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(5, 5, 5, 5),
        legend.position="none")

# Combine using patchwork (put rotated density plot on the right)
combined_plot <- GPP_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```


```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/GPP_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```



#### GPP: Look when synchrony dips below the sample average

This gives an idea for at what distances stream GPP is no longer synchronous 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df_all %>%
  mutate(
    mean_synch = round(mean_synch, 3),
    y = round(y, 3),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(time_scale) %>%
  # Filter only rows where y is approximately equal to mean_synch
  filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```


### ER synchrony


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

river_metab_dat<-as.data.frame(river_metab_dat)

ER_long <- river_metab_dat_filled %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day <- ER_long %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
ER_df.corr_day <- Sncf(x=ER_wide_day$Lon,
                   y=ER_wide_day$Lat,
                   z=ER_wide_day[,4:ncol(ER_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(ER_df.corr_day)


ER_corrgram_summary.table <- summary(ER_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(ER_df.corr_day$real$predicted$y)),
                               time_scale = "daily")


## Sample average line
ER.corr.pred.df.day$mean_synch <- mean(ER.corr.pred.df.day$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(ER_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "daily")


```



## Weekly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}


ER_week <-  river_metab_dat %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

ER_wide_week <- ER_week %>%
  dplyr::select(Site_ID,Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled)

river_metab_dat_filled<- as.data.frame(river_metab_dat_filled)

ER_wide_week <- ER_week %>%
 # filter(year>2011) %>%
 dplyr::select(Site_ID, Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=ER_wide_week$Lon,
                   y=ER_wide_week$Lat,
                   z=ER_wide_week[,4:ncol(ER_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)



ER_corrgram_summary.table <- summary(RR_df.corr_wk)
ER_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly")
## Sample average line
ER.corr.pred.df.wk$mean_synch <- mean(ER.corr.pred.df.wk$y, na.rm = TRUE)

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly")
```





## Monthly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

ER_month <-  river_metab_dat %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

river_metab_dat_filled<- as.data.frame(river_metab_dat_filled)

ER_wide_month <- ER_month %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, ER_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=ER_wide_month$Lon,
                   y=ER_wide_month$Lat,
                   z=ER_wide_month[,4:ncol(ER_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


ER_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly")

## Sample average line
ER.corr.pred.df.mon$mean_synch <- mean(ER.corr.pred.df.mon$y, na.rm = TRUE)

ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                time_scale = "monthly")

```



## Quarterly synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

ER_qrt <- river_metab_dat %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


river_metab_dat_filled<- as.data.frame(river_metab_dat_filled)

ER_wide_qrt <- ER_qrt %>%
 # filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, ER_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=ER_wide_qrt$Lon,
                   y=ER_wide_qrt$Lat,
                   z=ER_wide_qrt[,4:ncol(ER_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)


ER_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly")

## Sample average line
ER.corr.pred.df.qrtly$mean_synch <- mean(ER.corr.pred.df.qrtly$y, na.rm = TRUE)


ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly")


```


## year synch

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

ER_yr <- river_metab_dat %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


ER_wide_yr <- ER_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, ER_filled) %>%
  pivot_wider(names_from = c(year), values_from = ER_filled)



# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=ER_wide_yr$Lon,
                   y=ER_wide_yr$Lat,
                   z=ER_wide_yr[,4:ncol(ER_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

ER_corrgram_summary.table <- summary(RR_df.corr_yr)
ER_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly")

## Sample average line
ER.corr.pred.df.yr$mean_synch <- mean(ER.corr.pred.df.yr$y, na.rm = TRUE)


ER_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly")

```



### combined synch 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_ER <- bind_rows(ER.corr.pred.df.day, ER.corr.pred.df.wk, ER.corr.pred.df.mon, ER.corr.pred.df.qrtly, ER.corr.pred.df.yr)
ci_df_all_ER   <- bind_rows(ER.corr.boot.df.day, ER.corr.boot.df.wk, ER.corr.boot.df.mon, ER.corr.boot.df.qrtly, ER.corr.boot.df.yr)

```

#### ER: Look when synchrony dips below the sample average

This gives an idea for at what distances stream ER is no longer synchronous 

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}
synchrony_approx_mean_lowest_x <- pred_df_all_ER %>%
  mutate(
    mean_synch = round(mean_synch, 3),
    y = round(y, 3),
    diff = abs(y - mean_synch)
  ) %>%
  group_by(time_scale) %>%
  # Filter only rows where y is approximately equal to mean_synch
  filter(diff <= 0.001) %>%
  # Within those, get the row with the lowest x
  slice_min(order_by = x, n = 1) %>%
  ungroup()

synchrony_approx_mean_lowest_x
```

Looks like ER snychrony breaks down at 1105 km for daily dynamics and 2429 km for yearly (which is bascall the whole length of the US) 





```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=12, fig.height=8}

ggplot() +
  geom_ribbon(data = ci_df_all_ER,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  geom_line(data = pred_df_all_ER,
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  ##
    labs(title = "Synchrony in ER across different time scales", 
       y= "Correlation", 
       x="Distance (km)") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic()


```



```{r, echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}


# Step 1: Spread the data so each time_scale becomes a column
pred_wide_ER <- pred_df_all_ER %>%
  dplyr::select(x,y,time_scale)%>%
  pivot_wider(names_from = time_scale, values_from = y)

# Step 2: Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_ER <- pred_wide_ER %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )



# Step 1: Pivot longer
pred_long_diff_ER <- pred_diff_ER %>%
 dplyr:: select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )


# Step 3: Plot histogram
ggplot(pred_long_diff_ER, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```






Add points for each stream to get a sense of geographic distance of predicted values.


```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)

### add point for mean ER to plot 
library(geosphere)
# Get site centroids (optional, or pick one reference site)
site_coords <- ER_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (e.g., the upstream-most one)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

ER_long_dist <- ER_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")


ER_long_dist_plt <- ER_long_dist %>%
  group_by(Site_ID, distance_km) %>%
  summarise(
    ER_mean = (median(ER_filled, na.rm = TRUE)*-1), 
    ER_low = (quantile(ER_filled, probs = 0.25, na.rm = TRUE)*-1), 
    ER_up = (quantile(ER_filled, probs = 0.75, na.rm = TRUE)*-1),
    .groups = "drop"
  )

###
ER_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_ER%>%filter(x<3250),
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_ER%>%filter(x<3250),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  # Horizontal mean synchrony lines
  geom_hline(data = pred_df_all_ER%>%filter(x<3100),
             aes(yintercept = mean_synch, color = time_scale),
             linetype = "dashed") +
  # Add ER points (mean)
  geom_point(data = ER_long_dist_plt,
             aes(x = distance_km, y = ER_mean/25),
             shape = 1, size = 2,
             alpha=0.5) +
  # Add ER error bars
  geom_errorbar(data = ER_long_dist_plt,
                aes(x = distance_km, ymin = ER_low/25, ymax = ER_up/25),
                width = 2, color = "black", 
                alpha=0.25) +
  ylim(-0.2, 0.75) +
  # Labels and theme
  labs(title = "Synchrony in ER across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff_ER%>%filter(x<3250), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(5, 5, 5, 5),
        legend.position="none")

# Combine using patchwork (put rotated density plot on the right)
combined_plot <- ER_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```




```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/ER_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```



```{r, size='small', echo=T, include=T, warning=FALSE}
### look at an element of environmental variation -- lets look at spring flow
RR_dat_sum <-  RR_dat%>%
  mutate(month=month(date)) %>%
  filter(month>1 & month<=6) %>%
  group_by(Site_ID, year, Lon, Lat) %>%
  summarise(Q_max= max(discharge, na.rm=T),
            Q_anl= mean(discharge, na.rm=T))

# choosing max flow:
RR_dat_TS_Q <- RR_dat_sum %>%
  dplyr::select(Site_ID,Lon, Lat, year, Q_max) %>%
  pivot_wider(names_from = c(year), values_from = Q_max)


RR_dat_TS_Q <-na.omit(RR_dat_TS_Q)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
Q_RR_df.corr <- Sncf(x=RR_dat_TS_Q$Lon,
                   y=RR_dat_TS_Q$Lat,
                   z=RR_dat_TS_Q[,4:ncol(RR_dat_TS_Q)],
                   resamp = 1000,
                   latlon = T)

```




```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# Q_corrgram_summary.table <- summary(Q_RR_df.corr)
# 
# # Plot using ggplot for export
# # create a dataframe of the predicted values
# Q.corr.pred.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$real$predicted$x)),
#                                y = matrix(unlist(Q_RR_df.corr$real$predicted$y)))
# 
# 
# # ...and a dataframe of the y values for the bootstrap prediction
# Q.corr.bootValues.df <- as.data.frame(t(Q_RR_df.corr$boot$boot.summary$predicted$y))
# 
# # merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
# Q.corr.boot.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$boot$boot.summary$predicted$x)),
#                                yminQ = Q.corr.bootValues.df$`0.025`,
#                                ymaxQ = Q.corr.bootValues.df$`0.975`)
# 
# 
# 
# RR_dat_sum <-  RR_dat%>%
#   group_by(Site_ID, year, Lon, Lat) %>%
#   summarise(GPP_an= mean(GPP_filled, na.rm=T),
#             GPP_max= max(GPP_filled, na.rm=T),
#         Q_anl= mean(discharge, na.rm=T),
#         wtemp_min= min(temp.water, na.rm=T),
#         wtemp_max= max(temp.water, na.rm=T),
#         wtemp_an= mean(temp.water, na.rm=T),
#         ppt_ann= mean(ppt_mm, na.rm=T),
#         Atemp_ann= max(TmeanC, na.rm=T))
# 
# 
# ### Alright lets look at Flow
# RR_dat_TS_temp <- RR_dat_sum %>%
#   dplyr::select(Site_ID,Lon, Lat, year, wtemp_an) %>%
#   pivot_wider(names_from = c(year), values_from = wtemp_an) 
# 
# 
# RR_dat_TS_temp <-na.omit(RR_dat_TS_temp)
# 
# 
# # Now, calculate spatial Synchrony in the time series using the Sncf() function
# temp_RR_df.corr <- Sncf(x=RR_dat_TS_temp$Lon,
#                      y=RR_dat_TS_temp$Lat,
#                      z=RR_dat_TS_temp[,4:ncol(RR_dat_TS_temp)],
#                      resamp = 1000, 
#                      latlon = T)
# 
# plot(temp_RR_df.corr)
# 
# temp_corrgram_summary.table <- summary(temp_RR_df.corr)
# 
# # Plot using ggplot for export
# # create a dataframe of the predicted values
# temp.corr.pred.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$real$predicted$x)),
#                              y = matrix(unlist(temp_RR_df.corr$real$predicted$y)))
# 
# 
# # ...and a dataframe of the y values for the bootstrap prediction
# temp.corr.bootValues.df <- as.data.frame(t(temp_RR_df.corr$boot$boot.summary$predicted$y))
# 
# # merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
# temp.corr.boot.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$boot$boot.summary$predicted$x)),
#                              yminWtemp = temp.corr.bootValues.df$`0.025`,
#                              ymaxWtemp = temp.corr.bootValues.df$`0.975`)
# 
# 
# 
# RR_dat_sum <-  RR_dat%>%
#   group_by(Site_ID, year, Lon, Lat) %>%
#   summarise(GPP_an= mean(GPP_filled, na.rm=T),
#             GPP_max= max(GPP_filled, na.rm=T),
#         Q_anl= mean(discharge, na.rm=T),
#         wtemp_min= min(temp.water, na.rm=T),
#         wtemp_max= max(temp.water, na.rm=T),
#         wtemp_an= mean(temp.water, na.rm=T),
#         ppt_ann= mean(ppt_mm, na.rm=T),
#         Atemp_ann= max(TmeanC, na.rm=T))
# 
# 
# RR_dat_TS_ppt <- RR_dat_sum %>%
#   dplyr::select(Site_ID,Lon, Lat, year, ppt_ann) %>%
#   pivot_wider(names_from = c(year), values_from = ppt_ann) 
# 
# 
# RR_dat_TS_ppt <-na.omit(RR_dat_TS_ppt)
# 
# 
# 
# # Now, calculate spatial Synchrony in the time series using the Sncf() function
# ppt_RR_df.corr <- Sncf(x=RR_dat_TS_ppt$Lon,
#                         y=RR_dat_TS_ppt$Lat,
#                         z=RR_dat_TS_ppt[,4:ncol(RR_dat_TS_ppt)],
#                         resamp = 1000, 
#                         latlon = T)
# 
# plot(ppt_RR_df.corr)
# 
# ppt_corrgram_summary.table <- summary(ppt_RR_df.corr)
# 
# # Plot using ggplot for export
# # create a dataframe of the predicted values
# ppt.corr.pred.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$real$predicted$x)),
#                                 y = matrix(unlist(ppt_RR_df.corr$real$predicted$y)))
# 
# 
# # ...and a dataframe of the y values for the bootstrap prediction
# ppt.corr.bootValues.df <- as.data.frame(t(ppt_RR_df.corr$boot$boot.summary$predicted$y))
# 
# # merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
# ppt.corr.boot.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$boot$boot.summary$predicted$x)),
#                                 yminPPT = ppt.corr.bootValues.df$`0.025`,
#                                 ymaxPPT = ppt.corr.bootValues.df$`0.975`)
# 
# ### Alright lets look max summer temp
# RR_dat_sum <-  RR_dat%>%
#   mutate(month =month(date)) %>%
#   filter(month>5 & month<10) %>%
#   group_by(Site_ID, year, Lon, Lat) %>%
#   summarise(wtemp_max= max(temp.water, na.rm=T),
#             wtemp_an= mean(temp.water, na.rm=T))
# 
# 
# 
# RR_dat_sum <-  RR_dat%>%
#   group_by(Site_ID, year, Lon, Lat) %>%
#   summarise(GPP_an= mean(GPP_filled, na.rm=T),
#             GPP_max= max(GPP_filled, na.rm=T),
#         Q_anl= mean(discharge, na.rm=T),
#         wtemp_min= min(temp.water, na.rm=T),
#         wtemp_max= max(temp.water, na.rm=T),
#         wtemp_an= mean(temp.water, na.rm=T),
#         ppt_ann= mean(ppt_mm, na.rm=T),
#         Atemp_ann= max(TmeanC, na.rm=T))
# 
# 
# RR_dat_TS_Atemp <- RR_dat_sum %>%
#   dplyr::select(Site_ID,Lon, Lat, year, Atemp_ann) %>%
#   pivot_wider(names_from = c(year), values_from = Atemp_ann) 
# 
# 
# RR_dat_TS_Atemp <-na.omit(RR_dat_TS_Atemp)
# 
# 
# 
# 
# 
# # Now, calculate spatial Synchrony in the time series using the Sncf() function
# Atemp_RR_df.corr <- Sncf(x=RR_dat_TS_Atemp$Lon,
#                        y=RR_dat_TS_Atemp$Lat,
#                        z=RR_dat_TS_Atemp[,4:ncol(RR_dat_TS_Atemp)],
#                        resamp = 1000, 
#                        latlon = T)
# 
# plot(Atemp_RR_df.corr)
# 
# Atemp_corrgram_summary.table <- summary(Atemp_RR_df.corr)
# 
# # Plot using ggplot for export
# # create a dataframe of the predicted values
# Atemp.corr.pred.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$real$predicted$x)),
#                                y = matrix(unlist(Atemp_RR_df.corr$real$predicted$y)))
# 
# 
# # ...and a dataframe of the y values for the bootstrap prediction
# Atemp.corr.bootValues.df <- as.data.frame(t(Atemp_RR_df.corr$boot$boot.summary$predicted$y))
# 
# # merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
# Atemp.corr.boot.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$boot$boot.summary$predicted$x)),
#                                yminAtemp = Atemp.corr.bootValues.df$`0.025`,
#                                ymaxAtemp = Atemp.corr.bootValues.df$`0.975`)
# 
# 
# 
# ############
# big_boot_df<- GPP.corr.boot.df %>%
#   left_join(Q.corr.boot.df, by=c("x")) %>%
#   left_join(temp.corr.boot.df, by=c("x")) %>%
#   left_join(ppt.corr.boot.df, by=c("x")) %>%
#   left_join(Atemp.corr.boot.df, by=c("x")) 
#   
```

## Correlogam of Synchrony of GPP + other hydro/climate/environmental parameters near streams

```{r, size = 'large', echo=FALSE, include=T, warning=FALSE}
# 
# pred_df <- bind_rows(
#   Q.corr.pred.df %>% mutate(group = "Flow"),
#   GPP.corr.pred.df %>% mutate(group = "GPP"),
#   temp.corr.pred.df %>% mutate(group = "Water Temp"),
#   ppt.corr.pred.df %>% mutate(group = "PPT"),
#   Atemp.corr.pred.df %>% mutate(group = "Air Temp")
# )
# 
# ribbon_df <- big_boot_df %>%
#   transmute(
#     x = x,
#     Flow_ymin = yminQ, Flow_ymax = ymaxQ,
#     GPP_ymin = yminGPP, GPP_ymax = ymaxGPP,
#     WaterTemp_ymin = yminWtemp, WaterTemp_ymax = ymaxWtemp,
#     PPT_ymin = yminPPT, PPT_ymax = ymaxPPT,
#     AirTemp_ymin = yminAtemp, AirTemp_ymax = ymaxAtemp
#   ) %>%
#   pivot_longer(cols = -x, names_to = c("group", ".value"), names_sep = "_")
# 
# synch.plot <- ggplot() +
#   geom_line(data = pred_df,
#             aes(x = x, y = y, color = group)) +
#   geom_hline(yintercept = 0) +
#   theme_classic() +
#   labs(title = "Synchrony across GPP, flow, temp, and ppt",
#               y= "Correlation", 
#        x="Distance (km)",
#        color = "Variable", fill = "Variable") +
#   theme(axis.title.x = element_blank(),
#         axis.title.y = element_blank(),
#         text = element_text(size=14),
#         legend.position = "bottom") +
#   scale_color_manual(values = c(
#     "Flow" = "#6293c4",
#     "GPP" = "#5a8f5e",
#     "Water Temp" = "#DAA520",
#     "PPT" = "#654cad",
#     "Air Temp" = "#b82c02"
#   )) +
#    geom_ribbon(data = big_boot_df,
#               aes(x = x, ymin = yminQ, ymax = ymaxQ),
#               fill = "#6293c4", alpha=0.1) +
#   geom_ribbon(data = big_boot_df,
#               aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
#               fill = "#5a8f5e", alpha=0.1) +
#   geom_ribbon(data = big_boot_df,
#               aes(x = x, ymin = yminWtemp, ymax = ymaxWtemp),
#               fill = "#DAA520", alpha=0.1) +
#   geom_ribbon(data = big_boot_df,
#               aes(x = x, ymin = yminPPT, ymax = ymaxPPT),
#               fill = "#654cad", alpha=0.1) +
#   geom_ribbon(data = big_boot_df,
#               aes(x = x, ymin = yminAtemp, ymax = ymaxAtemp),
#               fill = "#b82c02", alpha=0.1)
# 
# 
# synch.plot
# 

```
