---
title: "Part 3. Spline Correlogram analysis for synchrony in aquatic ecosystem productivity - Lanscape position"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables
library(broom)
library(grid)

```

## Background: Synchrony estimates using Sncf{ncf}

#### How it works:

Sncf computes the nonparametric spatial (auto/cross-)correlation
function for multivariate time series data observed across spatial
locations.

1.  Takes time series data z from n locations (one row per location, one
    column per time step).

2.  Calculates the pairwise correlations between time series at each
    location pair.

3.  Bins those correlations based on distance between sites (using
    latitude and longitude if latlon = TRUE).

4.  Fits a smoothing spline to estimate the average correlation as a
    function of distance
    
It can handle missing data, where:
na.rm = TRUE: Missing values are allowed; only overlapping time steps are used in correlations.


#### Metrics to consider:

-   '\real$predicted\$y[1]' = Synchrony at (almost) zero distance –
    often max synchrony.
      
      - This is the first predicted value from the fitted nonparametric spline of the spatial correlation function
      
      - Estimated synchrony at the smallest distance bin


-   '\$y.intercept' = Extrapolated synchrony at exactly zero distance 
      
      - The theoretical synchrony among perfectly co-located sites

-   '\$cbar' = Regional mean synchrony across all distances

-   '\$x.intercept'	= Distance at which synchrony drops to 0

-   '\$e.intercept' = Distance where synchrony drops to 1/e of max (decay scale)


###  Fletcher, R., & Fortin, M. (2018):
we should truncate the range of lag distances at which we consider spatial 
dependence to approximately ½ to 2 = 3 the total distance observed.

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

river_metab_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/stream_spline_cor_df.rds") %>% mutate(GPP_filled = if_else(GPP_filled <= 0, 0.0001, GPP_filled),
          ER_filled = if_else(ER_filled >= 0, -0.0001, ER_filled))

names(river_metab_dat)
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
### read in prism data for PPT

PPT_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Clim_data/PRISM_stream.rds")%>%
  mutate(Site_ID=as.character(Name))


StreamLight_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/7_river_light.csv") %>%
  mutate(date = lubridate::as_date(date, format="%Y-%m-%d")) %>%
  mutate(Site_ID1 = str_remove(site, "^nwis_")) %>%
  mutate(Site_ID = str_remove(Site_ID1, "^0+"))


river_metab_dat1 <- river_metab_dat%>%
  left_join(PPT_dat, by=c("date"="Date", "Site_ID")) %>%
  dplyr::select(site:TmaxC) 

river_metab_dat1 <- river_metab_dat1%>%
  left_join(StreamLight_dat, by=c("date", "site", "Site_ID"))

intersect(unique(river_metab_dat1$site), unique(StreamLight_dat$site))


setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab



sites_keep <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Map_data_for_nested_reach.csv")%>%
  filter(keep == 1) %>%
  dplyr::select(Site_ID:stream_order)


# Filter river_metab_dat to keep only Site_IDs in sites_keep
river_metab_dat1 <- river_metab_dat1 %>%
  filter(Site_ID %in% sites_keep$Site_ID)

unique(river_metab_dat1$Site_ID)
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
library(dplyr)
library(geodist)

# Get site centroids
site_coords <- river_metab_dat1 %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Calculate the overall mean coordinates
mean_lat <- median(site_coords$Lat)
mean_lon <- median(site_coords$Lon)

# Find site closest to mean coordinates using geodist
mean_coords <- data.frame(Lat = mean_lat, Lon = mean_lon)
site_coords_df <- data.frame(Lat = site_coords$Lat, Lon = site_coords$Lon)

distances_to_mean <- geodist(site_coords_df, mean_coords, measure = "geodesic") / 1000
closest_index <- which.min(distances_to_mean)

# Use the closest site as the reference
ref_site <- site_coords[closest_index, ]

# Compute distances from the reference site to all others
ref_coords <- data.frame(Lat = ref_site$Lat, Lon = ref_site$Lon)
distances_from_ref <- geodist(site_coords_df, ref_coords, measure = "geodesic") / 1000

# Add distance column
site_coords <- site_coords %>%
  mutate(distance_km = as.numeric(distances_from_ref))

hist(site_coords$distance_km)


# Get 2/3 cut off 

half_dist_km <- max(site_coords$distance_km)/ 2
half_dist_km
twothirds_dist_km <- max(site_coords$distance_km) * (2/3)
twothirds_dist_km

```

## Part 1: Synchrony in metabolism at differnt timescales

Compare synchrony for GPP and ER from 56 different sites in Marzolf
dataset at different timescales:

### A. GPP synchrony at differnt timescales:

#### GPP- Daily synchrony

```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}

#### Daily synchrony

river_metab_dat<-as.data.frame(river_metab_dat1)

GPP_long <- river_metab_dat1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day <- GPP_long %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_day <- Sncf(x=GPP_wide_day$Lon,
                   y=GPP_wide_day$Lat,
                   z=GPP_wide_day[,4:ncol(GPP_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year

plot(RR_df.corr_day)

GPP_corrgram_summary.table <- summary(RR_df.corr_day)
Regional.synch <- GPP_corrgram_summary.table$Regional.synch


# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP")

## Sample average line
GPP.corr.pred.df.day$mean_synch <- round(Regional.synch,2)
Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.day$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP")



```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```


## Part 2 - Moran Effect

-   MODIS data for leaf area index (LAI)

-   Daily streamflow (Q)

-   Daily water temperature (water_temp)

-   Daily air temp (air_temp)

-   Daily precipitation (PPT)

#### Bring in MODIS data for LAI to get at

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

LAI_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Modis_data/StreamDat_MOD_combined.rds")%>%
  mutate(date=as.Date(Date)) %>%
    filter(Site_ID %in% sites_keep$Site_ID)


# intersect(unique(river_metab_dat1$Site_ID), unique(LAI_dat$Site_ID))
# 
# 
# Metat_sites_notinlight <- setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
# Light_sites_notinmetab<- setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab
# 
# LAI_sites_notinlight<- setdiff(unique(LAI_dat$Site_ID), unique(StreamLight_dat$Site_ID)) # Sites in light but not in metab

```

## Part 3. Landscape level factors as explaining synchrony:

A. left join time series dat with landscape parameters

B. ER / GPP explained by stream size (via regression)

i.  Dominant land cover explained by stream size


iii. Slope/ Catchment size/ basin characteristics

iv. Water temperature/ flow

C. Create landscape level groups and then run synchronous

i.  landcover (PCA? developed to forest?) - possibly cropland with
    moderate development

iii. Slope? or Catchment size

iv. Water temperature/ depth/flow?

<!-- -->

5.  When landscape character explains some synchronous, visualize with
    synch splines (Q1-Q3 for i to iv splines)

```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

dat_NHDinfo<- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata_25v2.rds")%>%
  mutate(Site_ID = str_remove(site_no, "^0+"))

river_metab_dat1 <- river_metab_dat1%>%
  left_join(dat_NHDinfo, by=c("Site_ID")) 

```

#####  Random forest to get at what should describe National Land Cover Database

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}
library(randomForest)

# Prepare data for RF
rf_LC_data <- river_metab_dat1[, c("Developed_OpenSpace_pct", "Developed_LowIntensity_pct",
                                   "Tree_Canopyin100mRip", #"daily_LAI",
                       "Developed_MedIntensity_pct", "Developed_HiIntensity_pct",
                       "Barren_Land_pct", "Forest_Deciduous_pct",
                       "Water_Open_pct", "Ice_Snow_Perennial_pct",
                       "Shrub_Scrub_pct", "Grass_landHerbaceous_pct",
                       "Forest_Evergreen_pct", "Forest_mixed_pct",
                       "Grass_PastureHay_pct", "Cultivated_Crops_pct",
                       "Wetlands_Woody_pct","Wetlands_EmergentHerb_pct",
                       "road_Rural_den", "road_private_den", "road_Total_den")]


rf_LC_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_LC_data)
rf_LC_input_clean <- na.omit(rf_LC_input)

# Now run the model
rf_LC_model <- randomForest(GPP_synch ~ ., data = rf_LC_input_clean,
                         importance = TRUE, ntree = 200)

# Plot importance
varImpPlot(rf_LC_model)

# Get importance metrics from your existing model
importance_vals <- randomForest::importance(rf_LC_model, type = 1)  # type = 1 for %IncMSE
importance_df <- data.frame(variable = rownames(importance_vals),
                            IncMSE = importance_vals[, 1])

importance_df

# Sort by importance and keep top 10
top_vars_LC <- importance_df %>%
  arrange(desc(IncMSE)) %>%
  slice(1:20) %>%
  pull(variable)

top_vars_LC

```

###### Create the composite variables for land cover/geology/basin characteristics

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Categorize summary columns based on their initial patterns
col_patterns <- c("alt_va" = "reach morphology", "drain_area_va" = "drainage_area", "Water_Open_pct" = "water",
                  "Ice_Snow_Perennial_pct" = "ice_snow", "Developed_OpenSpace_pct" = "open_space",
                  "Developed_LowIntensity_pct" = "open_space", "Developed_MedIntensity_pct" = "developed",
                  "Developed_HiIntensity_pct" = "developed", "Barren_Land_pct" = "open_space",
                  "Forest_Deciduous_pct" = "Forest_Deciduous_pct", "Forest_Evergreen_pct" = "Forest_Evergreen_pct", 
                  "Forest_mixed_pct" = "forest", "Shrub_Scrub_pct" = "shrub", 
                  "Grass_landHerbaceous_pct" = "shrub", "Grass_PastureHay_pct" = "crops", 
                  "Cultivated_Crops_pct" = "crops", "Wetlands_Woody_pct" = "wetlands", 
                  "Wetlands_EmergentHerb_pct" = "wetlands", "geol_gneiss" = "geology_metamorphic", 
                  "geol_granitic" = "geology_igneous", "geol_ultramafic" = "geology_metamorphic", 
                  "geol_quarternary" = "geology_sedimentary", "geol_sedimentary" = "geology_sedimentary", 
                  "geol_volcanic" = "geology_igneous", "geol_intermediate_pultonic" = "geology_igneous", 
                  "CAT_BASIN_AREA" = "drainage_area", "CAT_BASIN_SLOPE" = "slope", 
                  "CAT_ELEV_MEAN" = "elevation", "CAT_ELEV_MIN" = "elevation", 
                  "CAT_ELEV_MAX" = "elevation", "CAT_STREAM_SLOPE" = "reach morphology", 
                  "CAT_STREAM_LENGTH" = "stream_length", "road_Rural_den" = "road_density", 
                  "road_private_den" = "road_density", "road_Total_den" = "road_density", 
                  "Tree_Canopyin100mRip" = "shading", 
                  "daily_LAI" = "shading",
                  "temp.water" = "water quality", "DO.obs" = "water quality",
                  "depth" = "reach morphology", "discharge" = "discharge",
                  "pathlength" = "hydrologic hierarchy", "lengthkm" = "geometry",
                 "shape_length"=  "geometry", "streamleve"= "hydrologic hierarchy",
                 "streamorde"=  "hydrologic hierarchy", 
                 "streamcalc" = "hydrologic hierarchy", 
                 "totdasqkm" = "drainage_area", "slope" = "slope", "pathlength" = "hydrologic hierarchy" )

```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=6}
### start the PCA 

LC_pca_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID, all_of(top_vars_LC)) %>%
  mutate(
    crop = c(Grass_landHerbaceous_pct+ Cultivated_Crops_pct),
    developed = c(Developed_MedIntensity_pct + Developed_HiIntensity_pct + Developed_LowIntensity_pct + road_Total_den), 
    forest = c(Forest_mixed_pct + Forest_Evergreen_pct + Forest_Deciduous_pct),
    water = c(Water_Open_pct + Ice_Snow_Perennial_pct),
    shrub= c(Shrub_Scrub_pct + Wetlands_EmergentHerb_pct + Barren_Land_pct),
    shading = c(Tree_Canopyin100mRip)
  ) %>%
  dplyr::select(Site_ID, crop, developed, forest, water, shrub, shading)

col_patterns <- c("crop" = "crop",
                  "developed" = "developed",
                  "forest" = "forest",
                  "water" = "water",
                  "shrub" = "shrub",
                   "shading" = "shading")

LC_pca_data1 <- LC_pca_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
LC_pca_data <- LC_pca_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

LC_pca_input_clean <- na.omit(LC_pca_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(LC_pca_data), function(col) {
  col_patterns[col]
}))

pca_LC <- prcomp(LC_pca_input_clean, center = TRUE, scale. = TRUE)
summary(pca_LC)

library(factoextra)

# Visualize PCA eigenvalues
scree_plt_LC <- fviz_eig(pca_LC)
scree_plt_LC

# ggsave(plot = scree_plt_LC, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_LC_25.png",sep=""),width=3,height=2.5,dpi=300)


biplot(prcomp(LC_pca_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_LC <- fviz_pca_biplot(pca_LC, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories,  ggtheme = theme_classic())
    


library(factoextra)
library(ggplot2)

watershed_loadplot_LC
# ggsave(plot = watershed_loadplot_LC, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_LC_25.png",sep=""),width=4.5,height=3,dpi=300)


```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)
LC_pca_scores <- as.data.frame(pca_LC$x)

LC_pca_input_clean_sites <- na.omit(LC_pca_data1)

# Add sites
LC_pca_with_site <- cbind(Site_ID=LC_pca_input_clean_sites$Site_ID, LC_pca_scores)

# Add GPP_synch back in (at the annual scale)
LC_pca_with_response <- river_metab_dat1 %>%
  left_join(LC_pca_with_site, by=c("Site_ID"))


hist(LC_pca_with_response$PC1)
summary(LC_pca_with_response$PC1)
```

###### A. How much variance in GPP is explained by landscape?

*Land cover (LC) PCA:* Represents gradient from developed to
natural/forested areas, to agricultural landscapes 

-   High PC1 scores (positive values): Associated with urbanized or developed areas, especially medium and high intensity development 

-   Moderate PC1 scores natural forested 

-   Low PC1 scores (negative values): Associated with agricultural and herbaceous/ landscapes, like crops, pasture, and grasslands.


```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=4}

# lm_model_LC <- lm(GPP_filled ~ PC1 +PC2,family=lognormal(), data = LC_pca_with_response)
# summary(lm_model_LC)
# 
# hist(residuals(lm_model_LC)) # meh
```

###### BRMs

Very similar to glm

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
library(brms)

# Bayesian linear model
brm_model <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = LC_pca_with_response,
  family = gaussian,  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)


summary(brm_model)
plot(brm_model)  # Diagnostic plots

# bayes_R2(brm_model)

```


###### Quantile of different land cover type:

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
pc1_vals <- na.omit(LC_pca_with_response$PC1)

# Get 1/3 cutoffs (33.3% and 66.6%)
cut_points <- quantile(pc1_vals, probs = c(1/3, 2/3))

# Create group labels
LC_pca_with_response$PC1_group <- cut(
  LC_pca_with_response$PC1,
  breaks = c(-Inf, cut_points[1], cut_points[2], Inf),
  labels = c("Q1", "Q2", "Q3"),
  right = TRUE
)


tab_LC_pca_with_response <- LC_pca_with_response%>%
  group_by(PC1_group) %>%
  summarise(PC1_min=round(min(PC1, na.rm=T),2),
            PC1_max=round(max(PC1, na.rm=T),2), 
            site_no =length(unique(Site_ID)))


tab_LC_pca_with_response

```

LC is fairly different across the 3 different groups

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Q1" = "#00BFC4",   
  "Q2" = "#7CAE00",
  "Q3" = "#F8766D"
)
```


```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(LC_pca_with_response, aes(x = PC1, fill = factor(PC1_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Forested","Developed", "Ag.")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by LC quantile",
       x = "PC1",
       y = "Count")
```


```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

metab_units_area <- (expression(paste('Metabolism (g C', m^-2,' ',y^-1,')', sep = ' ')))


low_p <- '#6e2f9e'
high_p <- '#f4e6ff'

low_er <- '#381d01'
high_er <- '#fa9600'


LC_pca_plot <- LC_pca_with_response%>%
  group_by(Site_ID) %>%
  summarise(
    GPP = mean(GPP_filled, na.rm=T),
    ER = mean(c(ER_filled*-1), na.rm=T),
    PC1 =mean(PC1, na.rm=T))


library(ggplot2)
library(ggnewscale)  

fig_combined <- ggplot(LC_pca_plot, aes(x = PC1)) +
  # GPP points (upward triangle)
  geom_point(aes(y = GPP, fill = GPP, color = GPP), shape = 24, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "GPP", low = low_p, high = high_p) +
  scale_color_gradient(name = "GPP", low = low_p, high = high_p) +
  # Reset fill scale so ER can have its own
  new_scale_fill() +
    new_scale_color() +
  # ER points (downward triangle)
  geom_point(aes(y = ER, fill = ER, color = ER), shape = 25, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "ER", low = low_er, high = high_er) +
  scale_color_gradient(name = "ER", low = low_er, high = high_er) +
  scale_y_continuous(name = metab_units_area, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
        theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    panel.grid = element_blank()
  )
# ggsave(plot = fig_combined, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/LC_fig1b_25.png",sep=""),width=5.25,height=3.5,dpi=300)

```

However daily GPP looks fairly similar across all land cover types -
despite the postive regression resuls with cropland

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(LC_pca_with_response, aes(x = GPP_filled, fill = factor(PC1_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
        labels = c("Forested","Developed", "Ag.")) +
  theme_minimal() +
  labs(title = "GPP by LC quantile",
       x = "GPP",
       y = "Count")
```

###### i. Land cover Q1

PC1 

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q1 <- LC_pca_with_response %>%
dplyr::filter(PC1_group=="Q1") 

unique(land_cover_q1$Site_ID)
unique(land_cover_q1$Latitude)

hist(land_cover_q1$GPP_filled,  breaks = 30)

```

###### ii. Land cover Q2

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q2 <- LC_pca_with_response %>%
dplyr::filter(PC1_group=="Q2") 


unique(land_cover_q2$Site_ID)
unique(land_cover_q2$Latitude)

hist(land_cover_q2$GPP_filled)

```

###### iii. Land cover Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

land_cover_q3 <- LC_pca_with_response %>%
dplyr::filter(PC1_group=="Q3")
unique(land_cover_q3$Site_ID)
unique(land_cover_q3$Latitude)

hist(land_cover_q3$GPP_filled)

```

###### Daily synchrony in GPP based PC of land cover Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- land_cover_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_LC_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_LC_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_LC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_LC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Developed")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.LC.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.LC.Q1$cbar <- Regional.synch_x




# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q1 <- as.data.frame(t(RR_df.corr_LC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Developed")

```



```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


###### Daily synchrony in GPP land cover Q2

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_Q2 <- land_cover_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_Q2 <- GPP_long_Q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_Moderate <- Sncf(x=GPP_wide_day_Q2$Lon,
                   y=GPP_wide_day_Q2$Lat,
                   z=GPP_wide_day_Q2[,4:ncol(GPP_wide_day_Q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_Moderate)


GPP_corrgram_summary.table <- summary(RR_df.corr_Moderate)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_Moderate$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_Moderate$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Moderate")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.LC.Q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.LC.Q2$cbar <- Regional.synch_x



# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q2 <- as.data.frame(t(RR_df.corr_Moderate$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_Moderate$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Moderate")

```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

###### Daily synchrony in GPP land cover Q3

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_Q3 <- land_cover_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_Q3 <- GPP_long_Q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_Natural <- Sncf(x=GPP_wide_day_Q3$Lon,
                   y=GPP_wide_day_Q3$Lat,
                   z=GPP_wide_day_Q3[,4:ncol(GPP_wide_day_Q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_Natural)


GPP_corrgram_summary.table <- summary(RR_df.corr_Natural)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_Natural$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_Natural$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Natural")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.LC.Q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.LC.Q3$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.LC.Q3 <- as.data.frame(t(RR_df.corr_Natural$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_Natural$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.LC.Q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.LC.Q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Natural")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all <- bind_rows(GPP.corr.pred.LC.Q1, GPP.corr.pred.LC.Q2, GPP.corr.pred.LC.Q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all   <- bind_rows(GPP.corr.boot.LC.Q1, GPP.corr.boot.LC.Q2, GPP.corr.boot.LC.Q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)

```

###### Calculate the percent difference between different land cover types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    LQ1_diff = (abs(Developed - Developed) / Developed) * 100,
    LQ2_diff = (abs(Developed - Moderate) / Developed) * 100,
    LQ3_diff = (abs(Developed - Natural) / Developed) * 100)

pred_diff
```

### Figure 4a (?): Spline Correlogram of GPP at different landcover types


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=F, include=T}

intersection_df <- pred_df_all %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Developed" = "#00BFC4",   
  "Moderate" = "#7CAE00",
  "Natural" = "#F8766D"
)
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_LC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%dplyr::filter(x<half_dist_km),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%dplyr::filter(x<half_dist_km),
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.25, yend = y,
                   color = strat),
               linetype = "dashed") +
  ylim(-0.25, 0.75) +
  labs(title = "GPP and land cover",
       y = "Correlation",
       x = NULL,
       fill = "Land cover PC1",
       color = "Land cover PC1") +
  geom_vline(xintercept = 807, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                           labels = c("Forested","Developed", "Ag."))+
  scale_fill_manual(values = strat_colors,
                        labels = c("Forested","Developed", "Ag."))+
  theme_classic() +
  theme(legend.position = "right")

GPP_synch_spline_LC

```





Agricultural and forest influences sites may have the most synchronous
GPP.

###### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all %>%
  group_by(strat) %>%
  filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 4), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```


### ER

###### Daily synchrony in ER based PC of land cover Q1

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q1 <- land_cover_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q1 <- ER_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_LC_q1 <- Sncf(x=ER_wide_day_q1$Lon,
                   y=ER_wide_day_q1$Lat,
                   z=ER_wide_day_q1[,4:ncol(ER_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_LC_q1)


ER_corrgram_summary.table <- summary(RR_df.corr_LC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_LC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Developed")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.LC.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.LC.Q1$cbar <- Regional.synch_x




# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.LC.Q1 <- as.data.frame(t(RR_df.corr_LC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.LC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_LC_q1$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.LC.Q1$`0.025`,
                                             ymaxER = ER.corr.bootValues.LC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Developed")

```



```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


###### Daily synchrony in ER land cover Q2

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_Q2 <- land_cover_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_Q2 <- ER_long_Q2 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_Moderate <- Sncf(x=ER_wide_day_Q2$Lon,
                   y=ER_wide_day_Q2$Lat,
                   z=ER_wide_day_Q2[,4:ncol(ER_wide_day_Q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_Moderate)


ER_corrgram_summary.table <- summary(RR_df.corr_Moderate)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_Moderate$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_Moderate$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Moderate")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.LC.Q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.LC.Q2$cbar <- Regional.synch_x



# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.LC.Q2 <- as.data.frame(t(RR_df.corr_Moderate$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.LC.Q2 <- data.frame(x = matrix(unlist(RR_df.corr_Moderate$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.LC.Q2$`0.025`,
                                             ymaxER = ER.corr.bootValues.LC.Q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Moderate")

```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

###### Daily synchrony in ER land cover Q3

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_Q3 <- land_cover_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_Q3 <- ER_long_Q3 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_Natural <- Sncf(x=ER_wide_day_Q3$Lon,
                   y=ER_wide_day_Q3$Lat,
                   z=ER_wide_day_Q3[,4:ncol(ER_wide_day_Q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_Natural)


ER_corrgram_summary.table <- summary(RR_df.corr_Natural)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_Natural$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_Natural$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Natural")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.LC.Q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.LC.Q3$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.LC.Q3 <- as.data.frame(t(RR_df.corr_Natural$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.LC.Q3 <- data.frame(x = matrix(unlist(RR_df.corr_Natural$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.LC.Q3$`0.025`,
                                             ymaxER = ER.corr.bootValues.LC.Q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Natural")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all <- bind_rows(ER.corr.pred.LC.Q1, ER.corr.pred.LC.Q2, ER.corr.pred.LC.Q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all   <- bind_rows(ER.corr.boot.LC.Q1, ER.corr.boot.LC.Q2, ER.corr.boot.LC.Q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)

```

###### Calculate the percent difference between different land cover types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    LQ1_diff = (abs(Developed - Developed) / Developed) * 100,
    LQ2_diff = (abs(Developed - Moderate) / Developed) * 100,
    LQ3_diff = (abs(Developed - Natural) / Developed) * 100)

pred_diff
```

### Figure 4a (?): Spline Correlogram of ER at different landcover types


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=F, include=T}

intersection_df <- pred_df_all %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Developed" = "#00BFC4",   
  "Moderate" = "#7CAE00",
  "Natural" = "#F8766D"
)
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
ER_synch_spline_LC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%dplyr::filter(x<half_dist_km),
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%dplyr::filter(x<half_dist_km),
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Horizontal mean synchrony lines
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.25, yend = y,
                   color = strat),
               linetype = "dashed") +
  ylim(-0.25, 0.75) +
  labs(title = "ER and land cover",
       y = "Correlation",
       x = NULL,
       fill = "Land cover PC1",
       color = "Land cover PC1") +
    geom_vline(xintercept = 1083, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                           labels = c("Forested","Developed", "Ag."))+
  scale_fill_manual(values = strat_colors,
                        labels = c("Forested","Developed", "Ag."))+
  theme_classic() +
  theme(legend.position = "right")

ER_synch_spline_LC

```


Agricultural and forest influences sites may have the most synchronous
ER.

###### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all %>%
  group_by(strat) %>%
  filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 6), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

###


### C. Basin characteristics

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=12, fig.height=8}

# Prepare data for RF
rf_basin_data <- river_metab_dat1[,c("lengthkm", "shape_length", "streamleve", 
                                     "streamorde", "streamcalc", "totdasqkm", "slope",
                                     "pathlength")]


rf_basin_input <- data.frame(GPP_synch = river_metab_dat1$GPP_filled, rf_basin_data)
rf_basin_input_clean <- na.omit(rf_basin_input)

# Now run the model
rf_basin_model <- randomForest(GPP_synch ~ ., data = rf_basin_input_clean,
                         importance = TRUE, ntree = 200)

# Plot importance
varImpPlot(rf_basin_model)

# Get importance metrics from your existing model
importance_vals <- randomForest ::importance(rf_basin_model, type = 1)  # type = 1 for %IncMSE
importance_df <- data.frame(variable = rownames(importance_vals),
                            IncMSE = importance_vals[, 1])

importance_df
# Sort by importance and keep top 10
top_vars_basin <- importance_df %>%
  arrange(desc(IncMSE)) %>%
  slice(1:10) %>%
  pull(variable)
```

###### Create the composite variable for basin characteristics

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
### start the PCA 
pca_basin_dat <- river_metab_dat1 %>%
  dplyr::select(Site_ID,  all_of(top_vars_basin))

pca_basin_data1 <- pca_basin_dat %>%
  group_by(Site_ID) %>%
      summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

# Prepare data for PCA by removing non-numeric and non-relevant columns
pca_basin_data <- pca_basin_data1 %>% 
  dplyr::select(-Site_ID) %>% 
  dplyr::select(where(~ sd(., na.rm = TRUE) != 0)) %>%  # Drop columns with no variance
  as.matrix()

pca_basin_input_clean <- na.omit(pca_basin_data)

domains <- dat_NHDinfo$state

# Apply the categorization to column names
smry_categories <- factor(sapply(colnames(pca_basin_data), function(col) {
  col_patterns[col]
}))


pca_basin <- prcomp(pca_basin_input_clean, center = TRUE, scale. = TRUE)

# Visualize PCA eigenvalues
scree_plt_basin <- fviz_eig(pca_basin)
scree_plt_basin

# ggsave(plot = scree_plt_basin, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_screeplt_basin_25.png",sep=""),width=4,height=3.5,dpi=300)


biplot(prcomp(pca_basin_input_clean, center = TRUE, scale. = TRUE))


# Visualize PCA biplot with variable categories
watershed_loadplot_basin<- fviz_pca_biplot(pca_basin, geom.var = 'arrow', geom.ind = 'point', title = '',
                col.var = smry_categories,  ggtheme = theme_classic())

watershed_loadplot_basin

# ggsave(plot = watershed_loadplot_basin, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Watershed_attributes_load_basin_25.png",sep=""),width=4.5,height=3,dpi=300)

```

*PC1:*

-   Higher values represent larger, higher order, more dendritic streams
    while lower values represent steeper, smaller, lower-order streams

*PC2:*

-   High PC2 shorter streams with more branching (higher
    streamcalc/streamorde) while lower represents longer more linear
    streams

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
# Get the scores (principal component axes)

basin_pca_scores <- as.data.frame(pca_basin$x)

basin_pca_input_clean_sites <- na.omit(pca_basin_data1)

# Add sites
basin_pca_with_site <- cbind(Site_ID=basin_pca_input_clean_sites$Site_ID, basin_pca_scores)

# Add GPP_synch back in (at the annual scale)
basin_pca_with_response <- river_metab_dat %>%
  left_join(basin_pca_with_site, by=c("Site_ID"))

```

#### B. How much variance in GPP is explained by geology?

GLM:

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_basin <- lm(GPP_filled ~ PC1 + PC2, data = basin_pca_with_response)
summary(lm_model_basin)

hist(residuals(lm_model_basin)) # meh
```

Slope PC1: Showing lower GPP in large, steep, high-elevation basins

Adjusted R-squared: 0.01126

#### BRMs slope + 0.25

Very similar to glm again

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
# Bayesian linear model
brm_model_basin <- brm(
  formula = GPP_filled ~ PC1 + PC2,
  data = basin_pca_with_response,
  family = gaussian(),  # same as lm
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 123
)

summary(brm_model_basin)
plot(brm_model_basin)  # Diagnostic plots

# bayes_R2(brm_model)

```

#### C. Quantile of different basinogies:

```{r, size='large', echo=F, include=F, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
pc1_vals <- na.omit(basin_pca_with_response$PC1)

# Get 1/3 cutoffs (33.3% and 66.6%)
cut_points <- quantile(pc1_vals, probs = c(1/3, 2/3))

# Create group labels
basin_pca_with_response$PC1_group <- cut(
  basin_pca_with_response$PC1,
  breaks = c(-Inf, cut_points[1], cut_points[2], Inf),
  labels = c("Q1", "Q2", "Q3"),
  right = TRUE
)

table(basin_pca_with_response$PC1_group)



tab_basin_pca_with_response <- basin_pca_with_response%>%
  group_by(PC1_group) %>%
  summarise(PC1_min=round(min(PC1, na.rm=T),2),
            PC1_max=round(max(PC1, na.rm=T),2), 
            site_no =length(unique(Site_ID)))


tab_basin_pca_with_response



```

basin character groups look relatively more similar

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colorsC <- c(
  "Q1" = "#00BFC4",   
  "Q2" = "#7CAE00",
  "Q3" = "#F8766D"
)
```

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(basin_pca_with_response, aes(x = PC1, fill = factor(PC1_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Small steep", "Q2", "Large branched")
  ) +
  theme_minimal() +
  labs(title = "PC1 values by character quantile",
       x = "PC1",
       y = "Count")
```



```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

metab_units_area <- (expression(paste('Metabolism (g C', m^-2,' ',y^-1,')', sep = ' ')))
low_p <- '#6e2f9e'
high_p <- '#f4e6ff'

low_er <- '#381d01'
high_er <- '#fa9600'


BN_pca_plot <- basin_pca_with_response%>%
  group_by(Site_ID) %>%
  summarise(
    GPP = mean(GPP_filled, na.rm=T),
    ER = mean(c(ER_filled*-1), na.rm=T),
    PC1 =mean(PC1, na.rm=T))


library(ggplot2)
library(ggnewscale)  

fig_combined1 <- ggplot(BN_pca_plot, aes(x = PC1)) +
  # GPP points (upward triangle)
  geom_point(aes(y = GPP, fill = GPP, color = GPP), shape = 24, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "GPP", low = low_p, high = high_p) +
  scale_color_gradient(name = "GPP", low = low_p, high = high_p) +
  # Reset fill scale so ER can have its own
  new_scale_fill() +
    new_scale_color() +
  # ER points (downward triangle)
  geom_point(aes(y = ER, fill = ER, color = ER), shape = 25, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "ER", low = low_er, high = high_er) +
  scale_color_gradient(name = "ER", low = low_er, high = high_er) +
  scale_y_continuous(name = metab_units_area, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
      theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    panel.grid = element_blank()
  )

# ggsave(plot = fig_combined1, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/LC_fig1c_25.png",sep=""),width=5.25,height=3.5,dpi=300)

```


However daily GPP looks fairly similar across all land cover types -
despite the postive regression resuls with cropland

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

ggplot(basin_pca_with_response, aes(x = GPP_filled, fill = factor(PC1_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colorsC,
    name = "PC1 Quantile",
    labels = c("Small steep", "Q2", "Large branched")) + theme_minimal() +
  labs(title = "GPP by basin character quantile",
       x = "GPP",
       y = "Count")
```

##### i. basin character Q1

All open space or forested

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q1 <- basin_pca_with_response %>%
dplyr::filter(PC1_group=="Q1") 

unique(basin_q1$Site_ID)
unique(basin_q1$Latitude)

hist(basin_q1$GPP_filled,  breaks = 30)

```

##### ii. basin character Q2

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q2 <- basin_pca_with_response %>%
dplyr::filter(PC1_group=="Q2") 


unique(basin_q2$Site_ID)
unique(basin_q2$Latitude)

hist(basin_q2$GPP_filled)

```

##### iii. basin character Q3

```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

basin_q3 <- basin_pca_with_response %>%
dplyr::filter(PC1_group=="Q3")
unique(basin_q3$Site_ID)
unique(basin_q3$Latitude)

hist(basin_q3$GPP_filled)

```

#### Daily synchrony basin Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- basin_pca_with_response %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Steep_small")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.BC.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.BC.Q1$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.Q1 <- as.data.frame(t(RR_df.corr_BC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Steep_small")

```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony basin character Q2

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- basin_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Moderate")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.BC.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.BC.q2$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.q2 <- as.data.frame(t(RR_df.corr_BC_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Moderate")
```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony basin character Q3

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- basin_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_BC_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "Large_branched")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.BC.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.BC.q3$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.BC.q3 <- as.data.frame(t(RR_df.corr_BC_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.BC.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.BC.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "Large_branched")

```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_basin <- bind_rows(GPP.corr.pred.BC.Q1, GPP.corr.pred.BC.q2, GPP.corr.pred.BC.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_basin   <- bind_rows(GPP.corr.boot.BC.Q1, GPP.corr.boot.BC.q2, GPP.corr.boot.BC.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different basin character types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_basin <- pred_df_all_basin %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_basin <- pred_wide_basin %>%
  mutate(
    Steep_small_diff = (abs(Steep_small - Steep_small) / Steep_small) * 100,
    Moderate_diff = (abs(Steep_small - Moderate) / Steep_small) * 100,
    Large_branched_diff = (abs(Steep_small - Large_branched) / Steep_small) * 100)

pred_diff_basin
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Steep_small" = "#00BFC4",   
  "Moderate" = "#7CAE00",
  "Large_branched" = "#F8766D"
)
```

#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_basin %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_BC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_basin,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_basin,
            aes(x = x, y = y, color = strat),
            size = 1) +
   # Horizontal mean synchrony lines
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.25, yend = y,
                   color = strat),
               linetype = "dashed") +
    ylim(-0.25, 0.75) +
  labs(title = "GPP and morphology",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Stream type PC1",
       color = "Stream type PC1") +
    geom_vline(xintercept = 807, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Large branched", "Moderate", "Steep small")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Large branched", "Moderate", "Steep small")) +
  theme_classic() +
  theme(legend.position = "right")

GPP_synch_spline_BC

```


Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_basin %>%
  group_by(strat) %>%
    filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 4), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

### ER.

#### Daily synchrony basin Q1

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q1 <- basin_q1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q1 <- ER_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q1 <- Sncf(x=ER_wide_day_q1$Lon,
                   y=ER_wide_day_q1$Lat,
                   z=ER_wide_day_q1[,4:ncol(ER_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q1)


ER_corrgram_summary.table <- summary(RR_df.corr_BC_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Steep_small")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.BC.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.BC.Q1$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.BC.Q1 <- as.data.frame(t(RR_df.corr_BC_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.BC.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q1$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.BC.Q1$`0.025`,
                                             ymaxER = ER.corr.bootValues.BC.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Steep_small")

```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony basin character Q2

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q2 <- basin_q2 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q2 <- ER_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q2 <- Sncf(x=ER_wide_day_q2$Lon,
                   y=ER_wide_day_q2$Lat,
                   z=ER_wide_day_q2[,4:ncol(ER_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q2)


ER_corrgram_summary.table <- summary(RR_df.corr_BC_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Moderate")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.BC.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.BC.q2$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.BC.q2 <- as.data.frame(t(RR_df.corr_BC_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.BC.q2 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q2$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.BC.q2$`0.025`,
                                             ymaxER = ER.corr.bootValues.BC.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Moderate")
```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony basin character Q3

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q3 <- basin_q3 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q3 <- ER_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_BC_q3 <- Sncf(x=ER_wide_day_q3$Lon,
                   y=ER_wide_day_q3$Lat,
                   z=ER_wide_day_q3[,4:ncol(ER_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_BC_q3)


ER_corrgram_summary.table <- summary(RR_df.corr_BC_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_BC_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "Large_branched")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.BC.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.BC.q3$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.BC.q3 <- as.data.frame(t(RR_df.corr_BC_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.BC.q3 <- data.frame(x = matrix(unlist(RR_df.corr_BC_q3$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.BC.q3$`0.025`,
                                             ymaxER = ER.corr.bootValues.BC.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "Large_branched")

```



```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_basin <- bind_rows(ER.corr.pred.BC.Q1, ER.corr.pred.BC.q2, ER.corr.pred.BC.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_basin   <- bind_rows(ER.corr.boot.BC.Q1, ER.corr.boot.BC.q2, ER.corr.boot.BC.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different basin character types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_basin <- pred_df_all_basin %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_basin <- pred_wide_basin %>%
  mutate(
    Steep_small_diff = (abs(Steep_small - Steep_small) / Steep_small) * 100,
    Moderate_diff = (abs(Steep_small - Moderate) / Steep_small) * 100,
    Large_branched_diff = (abs(Steep_small - Large_branched) / Steep_small) * 100)

pred_diff_basin
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Steep_small" = "#00BFC4",   
  "Moderate" = "#7CAE00",
  "Large_branched" = "#F8766D"
)
```

#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_basin %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
ER_synch_spline_BC <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_basin,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_basin,
            aes(x = x, y = y, color = strat),
            size = 1) +
   # Horizontal mean synchrony lines
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.25, yend = y,
                   color = strat),
               linetype = "dashed") +
    ylim(-0.25, 0.75) +
  labs(title = "ER and morphology",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Stream type PC1",
       color = "Stream type PC1") +
      geom_vline(xintercept = 1083, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Large branched", "Moderate", "Steep small")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Large branched", "Moderate", "Steep small")) +
  theme_classic() +
  theme(legend.position = "right")

ER_synch_spline_BC

```


Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_basin %>%
  group_by(strat) %>%
      filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 4), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```


####

```{r, warning=F, size = 'small', echo=F,include=TRUE, fig.width=6, fig.height=9}
# Arrange plots with shared legend at the bottom
library(ggpubr)

spline_grid_landscape <- ggarrange(
  GPP_synch_spline_LC,
  ER_synch_spline_LC,
GPP_synch_spline_BC,
ER_synch_spline_BC,
  ncol = 1, nrow = 4,
  common.legend = F, 
  legend = "right",
  align = "v")

spline_grid_landscape
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/metab_LC_GL_BC_splines.png", plot = spline_grid_landscape, width = 4.75, height = 9.5, units = "in")

```
