---
title: "Part 1 & 2 : Spline Correlogram analysis for synchrony in aquatic ecosystem productivity - Temporal scales"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables
library(broom)
library(grid)

```

## Background: Synchrony estimates using Sncf{ncf}

#### How it works:

Sncf computes the nonparametric spatial (auto/cross-)correlation
function for multivariate time series data observed across spatial
locations.

1.  Takes time series data z from n locations (one row per location, one
    column per time step).

2.  Calculates the pairwise correlations between time series at each
    location pair.

3.  Bins those correlations based on distance between sites (using
    latitude and longitude if latlon = TRUE).

4.  Fits a smoothing spline to estimate the average correlation as a
    function of distance
    
It can handle missing data, where:
na.rm = TRUE: Missing values are allowed; only overlapping time steps are used in correlations.


#### Metrics to consider:

-   '\real$predicted\$y[1]' = Synchrony at (almost) zero distance –
    often max synchrony.
      
      - This is the first predicted value from the fitted nonparametric spline of the spatial correlation function
      
      - Estimated synchrony at the smallest distance bin


-   '\$y.intercept' = Extrapolated synchrony at exactly zero distance 
      
      - The theoretical synchrony among perfectly co-located sites

-   '\$cbar' = Regional mean synchrony across all distances

-   '\$x.intercept'	= Distance at which synchrony drops to 0

-   '\$e.intercept' = Distance where synchrony drops to 1/e of max (decay scale)


###  Fletcher, R., & Fortin, M. (2018):
we should truncate the range of lag distances at which we consider spatial 
dependence to approximately ½ to 2 = 3 the total distance observed.

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

river_metab_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/stream_spline_cor_df.rds") %>% mutate(GPP_filled = if_else(GPP_filled <= 0, 0.0001, GPP_filled),
          ER_filled = if_else(ER_filled >= 0, -0.0001, ER_filled))

names(river_metab_dat)
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
### read in prism data for PPT

PPT_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Clim_data/PRISM_stream.rds")%>%
  mutate(Site_ID=as.character(Name))


StreamLight_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/7_river_light.csv") %>%
  mutate(date = lubridate::as_date(date, format="%Y-%m-%d")) %>%
  mutate(Site_ID1 = str_remove(site, "^nwis_")) %>%
  mutate(Site_ID = str_remove(Site_ID1, "^0+"))


river_metab_dat1 <- river_metab_dat%>%
  left_join(PPT_dat, by=c("date"="Date", "Site_ID")) %>%
  dplyr::select(site:TmaxC) 

river_metab_dat1 <- river_metab_dat1%>%
  left_join(StreamLight_dat, by=c("date", "site", "Site_ID"))

intersect(unique(river_metab_dat1$site), unique(StreamLight_dat$site))


setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab



sites_keep <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Map_data_for_nested_reach.csv")%>%
  filter(keep == 1) %>%
  dplyr::select(Site_ID:stream_order)


# Filter river_metab_dat to keep only Site_IDs in sites_keep
river_metab_dat1 <- river_metab_dat1 %>%
  filter(Site_ID %in% sites_keep$Site_ID)

unique(river_metab_dat1$Site_ID)
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
library(dplyr)
library(geodist)

# Get site centroids
site_coords <- river_metab_dat1 %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Calculate the overall mean coordinates
mean_lat <- median(site_coords$Lat)
mean_lon <- median(site_coords$Lon)

# Find site closest to mean coordinates using geodist
mean_coords <- data.frame(Lat = mean_lat, Lon = mean_lon)
site_coords_df <- data.frame(Lat = site_coords$Lat, Lon = site_coords$Lon)

distances_to_mean <- geodist(site_coords_df, mean_coords, measure = "geodesic") / 1000
closest_index <- which.min(distances_to_mean)

# Use the closest site as the reference
ref_site <- site_coords[closest_index, ]

# Compute distances from the reference site to all others
ref_coords <- data.frame(Lat = ref_site$Lat, Lon = ref_site$Lon)
distances_from_ref <- geodist(site_coords_df, ref_coords, measure = "geodesic") / 1000

# Add distance column
site_coords <- site_coords %>%
  mutate(distance_km = as.numeric(distances_from_ref))

hist(site_coords$distance_km)


# Get 2/3 cut off 

half_dist_km <- max(site_coords$distance_km)/ 2
half_dist_km
twothirds_dist_km <- max(site_coords$distance_km) * (2/3)
twothirds_dist_km

```

## Part 1: Synchrony in metabolism at differnt timescales

Compare synchrony for GPP and ER from 56 different sites in Marzolf
dataset at different timescales:

-   Daily

-   Weekly (mean)

-   Monthly (mean)

-   Quarterly (3 month mean)

-   Yearly (mean)

### A. GPP synchrony at differnt timescales:

#### GPP- Daily synchrony

```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}

#### Daily synchrony

river_metab_dat<-as.data.frame(river_metab_dat1)

GPP_long <- river_metab_dat1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day <- GPP_long %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_day <- Sncf(x=GPP_wide_day$Lon,
                   y=GPP_wide_day$Lat,
                   z=GPP_wide_day[,4:ncol(GPP_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year

plot(RR_df.corr_day)

GPP_corrgram_summary.table <- summary(RR_df.corr_day)
Regional.synch <- GPP_corrgram_summary.table$Regional.synch


# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP")

## Sample average line
GPP.corr.pred.df.day$mean_synch <- round(Regional.synch,2)
Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.day$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP")



```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### GPP- Weekly synchrony

```{r, warning=F, size = 'small', echo=F, include=F, cache=TRUE}
#### Weekly synchrony

GPP_week <-  river_metab_dat1 %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))

GPP_wide_week <- GPP_week %>%
  dplyr::select(Site_ID, Lon, Lat, week, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }

GPP_wide_week_1 <-na.omit(GPP_wide_week)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=GPP_wide_week$Lon,
                   y=GPP_wide_week$Lat,
                   z=GPP_wide_week[,4:ncol(GPP_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)


GPP_corrgram_summary.table <- summary(RR_df.corr_wk)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly",
                                     parameter = "GPP")

## Sample average line

Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.df.wk$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.wk$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly",
                                     parameter = "GPP")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### GPP- Monthly synchrony

```{r, warning=F, size = 'small', echo=F, include=F, cache=TRUE}

GPP_month <-  river_metab_dat1 %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm=T))


GPP_wide_month <- GPP_month %>%
 # dplyr::filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=GPP_wide_month$Lon,
                   y=GPP_wide_month$Lat,
                   z=GPP_wide_month[,4:ncol(GPP_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


GPP_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly",
                                   parameter = "GPP")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.df.mon$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.mon$cbar <- Regional.synch_x


GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                time_scale = "monthly",
                                   parameter = "GPP")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### GPP - Quarterly synch

```{r, warning=F, size = 'small', echo=F, include=F}
#### Quarterly synch

GPP_qrt <- river_metab_dat1 %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_qrt <- GPP_qrt %>%
 # dplyr::filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, GPP_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = GPP_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(GPP_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=GPP_wide_qrt$Lon,
                   y=GPP_wide_qrt$Lat,
                   z=GPP_wide_qrt[,4:ncol(GPP_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)

GPP_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly",   
                                parameter = "GPP")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.df.qrtly$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.qrtly$cbar <- Regional.synch_x


GPP_corrgram_summary.table
# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly",    
                               parameter = "GPP")


```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### GPP - Yearly synchrony

```{r, warning=F, size = 'small', echo=F, include=F, cache=TRUE}

GPP_yr <- river_metab_dat1 %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")


GPP_wide_yr <- GPP_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, GPP_filled) %>%
  pivot_wider(names_from = c(year), values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=GPP_wide_yr$Lon,
                   y=GPP_wide_yr$Lat,
                   z=GPP_wide_yr[,4:ncol(GPP_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

GPP_corrgram_summary.table <- summary(RR_df.corr_yr)
GPP_corrgram_summary.table
# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly",
                                  parameter = "GPP")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.df.yr$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.yr$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly",
                                     parameter = "GPP")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
##### combined synchonry df 
pred_df_all <- bind_rows(GPP.corr.pred.df.day, GPP.corr.pred.df.wk, GPP.corr.pred.df.mon, GPP.corr.pred.df.qrtly, GPP.corr.pred.df.yr)
ci_df_all   <- bind_rows(GPP.corr.boot.df.day, GPP.corr.boot.df.wk, GPP.corr.boot.df.mon, GPP.corr.boot.df.qrtly, GPP.corr.boot.df.yr)

```

###### Calculate the percent difference in synchrony between time scales from daily

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each time_scale becomes a column
pred_wide <- pred_df_all %>%
  dplyr::select(x, y, time_scale) %>%
  pivot_wider(names_from = time_scale, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff <- pred_wide %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )

pred_diff
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
scale_colors <- c(
  "daily" = "#E69F00",     
  "quarterly" = "#56B4E9",    
  "monthly" = "#009E73",   
  "weekly" = "#F0E442", 
  "yearly" = "#0072B2"    
)
```

###### Precent difference in GPP snych signal from daily 

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Pivot longer
pred_long_diff <- pred_diff %>%
  dplyr::select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )

# Plot histogram
ggplot(pred_long_diff, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```


###### GPP: Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances stream GPP is no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all %>%
  group_by(time_scale) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```

### Figure 2 (a): Spline Correlogram of GPP at different timescales

Points represent sites with metabolism data 

```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(grid)
library(ggplot2)
library(patchwork)

### add point for mean GPP to plot 
library(geosphere)
# Get site centroids
site_coords <- GPP_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (the first one in the dataset)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

GPP_long_dist <- GPP_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")


GPP_long_dist_plt <- GPP_long_dist %>%
  mutate(GPP_P = ifelse(!is.na(GPP_filled), -0.2, 0)) %>%
  group_by(Site_ID, distance_km, GPP_P) %>%
  summarise(
    GPP_mean = median(GPP_filled, na.rm = TRUE), 
    GPP_low = quantile(GPP_filled, probs = 0.25, na.rm = TRUE), 
    GPP_up = quantile(GPP_filled, probs = 0.75, na.rm = TRUE),
    .groups = "drop"
  )

###
GPP_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all%>%dplyr::filter(x<3150),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all%>%dplyr::filter(x<3150),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = time_scale),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = time_scale),
               linetype = "dashed") +
  geom_vline(xintercept = twothirds_dist_km, color="grey40") +
  # Add GPP points (mean)
  geom_point(data = GPP_long_dist_plt,
             aes(x = distance_km, y = GPP_P),
             shape = 4, size = 2,
             alpha=0.5) +
    ylim(-0.2, 0.75) +
  # Labels and theme
  labs(title = "Synchrony in GPP across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff%>%dplyr::filter(x<3150), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic() +
theme(
  axis.title.x = element_blank(),
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  plot.margin = ggplot2::margin(5, 5, 5, 5, unit = "pt"),
legend.position = "none"
)


# Combine using patchwork (put rotated density plot on the right)
combined_plot <- GPP_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/GPP_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```


###### Rate of GPP synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all %>%
  group_by(time_scale) %>%
    filter(x<cbar) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(time_scale, term,rounded_estimate: rounded_p.value)

fit_models_tab


```

### B. ER synchrony at differnt timescales:

#### ER - Daily synchrony:

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8, cache=TRUE}
ER_long <- river_metab_dat1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day <- ER_long %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
ER_df.corr_day <- Sncf(x=ER_wide_day$Lon,
                   y=ER_wide_day$Lat,
                   z=ER_wide_day[,4:ncol(ER_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(ER_df.corr_day)


ER_corrgram_summary.table <- summary(ER_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(ER_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                                  parameter = "ER")


## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.df.day$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.df.day$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(ER_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.day <- data.frame(x = matrix(unlist(ER_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "daily",     
                               parameter = "ER")


```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### ER - Weekly synchrony

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8, cache=TRUE}
ER_week <-  river_metab_dat1 %>%
  mutate(week=week(date))%>%
  group_by(Site_ID, week, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

ER_wide_week <- ER_week %>%
  dplyr::select(Site_ID,Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled)

river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_week <- ER_week %>%
 # dplyr::filter(year>2011) %>%
 dplyr::select(Site_ID, Lon, Lat, week, year, ER_filled) %>%
  pivot_wider(names_from = c(year, week), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_wk <- Sncf(x=ER_wide_week$Lon,
                   y=ER_wide_week$Lat,
                   z=ER_wide_week[,4:ncol(ER_wide_week)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm=T)


plot(RR_df.corr_wk)


ER_corrgram_summary.table <- summary(RR_df.corr_wk)
# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_wk$real$predicted$y)),
                                     time_scale = "weekly",     
                                 parameter = "ER")
## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.df.wk$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.df.wk$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_wk$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.wk <- data.frame(x = matrix(unlist(RR_df.corr_wk$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                     time_scale = "weekly",
                                     parameter = "ER")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```

#### ER - Monthly synchrony

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8, cache=TRUE}

ER_month <-  river_metab_dat1 %>%
  mutate(month=month(date))%>%
  group_by(Site_ID, month, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm=T))

river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_month <- ER_month %>%
 # dplyr::filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, month, year, ER_filled) %>%
  pivot_wider(names_from = c(year, month), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_month)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_mon <- Sncf(x=ER_wide_month$Lon,
                   y=ER_wide_month$Lat,
                   z=ER_wide_month[,4:ncol(ER_wide_month)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_mon)


ER_corrgram_summary.table <- summary(RR_df.corr_mon)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_mon$real$predicted$y)),
                                time_scale = "monthly",
                                    parameter = "ER")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.df.mon$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.df.mon$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_mon$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.mon <- data.frame(x = matrix(unlist(RR_df.corr_mon$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                time_scale = "monthly",
                                    parameter = "ER")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### ER - Quarterly synchrony

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8, cache=TRUE}

ER_qrt <- river_metab_dat1 %>%
  mutate(
    month = month(date),
    year = year(date),
    quarter = case_when(
      month %in% 1:3 ~ "Q1",
      month %in% 4:6 ~ "Q2",
      month %in% 7:9 ~ "Q3",
      month %in% 10:12 ~ "Q4"
    )
  ) %>%
  group_by(Site_ID, quarter, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


river_metab_dat<- as.data.frame(river_metab_dat)

ER_wide_qrt <- ER_qrt %>%
 # dplyr::filter(year>2011) %>%
  dplyr::select(Site_ID, Lon, Lat, quarter, year, ER_filled) %>%
  pivot_wider(names_from = c(year, quarter), values_from = ER_filled) %>%
  # Reorder columns: keep Site_ID, Lon, Lat first, then order year_week columns
  {
    static_cols <- c("Site_ID", "Lon", "Lat")
    dynamic_cols <- setdiff(names(.), static_cols)
    
    # Extract year and week and sort
    sorted_dynamic_cols <- dynamic_cols[order(
      as.integer(sub("^(\\d+)_\\d+$", "\\1", dynamic_cols)),   # year part
      as.integer(sub("^\\d+_(\\d+)$", "\\1", dynamic_cols))    # week part
    )]
    
    dplyr::select(., all_of(static_cols), all_of(sorted_dynamic_cols))
  }


str(ER_wide_qrt)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_qrt <- Sncf(x=ER_wide_qrt$Lon,
                   y=ER_wide_qrt$Lat,
                   z=ER_wide_qrt[,4:ncol(ER_wide_qrt)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_qrt)


ER_corrgram_summary.table <- summary(RR_df.corr_qrt)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_qrt$real$predicted$y)),
                                time_scale = "quarterly",     
                                parameter = "ER")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.df.qrtly$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.df.qrtly$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_qrt$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.qrtly <- data.frame(x = matrix(unlist(RR_df.corr_qrt$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                               time_scale = "quarterly",     
                               parameter = "ER")
```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### ER - Yearly synchrony

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8, cache=TRUE}

#### Yearly synchrony
ER_yr <- river_metab_dat1 %>%
  mutate(
    year = year(date)) %>%
  group_by(Site_ID, year, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")


ER_wide_yr <- ER_yr %>%
  dplyr::select(Site_ID,Lon, Lat, year, ER_filled) %>%
  pivot_wider(names_from = c(year), values_from = ER_filled)


# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_yr <- Sncf(x=ER_wide_yr$Lon,
                   y=ER_wide_yr$Lat,
                   z=ER_wide_yr[,4:ncol(ER_wide_yr)],
                   resamp = 1000, 
                   latlon = T, 
                   na.rm=T)
plot(RR_df.corr_yr)

ER_corrgram_summary.table <- summary(RR_df.corr_yr)

# create a dataframe of the predicted values
ER.corr.pred.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_yr$real$predicted$y)),
                               time_scale = "yearly",
                                   parameter = "ER")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.df.yr$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.df.yr$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.df <- as.data.frame(t(RR_df.corr_yr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.df.yr <- data.frame(x = matrix(unlist(RR_df.corr_yr$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.df$`0.025`,
                                             ymaxER = ER.corr.bootValues.df$`0.975`,
                                  time_scale = "yearly",    
                                 parameter = "ER")

```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
##### combined synch 

pred_df_all_ER <- bind_rows(ER.corr.pred.df.day, ER.corr.pred.df.wk, ER.corr.pred.df.mon, ER.corr.pred.df.qrtly, ER.corr.pred.df.yr)
ci_df_all_ER   <- bind_rows(ER.corr.boot.df.day, ER.corr.boot.df.wk, ER.corr.boot.df.mon, ER.corr.boot.df.qrtly, ER.corr.boot.df.yr)

```

###### ER: Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances stream ER is no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_ER %>%
  group_by(time_scale) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


###### Rate of ER synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}
library(dplyr)
library(purrr)
library(broom)


cbar_df <- pred_df_all_ER %>%
  group_by(time_scale) %>%
  summarise(cbar = unique(cbar))  # or max, mean, etc. depending on your case

fit_models <- pred_df_all_ER %>%
  group_by(time_scale) %>%
  filter(x<cbar) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(time_scale, term,rounded_estimate: rounded_p.value)

fit_models_tab


```

```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

ggplot() +
  geom_ribbon(data = ci_df_all_ER,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  geom_line(data = pred_df_all_ER,
            aes(x = x, y = y, color = time_scale),
            size = 1) +
  ##
    labs(title = "Synchrony in ER across different time scales", 
       y= "Correlation", 
       x="Distance (km)") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic()


```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=4}


# Spread the data so each time_scale becomes a column
pred_wide_ER <- pred_df_all_ER %>%
  dplyr::select(x,y,time_scale)%>%
  pivot_wider(names_from = time_scale, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_ER <- pred_wide_ER %>%
  mutate(
    weekly_diff = (abs(weekly - daily) / daily) * 100,
    monthly_diff = (abs(monthly - daily) / daily) * 100,
    quarterly_diff = (abs(quarterly - daily) / daily) * 100,
    yearly_diff = (abs(yearly - daily) / daily) * 100
  )



# Pivot longer
pred_long_diff_ER <- pred_diff_ER %>%
 dplyr:: select(x, ends_with("_diff")) %>%
  pivot_longer(
    cols = ends_with("_diff"),
    names_to = "time_scale",
    values_to = "percent_diff"
  ) %>%
  mutate(
    time_scale = gsub("_diff", "", time_scale)
  )


# Plot histogram
ggplot(pred_long_diff_ER, aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = scale_colors) +
  labs(
    title = "Percent Difference from Daily by Time Scale",
    x = "Percent Difference from Daily (%)",
    y = "Count",
    fill = "Time Scale"
  ) +
  theme_minimal()

```

### Figure 2 (b): Spline Correlogram of ER at different timescales

Points represent site with metabolism estimates


```{r, echo=FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)

library(broom)
### add point for mean ER to plot 
library(geosphere)

site_coords <- ER_long %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Choose a reference site (e.g., the upstream-most one)
ref_site <- site_coords[1, ]

# Compute distance in km from reference site
site_coords <- site_coords %>%
  mutate(distance_km = distHaversine(matrix(c(Lon, Lat), ncol = 2),
                                     matrix(c(ref_site$Lon, ref_site$Lat), ncol = 2)) / 1000)

ER_long_dist <- ER_long %>%
  left_join(site_coords %>% dplyr::select(Site_ID, distance_km), by = "Site_ID")




ER_long_dist_plt <- ER_long_dist %>%
  mutate(ER_P = ifelse(!is.na(ER_filled), -0.2, 0)) %>%
  group_by(Site_ID, distance_km, ER_P) %>%
  summarise(
    ER_mean = median(ER_filled, na.rm = TRUE), 
    ER_low = quantile(ER_filled, probs = 0.25, na.rm = TRUE), 
    ER_up = quantile(ER_filled, probs = 0.75, na.rm = TRUE),
    .groups = "drop"
  )


###
ER_synch_spline <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_ER%>%dplyr::filter(x<3150),
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = time_scale),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_ER%>%dplyr::filter(x<3150),
            aes(x = x, y = y, color = time_scale),
            size = 1) +
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = time_scale),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = time_scale),
               linetype = "dashed") +
  # Add ER points (mean)
  geom_point(data = ER_long_dist_plt,
             aes(x = distance_km, y = ER_P),
             shape = 4, size = 2,
             alpha=0.5) +
    geom_vline(xintercept = twothirds_dist_km, color="grey40") +
  ylim(-0.2,0.75) +
  # Labels and theme
  labs(title = "Synchrony in ER across different time scales",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Time Scale",
       color = "Time Scale") +
  scale_color_manual(values = scale_colors) +
  scale_fill_manual(values = scale_colors) +
  theme_classic() +
  theme(legend.position = "bottom")

# Rotated density plot (marginal plot for y-axis)
density_plot_y <- ggplot(pred_long_diff_ER%>%dplyr::filter(x<3150), aes(x = percent_diff, fill = time_scale)) +
  geom_density(position = "identity", alpha = 0.6) +
  coord_flip() +  # Rotate 90 degrees
  scale_fill_manual(values = scale_colors) +
  labs(
    x = "Percent difference from daily synchrony",
    y = "Density") +
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
                plot.margin = ggplot2::margin(5, 5, 5, 5, unit = "pt"),,
        legend.position="none")

# Combine using patchwork (put rotated density plot on the right)
combined_plot <- ER_synch_spline + density_plot_y +
  plot_layout(ncol = 2, widths = c(3, 1))
combined_plot
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/ER_spline_correlogram.png", plot = combined_plot, width =6.5, height = 4.25, units = "in")

```


## FIG 2


```{r, warning=F, size = 'small', echo=F,include=TRUE, fig.width=7, fig.height=7}
# Arrange plots with shared legend at the bottom
library(ggpubr)

spline_grid_combo <- ggarrange(
GPP_synch_spline,
ER_synch_spline,
  ncol = 1, nrow = 2,
  common.legend = T, 
  legend = "bottom"
)

spline_grid_combo
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Timeseiries_spline_correlogram_v2.png", plot = spline_grid_combo, width =5.5, height = 6.5, units = "in")

```



## Part 2 - Moran Effect

-   MODIS data for leaf area index (LAI)

-   Daily streamflow (Q)

-   Daily water temperature (water_temp)

-   Daily air temp (air_temp)

-   Daily precipitation (PPT)

#### Bring in MODIS data for LAI to get at

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

LAI_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Modis_data/StreamDat_MOD_combined.rds")%>%
  mutate(date=as.Date(Date)) %>%
    filter(Site_ID %in% sites_keep$Site_ID)


# intersect(unique(river_metab_dat1$Site_ID), unique(LAI_dat$Site_ID))
# 
# 
# Metat_sites_notinlight <- setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
# Light_sites_notinmetab<- setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab
# 
# LAI_sites_notinlight<- setdiff(unique(LAI_dat$Site_ID), unique(StreamLight_dat$Site_ID)) # Sites in light but not in metab

```

#### LAI Daily synch (filled)

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}

str(LAI_dat)

stream_LAI_dat <- LAI_dat %>%
  left_join(river_metab_dat, by = c("Site_ID", "date"))

### 

LAI_long <- stream_LAI_dat %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(Lai = mean(Lai, na.rm = TRUE), .groups = "drop")


LAI_filled_coords <- LAI_long %>%
  group_by(Site_ID) %>%
  fill(Lat, Lon, .direction = "downup") %>%
  ungroup()


LAI_filled_coords1 <- LAI_filled_coords %>%
  group_by(Site_ID) %>%
  fill(Lai, .direction = "downup") %>%
  ungroup()

LAI_wide_day <- LAI_filled_coords1 %>%
  pivot_wider(
    names_from = date,
    values_from = Lai)

LAI_wide_day<- na.omit(LAI_wide_day)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
LAI_df.corr_day <- Sncf(x=LAI_wide_day$Lon,
                   y=LAI_wide_day$Lat,
                   z=LAI_wide_day[,4:ncol(LAI_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T)


plot(LAI_df.corr_day)


LAI_corrgram_summary.table <- summary(LAI_df.corr_day)

# Plot using ggplot for export
# create a dataframe of the predicted values
LAI.corr.pred.df.day <- data.frame(x = matrix(unlist(LAI_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(LAI_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter="LAI")

## Sample average line
Regional.synch <- LAI_corrgram_summary.table$Regional.synch
LAI.corr.pred.df.day$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(LAI_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
LAI.corr.pred.df.day$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
LAI.corr.bootValues.df <- as.data.frame(t(LAI_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
LAI.corr.boot.df.day <- data.frame(x = matrix(unlist(LAI_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminLAI = LAI.corr.bootValues.df$`0.025`,
                                             ymaxLAI = LAI.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                parameter="LAI")

```

```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(LAI_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(LAI_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(LAI_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(LAI_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily flow:

```{r, size='small', echo=F, include=T, warning=FALSE}
# choosing max flow:
RR_dat_TS_Q <- river_metab_dat1 %>%
  dplyr::select(Site_ID,Lon, Lat, date, discharge) %>%
  pivot_wider(names_from = c(date), values_from = discharge)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
Q_RR_df.corr <- Sncf(x=RR_dat_TS_Q$Lon,
                   y=RR_dat_TS_Q$Lat,
                   z=RR_dat_TS_Q[,4:ncol(RR_dat_TS_Q)],
                   resamp = 1000,
                   latlon = T,
                     na.rm = T)

plot(Q_RR_df.corr)

Q_corrgram_summary.table <- summary(Q_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Q.corr.pred.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$real$predicted$x)),
                               y = matrix(unlist(Q_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="Q")

## Sample average line
Regional.synch <- Q_corrgram_summary.table$Regional.synch
Q.corr.pred.df$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(Q_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
Q.corr.pred.df$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
Q.corr.bootValues.df <- as.data.frame(t(Q_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Q.corr.boot.df <- data.frame(x = matrix(unlist(Q_RR_df.corr$boot$boot.summary$predicted$x)),
                               yminQ = Q.corr.bootValues.df$`0.025`,
                               ymaxQ = Q.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="Q")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(Q_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(Q_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(Q_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(Q_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily water temperature:

```{r, size='small', echo=F, include=T, warning=FALSE}

### Water temperature
RR_dat_TS_temp <- river_metab_dat1 %>%
  dplyr::select(Site_ID, Lon, Lat, date, temp.water) %>%
  pivot_wider(names_from = c(date), values_from = temp.water)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
temp_RR_df.corr <- Sncf(x=RR_dat_TS_temp$Lon,
                     y=RR_dat_TS_temp$Lat,
                     z=RR_dat_TS_temp[,4:ncol(RR_dat_TS_temp)],
                     resamp = 1000,
                     latlon = T,
                     na.rm = T)

plot(temp_RR_df.corr)
temp_corrgram_summary.table <- summary(temp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
temp.corr.pred.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$real$predicted$x)),
                             y = matrix(unlist(temp_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="water_temp")

## Sample average line
Regional.synch <- temp_corrgram_summary.table$Regional.synch
temp.corr.pred.df$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(temp_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
temp.corr.pred.df$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
temp.corr.bootValues.df <- as.data.frame(t(temp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
temp.corr.boot.df <- data.frame(x = matrix(unlist(temp_RR_df.corr$boot$boot.summary$predicted$x)),
                             yminWtemp = temp.corr.bootValues.df$`0.025`,
                             ymaxWtemp = temp.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="water_temp")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(temp_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(temp_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(temp_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(temp_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily air temperature:

```{r, size='small', echo=T, include=T, warning=FALSE}

### Water temperature
RR_dat_TS_Atemp <- river_metab_dat1 %>%
  dplyr::select(Site_ID, Lon, Lat, date, TmeanC) %>%
  pivot_wider(names_from = c(date), values_from = TmeanC)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
Atemp_RR_df.corr <- Sncf(x=RR_dat_TS_Atemp$Lon,
                     y=RR_dat_TS_Atemp$Lat,
                     z=RR_dat_TS_Atemp[,4:ncol(RR_dat_TS_Atemp)],
                     resamp = 1000,
                     latlon = T,
                     na.rm = T)

plot(Atemp_RR_df.corr)
Atemp_corrgram_summary.table <- summary(Atemp_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
Atemp.corr.pred.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$real$predicted$x)),
                             y = matrix(unlist(Atemp_RR_df.corr$real$predicted$y)),
                             time_scale = "daily",
                             parameter="air_temp")

## Sample average line
Regional.synch <- Atemp_corrgram_summary.table$Regional.synch
Atemp.corr.pred.df$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(Atemp_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
Atemp.corr.pred.df$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
Atemp.corr.bootValues.df <- as.data.frame(t(Atemp_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
Atemp.corr.boot.df <- data.frame(x = matrix(unlist(Atemp_RR_df.corr$boot$boot.summary$predicted$x)),
                             yminAtemp = Atemp.corr.bootValues.df$`0.025`,
                             ymaxAtemp = Atemp.corr.bootValues.df$`0.975`,
                             time_scale = "daily",
                             parameter="air_temp")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(Atemp_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(Atemp_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(Atemp_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(Atemp_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily PPT:

```{r, size='small', echo=T, include=T, warning=FALSE}

RR_dat_TS_ppt <- river_metab_dat1 %>%
  dplyr::select(Site_ID, Lon, Lat, date, ppt_mm) %>%
  pivot_wider(names_from = c(date), values_from = ppt_mm)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
ppt_RR_df.corr <- Sncf(x=RR_dat_TS_ppt$Lon,
                        y=RR_dat_TS_ppt$Lat,
                        z=RR_dat_TS_ppt[,4:ncol(RR_dat_TS_ppt)],
                        resamp = 1000,
                        latlon = T,
                       na.rm = T)

plot(ppt_RR_df.corr)

ppt_corrgram_summary.table <- summary(ppt_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
ppt.corr.pred.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$real$predicted$x)),
                                y = matrix(unlist(ppt_RR_df.corr$real$predicted$y)),
                            time_scale = "daily",
                             parameter="PPT")

## Sample average line
Regional.synch <- ppt_corrgram_summary.table$Regional.synch
ppt.corr.pred.df$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ppt_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ppt.corr.pred.df$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
ppt.corr.bootValues.df <- as.data.frame(t(ppt_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ppt.corr.boot.df <- data.frame(x = matrix(unlist(ppt_RR_df.corr$boot$boot.summary$predicted$x)),
                                yminPPT = ppt.corr.bootValues.df$`0.025`,
                                ymaxPPT = ppt.corr.bootValues.df$`0.975`,
                                    time_scale = "daily",
                             parameter="PPT")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ppt_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ppt_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ppt_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ppt_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```



#### Light :

```{r, size='small', echo=T, include=T, warning=FALSE}

RR_dat_TS_light <- river_metab_dat1 %>%
  dplyr::select(Site_ID, Lon, Lat, date, daily_PAR) %>%
  pivot_wider(names_from = c(date), values_from = daily_PAR)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
light_RR_df.corr <- Sncf(x=RR_dat_TS_light$Lon,
                        y=RR_dat_TS_light$Lat,
                        z=RR_dat_TS_light[,4:ncol(RR_dat_TS_light)],
                        resamp = 1000,
                        latlon = T,
                       na.rm = T)

plot(light_RR_df.corr)

light_corrgram_summary.table <- summary(light_RR_df.corr)

# Plot using ggplot for export
# create a dataframe of the predicted values
light.corr.pred.df <- data.frame(x = matrix(unlist(light_RR_df.corr$real$predicted$x)),
                                y = matrix(unlist(light_RR_df.corr$real$predicted$y)),
                            time_scale = "daily",
                             parameter="Light")

## Sample average line
Regional.synch <- light_corrgram_summary.table$Regional.synch
light.corr.pred.df$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(light_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
light.corr.pred.df$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
light.corr.bootValues.df <- as.data.frame(t(light_RR_df.corr$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
light.corr.boot.df <- data.frame(x = matrix(unlist(light_RR_df.corr$boot$boot.summary$predicted$x)),
                                yminlight = light.corr.bootValues.df$`0.025`,
                                ymaxlight = light.corr.bootValues.df$`0.975`,
                                    time_scale = "daily",
                             parameter="Light")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(light_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(light_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(light_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(light_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, size='small', echo=F, include=F, warning=FALSE}

big_boot_df<- GPP.corr.boot.df.day %>%
    left_join(ER.corr.boot.df.day, by=c("x")) %>%
  left_join(Q.corr.boot.df, by=c("x")) %>%
  left_join(temp.corr.boot.df, by=c("x")) %>%
  left_join(ppt.corr.boot.df, by=c("x")) %>%
  left_join(Atemp.corr.boot.df, by=c("x")) %>%
   left_join(light.corr.boot.df, by=c("x"))

# Create a vector of unique x values from big_boot_df
big_x_vals <- unique(big_boot_df$x)

# Add a column to LAI.corr.boot.df.day with the nearest x from big_boot_df
LAI.corr.boot.df.day <- LAI.corr.boot.df.day %>%
  mutate(x_nearest = map_dbl(x, ~big_x_vals[which.min(abs(big_x_vals - .x))]))

# light.corr.boot.df.day <- light.corr.boot.df %>%
#   mutate(x_nearest = map_dbl(x, ~big_x_vals[which.min(abs(big_x_vals - .x))]))

# Now join with big_boot_df using the nearest x value
joined_df <- LAI.corr.boot.df.day %>%
  left_join(big_boot_df, by = c("x_nearest" = "x"))

# joined_df <- light.corr.boot.df.day %>%
#   left_join(big_boot_df, by = c("x_nearest" = "x"))

```

## Figure 3: correlogram of Synchrony of GPP + other hydro/climate/environmental parameters near streams

```{r, size = 'large', echo=FALSE, include=F, warning=FALSE}

pred_df <- bind_rows(
  Q.corr.pred.df %>% mutate(group = "Flow"),
  GPP.corr.pred.df.day %>% mutate(group = "GPP"),
    ER.corr.pred.df.day %>% mutate(group = "ER"),
  temp.corr.pred.df %>% mutate(group = "Water Temp"),
  ppt.corr.pred.df %>% mutate(group = "PPT"),
  Atemp.corr.pred.df %>% mutate(group = "Air Temp"),
 LAI.corr.pred.df.day %>% mutate(group = "LAI"),
  light.corr.pred.df%>% mutate(group = "Light")
)

```


###### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df %>%
  group_by(group) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


```{r, size = 'large', echo=FALSE, include=T, warning=FALSE}
ribbon_df <- joined_df %>%
  transmute(
    x = x,
    Flow_ymin = yminQ, Flow_ymax = ymaxQ,
    GPP_ymin = yminGPP, GPP_ymax = ymaxGPP,
    ER_ymin = yminER, ER_ymax = ymaxER,
    WaterTemp_ymin = yminWtemp, WaterTemp_ymax = ymaxWtemp,
    PPT_ymin = yminPPT, PPT_ymax = ymaxPPT,
    AirTemp_ymin = yminAtemp, AirTemp_ymax = ymaxAtemp,
    yminLAI = yminLAI, ymaxLAI = ymaxLAI
  ) %>%
  pivot_longer(cols = -x, names_to = c("group", ".value"), names_sep = "_")

synch.plot <- ggplot() +
  geom_line(data = pred_df%>%dplyr::filter(x<twothirds_dist_km),
            aes(x = x, y = y, color = group)) +
     geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = group),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = group),
               linetype = "dashed") +
  theme_classic() +
  scale_color_manual(values = c(
    "Flow" = "#0072B2",
    "GPP" = "#1aab1a",
    "ER" = "#f5880c",
    "Water Temp" = "#ed1111",
    "PPT" = "#56B4E9",
    "Air Temp" = "#de21bb",
    "Light" = "#E2D523", 
    "LAI"= "#8f21de"
  )) +
   geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminQ, ymax = ymaxQ),
              fill = "#0072B2", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
              fill = "#1aab1a", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminWtemp, ymax = ymaxWtemp),
              fill = "#ed1111", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminPPT, ymax = ymaxPPT),
              fill = "#56B4E9", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminAtemp, ymax = ymaxAtemp),
              fill = "#de21bb", alpha=0.2)+
      geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminlight, ymax = ymaxlight),
              fill = "#E2D523", alpha=0.2) +
      geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminER, ymax = ymaxER),
              fill = "#f5880c", alpha=0.2) +
    geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminLAI, ymax = ymaxLAI),
              fill = "#8f21de", alpha=0.2) +
    labs(title = "Synchrony of metabolism and\nenvironmental conditions",
              y= "Correlation",
       x="Distance (km)",
       color = "Variable", fill = "Variable") +
  theme(text = element_text(size=14),
        legend.position = "bottom") 

synch.plot

```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Moran_spline_correlogram.png", plot = synch.plot, width =5.5, height = 4.5, units = "in")

```

###### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}
library(dplyr)
library(purrr)
library(broom)


fit_models <- pred_df%>%dplyr::filter(x<2500) %>% # adjust to x at sample dist
  group_by(parameter) %>%
  filter(x<cbar) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params)

fit_models

```

| Variable | a (initial synch) | b (decay rate) | Interpretation |
|-----------|----------:|----------:|-----------------------------------------|
| **Q (Flow)** | 0.6937 | 0.00223 | Starts moderately synchronized, but decays rapidly over space. |
| **GPP** | 0.2597 | 0.000037 | Low synchrony, and it barely decays — very stable across space, suggesting local factors |
| **Water Temp** | 0.9361 | 0.0000285 | High synchrony, — very stable across space |
| **PPT (Precip)** | 0.9009 | 0.00277 | High synchrony, but decays quickly — similar to flow. Tight confidence intervals. |
| **Air Temp** | 0.9853 | 0.000103 | The most synchronous slowly decaying — regional coherence in air temp. Tight confidence intervals |
| **LAI** | 0.6879 | 0.000153 | Moderate synchrony and mild decay — stable across space, suggesting local factors |


###### Main take-aways from Fig 3

A. Looks like GPP has the lowest synchrony, showing it is the most
influenced by localized differences.

B. GPP becomes asynchronous at much smaller distances (797 km shorter
\~64% than LAI), and at faster rates (64% delta b: 0.0004).



### defense presentation photo:


## Figure 3: correlogram of Synchrony of GPP + other hydro/climate/environmental parameters near streams

```{r, size = 'large', echo=FALSE, include=F, warning=FALSE}

pred_df <- bind_rows(
  Q.corr.pred.df %>% mutate(group = "Flow"),
  GPP.corr.pred.df.day %>% mutate(group = "GPP"),
    ER.corr.pred.df.day %>% mutate(group = "ER"),
  temp.corr.pred.df %>% mutate(group = "Water Temp"),
  ppt.corr.pred.df %>% mutate(group = "PPT"),
  Atemp.corr.pred.df %>% mutate(group = "Air Temp"),
 LAI.corr.pred.df.day %>% mutate(group = "LAI"),
  light.corr.pred.df%>% mutate(group = "Light")
)

```


###### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df %>%
  group_by(group) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))


intersection_df
```


```{r, size = 'large', echo=FALSE, include=T, warning=FALSE}
ribbon_df <- joined_df %>%
  transmute(
    x = x,
    Flow_ymin = yminQ, Flow_ymax = ymaxQ,
    GPP_ymin = yminGPP, GPP_ymax = ymaxGPP,
    ER_ymin = yminER, ER_ymax = ymaxER,
    WaterTemp_ymin = yminWtemp, WaterTemp_ymax = ymaxWtemp,
    PPT_ymin = yminPPT, PPT_ymax = ymaxPPT,
    AirTemp_ymin = yminAtemp, AirTemp_ymax = ymaxAtemp,
    yminLAI = yminLAI, ymaxLAI = ymaxLAI
  ) %>%
  pivot_longer(cols = -x, names_to = c("group", ".value"), names_sep = "_")

synch.plot <- ggplot() +
  geom_line(data = pred_df%>%dplyr::filter(x<twothirds_dist_km),
            aes(x = x, y = y, color = group)) +
     geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = group),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = group),
               linetype = "dashed") +
  theme_classic() +
  ylim(-0.2,1)+
  scale_color_manual(values = c(
    "Flow" = "#0072B2",
    "GPP" = "#1aab1a",
    "ER" = "#f5880c",
    "Water Temp" = "#ed1111",
    "PPT" = "#56B4E9",
    "Air Temp" = "#de21bb",
    "Light" = "#E2D523", 
    "LAI"= "#8f21de"
  )) +
   geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminQ, ymax = ymaxQ),
              fill = "#0072B2", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP),
              fill = "#1aab1a", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminWtemp, ymax = ymaxWtemp),
              fill = "#ed1111", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminPPT, ymax = ymaxPPT),
              fill = "#56B4E9", alpha=0.2) +
  geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
             aes(x = x, ymin = yminAtemp, ymax = ymaxAtemp),
              fill = "#de21bb", alpha=0.2)+
      geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminlight, ymax = ymaxlight),
              fill = "#E2D523", alpha=0.2) +
      geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminER, ymax = ymaxER),
              fill = "#f5880c", alpha=0.2) +
    geom_ribbon(data = joined_df%>%dplyr::filter(x<twothirds_dist_km),
              aes(x = x, ymin = yminLAI, ymax = ymaxLAI),
              fill = "#8f21de", alpha=0.2) +
    labs(y= "Correlation",
       x="Distance (km)",
       color = "Variable", fill = "Variable") +
  theme(text = element_text(size=14),
        legend.position = "bottom") 

synch.plot

```
```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Moran_spline_correlogram_p4.png", plot = synch.plot, width =6, height = 5, units = "in")

```


