---
title: "Part 4. Spline Correlogram analysis for synchrony in aquatic ecosystem productivity - Fine scale conditions"
author: "K. Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo=FALSE, include=FALSE}
library(data.table) # Allows fast reading and writing of files
library(ggplot2)    # For plotting
library(tidyverse)  # For data processing
library(lubridate)  # For date processing
library(reshape2)   # For reshaping the data into long format
library(ncf)        # For Synchrony analysis
library(grid)       # for labeling the gridded plot
library(gridExtra)  # for creating a labeled, gridded plot
library(FSA)        # For making depletion estimates
library(knitr)      # For making tidy tables
library(broom)
library(grid)

```

## Background: Synchrony estimates using Sncf{ncf}

#### How it works:

Sncf computes the nonparametric spatial (auto/cross-)correlation
function for multivariate time series data observed across spatial
locations.

1.  Takes time series data z from n locations (one row per location, one
    column per time step).

2.  Calculates the pairwise correlations between time series at each
    location pair.

3.  Bins those correlations based on distance between sites (using
    latitude and longitude if latlon = TRUE).

4.  Fits a smoothing spline to estimate the average correlation as a
    function of distance
    
It can handle missing data, where:
na.rm = TRUE: Missing values are allowed; only overlapping time steps are used in correlations.


#### Metrics to consider:

-   '\real$predicted\$y[1]' = Synchrony at (almost) zero distance –
    often max synchrony.
      
      - This is the first predicted value from the fitted nonparametric spline of the spatial correlation function
      
      - Estimated synchrony at the smallest distance bin


-   '\$y.intercept' = Extrapolated synchrony at exactly zero distance 
      
      - The theoretical synchrony among perfectly co-located sites

-   '\$cbar' = Regional mean synchrony across all distances

-   '\$x.intercept'	= Distance at which synchrony drops to 0

-   '\$e.intercept' = Distance where synchrony drops to 1/e of max (decay scale)


###  Fletcher, R., & Fortin, M. (2018):
we should truncate the range of lag distances at which we consider spatial 
dependence to approximately ½ to 2 = 3 the total distance observed.

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

river_metab_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/stream_spline_cor_df.rds") %>% mutate(GPP_filled = if_else(GPP_filled <= 0, 0.0001, GPP_filled),
          ER_filled = if_else(ER_filled >= 0, -0.0001, ER_filled))

names(river_metab_dat)
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
### read in prism data for PPT

PPT_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Clim_data/PRISM_stream.rds")%>%
  mutate(Site_ID=as.character(Name))


StreamLight_dat <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/doi_10_5061_dryad_bcc2fqzj2__v20230622/data_citation/7_river_light.csv") %>%
  mutate(date = lubridate::as_date(date, format="%Y-%m-%d")) %>%
  mutate(Site_ID1 = str_remove(site, "^nwis_")) %>%
  mutate(Site_ID = str_remove(Site_ID1, "^0+"))


river_metab_dat1 <- river_metab_dat%>%
  left_join(PPT_dat, by=c("date"="Date", "Site_ID")) %>%
  dplyr::select(site:TmaxC) 

river_metab_dat1 <- river_metab_dat1%>%
  left_join(StreamLight_dat, by=c("date", "site", "Site_ID"))

intersect(unique(river_metab_dat1$site), unique(StreamLight_dat$site))


setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab



sites_keep <- read.csv("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Map_data_for_nested_reach.csv")%>%
  filter(keep == 1) %>%
  dplyr::select(Site_ID:stream_order)


# Filter river_metab_dat to keep only Site_IDs in sites_keep
river_metab_dat1 <- river_metab_dat1 %>%
  filter(Site_ID %in% sites_keep$Site_ID)

unique(river_metab_dat1$Site_ID)
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
library(dplyr)
library(geodist)

# Get site centroids
site_coords <- river_metab_dat1 %>%
  group_by(Site_ID) %>%
  summarise(Lat = mean(Lat), Lon = mean(Lon), .groups = "drop")

# Calculate the overall mean coordinates
mean_lat <- median(site_coords$Lat)
mean_lon <- median(site_coords$Lon)

# Find site closest to mean coordinates using geodist
mean_coords <- data.frame(Lat = mean_lat, Lon = mean_lon)
site_coords_df <- data.frame(Lat = site_coords$Lat, Lon = site_coords$Lon)

distances_to_mean <- geodist(site_coords_df, mean_coords, measure = "geodesic") / 1000
closest_index <- which.min(distances_to_mean)

# Use the closest site as the reference
ref_site <- site_coords[closest_index, ]

# Compute distances from the reference site to all others
ref_coords <- data.frame(Lat = ref_site$Lat, Lon = ref_site$Lon)
distances_from_ref <- geodist(site_coords_df, ref_coords, measure = "geodesic") / 1000

# Add distance column
site_coords <- site_coords %>%
  mutate(distance_km = as.numeric(distances_from_ref))

hist(site_coords$distance_km)


# Get 2/3 cut off 

half_dist_km <- max(site_coords$distance_km)/ 2
half_dist_km
twothirds_dist_km <- max(site_coords$distance_km) * (2/3)
twothirds_dist_km

```

## Part 1: Synchrony in metabolism at differnt timescales

Compare synchrony for GPP and ER from 56 different sites in Marzolf
dataset at different timescales:

### A. GPP synchrony at differnt timescales:

#### GPP- Daily synchrony

```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}

#### Daily synchrony

river_metab_dat<-as.data.frame(river_metab_dat1)

GPP_long <- river_metab_dat1 %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day <- GPP_long %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_day <- Sncf(x=GPP_wide_day$Lon,
                   y=GPP_wide_day$Lat,
                   z=GPP_wide_day[,4:ncol(GPP_wide_day)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year

plot(RR_df.corr_day)

GPP_corrgram_summary.table <- summary(RR_df.corr_day)
Regional.synch <- GPP_corrgram_summary.table$Regional.synch


# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_day$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP")

## Sample average line
GPP.corr.pred.df.day$mean_synch <- round(Regional.synch,2)
Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.df.day$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.df <- as.data.frame(t(RR_df.corr_day$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.df.day <- data.frame(x = matrix(unlist(RR_df.corr_day$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.df$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.df$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP")



daily_dat <- data.frame(
  strat = "daily",
  y = max(GPP.corr.pred.df.day$y, na.rm = TRUE),
  ymin = mean(GPP.corr.boot.df.day$yminGPP),
  ymax = mean(GPP.corr.boot.df.day$ymaxGPP)
)

```




```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}

cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))

```


## Part 2 - Moran Effect

-   MODIS data for leaf area index (LAI)

-   Daily streamflow (Q)

-   Daily water temperature (water_temp)

-   Daily air temp (air_temp)

-   Daily precipitation (PPT)

#### Bring in MODIS data for LAI to get at

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}
# https://github.com/gpvalentine/SE_Trout/blob/main/Analysis/SE_Trout_Correlograms.R

LAI_dat <-readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Modis_data/StreamDat_MOD_combined.rds")%>%
  mutate(date=as.Date(Date)) %>%
    filter(Site_ID %in% sites_keep$Site_ID)


# intersect(unique(river_metab_dat1$Site_ID), unique(LAI_dat$Site_ID))
# 
# 
# Metat_sites_notinlight <- setdiff(unique(river_metab_dat1$site), unique(StreamLight_dat$site)) # Sites in metab but not in light
# Light_sites_notinmetab<- setdiff(unique(StreamLight_dat$site), unique(river_metab_dat1$site)) # Sites in light but not in metab
# 
# LAI_sites_notinlight<- setdiff(unique(LAI_dat$Site_ID), unique(StreamLight_dat$Site_ID)) # Sites in light but not in metab

```

## Part 3. Landscape level factors as explaining synchrony:

A. left join time series dat with landscape parameters

B. ER / GPP explained by stream size (via regression)

i.  Dominant land cover explained by stream size


iii. Slope/ Catchment size/ basin characteristics

iv. Water temperature/ flow

C. Create landscape level groups and then run synchronous

i.  landcover (PCA? developed to forest?) - possibly cropland with
    moderate development

iii. Slope? or Catchment size

iv. Water temperature/ depth/flow?

<!-- -->

5.  When landscape character explains some synchronous, visualize with
    synch splines (Q1-Q3 for i to iv splines)

```{r, size='large', echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

dat_NHDinfo<- readRDS("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/data/Metab_metadata_25v2.rds")%>%
  mutate(Site_ID = str_remove(site_no, "^0+"))

river_metab_dat1 <- river_metab_dat1%>%
  left_join(dat_NHDinfo, by=c("Site_ID")) 


river_metab_dat1 <- river_metab_dat1 %>%
  mutate(areasqkm = if_else(is.na(areasqkm) & Site_ID == 71912213,
                            422.168,
                              areasqkm)) %>%
  mutate(flow_km = c(discharge/areasqkm))

summary(river_metab_dat1)
```



## Part 4: How much variance in GPP/ER is explained by temperature, flow, and streamlight

#### A. Water temp.

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

river_LF_sum <- river_metab_dat1%>%
  group_by(Site_ID) %>%
  summarise(
    m_temp =median(temp.water, na.rm = T), 
    m_PAR =median(daily_PAR, na.rm = T),
    m_LAI =median(daily_LAI, na.rm = T), 
    m_flow =median(discharge, na.rm = T))

river_LF_sum

river_LF_sum <- river_metab_dat1%>%
  group_by(Site_ID) %>%
  summarise(
    m_temp =mean(temp.water, na.rm = T), 
    m_PAR =mean(daily_PAR, na.rm = T),
    m_LAI =mean(daily_LAI, na.rm = T), 
    m_flow =mean(discharge, na.rm = T), 
    m_flow_km =mean(flow_km, na.rm = T)
    )

river_LF_sum

summary(river_LF_sum)
```



###### Quantile of median temps at each site :

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
river_metab_t <- na.omit(river_LF_sum$m_temp)

# Get 1/3 cutoffs (33.3% and 66.6%)
cut_points <- quantile(river_metab_t, probs = c(1/3, 2/3))

# Create group labels
river_LF_sum$Temp_group <- cut(
  river_LF_sum$m_temp,
  breaks = c(-Inf, cut_points[1], cut_points[2], Inf),
  labels = c("Q1", "Q2", "Q3"),
  right = TRUE
)

table(river_LF_sum$Temp_group)



tab_river_LF_sum <- river_LF_sum%>%
  group_by(Temp_group) %>%
  summarise(X_min=round(min(m_temp, na.rm=T),2),
            X_max=round(max(m_temp, na.rm=T),2), 
            site_no =length(unique(Site_ID)))


tab_river_LF_sum
```


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

df_wtemp <- river_metab_dat1 %>%
  left_join(river_LF_sum%>%dplyr::select(Site_ID, Temp_group), by=c("Site_ID"))

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Q1" = "#00BFC4",   
  "Q2" = "#7CAE00",
  "Q3" = "#F8766D"
)
```


```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_wtemp, aes(x = temp.water, fill = factor(Temp_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colors,
    name = "temp category",
    labels = c("Cold", "Q2", "Warm")
  ) +
  theme_minimal() +
  labs(title = "Water temp. values by quantile",
       x = "Water temp.",
       y = "Count")



# # Plot histogram
# Temp_dist_plt <- ggplot(df_wtemp, aes(x = temp.water, fill = Temp_group)) +
#   geom_density(position = "identity", alpha = 0.6, bins = 30) +
#   scale_fill_manual(
#     values = strat_colors,
#     name = "Water temp. group",
#     labels = c("Low", "Moderate", "High")
#   ) +
#   #scale_x_reverse() +
#   scale_y_reverse() +
#   theme_minimal() +
#   labs(title = "Cold, moderate, and warm sites",
#        x = "Water temp.",
#        y = "Density") +theme_classic()


Temp_dist_plt <- ggplot(df_wtemp, aes(x = temp.water, fill = Temp_group)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(
    values = strat_colors,
    name = "Water temp. group",
    labels = c("Low", "Moderate", "High")
  ) +
 # coord_flip() +  # Rotate 90 degrees
  theme_classic() +
  labs(title = "Cold, moderate, and\nwarm sites",
       x = "Water temp.",
       y = "Density") +
        theme(plot.title = element_text(size = 10))


```



```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

metab_units_area <- (expression(paste('Metabolism (g C', m^-2,' ',y^-1,')', sep = ' ')))
low_p <- '#6e2f9e'
high_p <- '#f4e6ff'

low_er <- '#381d01'
high_er <- '#fa9600'


WQ_temp_plot <- df_wtemp%>%
  group_by(Site_ID) %>%
  summarise(
    GPP = mean(GPP_filled, na.rm=T),
    ER = mean(c(ER_filled*-1), na.rm=T),
    Wtemp =mean(temp.water, na.rm=T))


library(ggplot2)
library(ggnewscale)  

fig_combined2 <- ggplot(WQ_temp_plot, aes(x = Wtemp)) +
  # GPP points (upward triangle)
  geom_point(aes(y = GPP, fill = GPP, color = GPP), shape = 24, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "GPP", low = low_p, high = high_p) +
  scale_color_gradient(name = "GPP", low = low_p, high = high_p) +
  # Reset fill scale so ER can have its own
  new_scale_fill() +
    new_scale_color() +
  # ER points (downward triangle)
  geom_point(aes(y = ER, fill = ER, color = ER), shape = 25, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "ER", low = low_er, high = high_er) +
  scale_color_gradient(name = "ER", low = low_er, high = high_er) +
  scale_y_continuous(name = metab_units_area, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
    theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    panel.grid = element_blank()
  )

# ggsave(plot = fig_combined2, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/WQ_fig1d_25.png",sep=""),width=5.25,height=3.5,dpi=300)

```

GLM:

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

library(lme4)
library(lmerTest)
library(MuMIn)

lm_model_temp <- lmer(log(GPP_filled) ~ temp.water + (1|Temp_group), data = df_wtemp)
summary(lm_model_temp)
hist(residuals(lm_model_temp))
r.squaredGLMM(lm_model_temp)

lm_model_tempn <- lm(log(GPP_filled) ~ temp.water, data = df_wtemp)
summary(lm_model_tempn)
hist(residuals(lm_model_tempn))

AIC(lm_model_temp, lm_model_tempn)

```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
library(lme4)
library(lmerTest)
library(MuMIn)

lm_model_temp <- lmer(log(ER_filled*-1) ~ temp.water + (1|Temp_group), data = df_wtemp)
summary(lm_model_temp)
hist(residuals(lm_model_temp))
r.squaredGLMM(lm_model_temp)

lm_model_tempn <- lm(log(ER_filled*-1) ~ temp.water, data = df_wtemp)
summary(lm_model_tempn)
hist(residuals(lm_model_tempn))

AIC(lm_model_temp, lm_model_tempn)
```

#### Temperature splines

#### Daily synchrony GPP and Water temp Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- df_wtemp %>%
  dplyr::filter(Temp_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_temp_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.temp.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q1")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.temp.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.temp.Q1$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.temp.Q1 <- as.data.frame(t(RR_df.corr_temp_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.temp.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.temp.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.temp.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q1")

```


```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony temp Q2

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- df_wtemp %>%
    dplyr::filter(Temp_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_temp_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.temp.q2 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q2")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.temp.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.temp.q2$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.temp.q2 <- as.data.frame(t(RR_df.corr_temp_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.temp.q2 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.temp.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.temp.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q2")
```


```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony temp Q3

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- df_wtemp %>%
      dplyr::filter(Temp_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_temp_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.temp.q3 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q3")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.temp.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.temp.q3$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.temp.q3 <- as.data.frame(t(RR_df.corr_temp_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.temp.q3 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.temp.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.temp.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q3")

```


```{r, warning=F, size = 'small', echo=T, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_temp <- bind_rows(GPP.corr.pred.temp.Q1, GPP.corr.pred.temp.q2, GPP.corr.pred.temp.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_temp   <- bind_rows(GPP.corr.boot.temp.Q1, GPP.corr.boot.temp.q2, GPP.corr.boot.temp.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different flow quantiles

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_temp <- pred_df_all_temp %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_temp <- pred_wide_temp %>%
  mutate(
    temp_Q1_diff = (abs(temp_Q1 - temp_Q1) / temp_Q1) * 100,
    temp_Q2_diff = (abs(temp_Q1 - temp_Q2) / temp_Q1) * 100,
    temp_Q3_diff = (abs(temp_Q1 - temp_Q3) / temp_Q1) * 100)

pred_diff_temp
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "temp_Q1" = "#00BFC4",   
  "temp_Q2" = "#7CAE00",
  "temp_Q3" = "#F8766D"
)
```


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_temp %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

ERbar_chart_temp  <-ci_df_all_temp %>%
  filter(x_rounded_5<635)%>%
  group_by(strat) %>%
  summarise(
    ymin= mean(yminGPP), 
    ymax=mean(ymaxGPP))

coeff_chart_temp <- pred_df_all_temp %>%
  group_by(strat) %>%
  summarise(
    y= max(mean_synch))

coef_plot_temp_dat <- coeff_chart_temp%>%
  left_join(ERbar_chart_temp, by =c("strat" = "strat"))

intersection_df
```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_temp <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_temp,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_temp,
            aes(x = x, y = y, color = strat),
            size = 1) +
  labs(title = "GPP across water temperature",
              y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
    geom_vline(xintercept = 807, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Low", "Moderate", "High")) +
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  theme_classic() +   ylim(-0.2, 0.75) +
  theme(legend.position = "bottom")

GPP_synch_spline_temp

```


Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_temp %>% dplyr::filter(x<half_dist_km)%>%
  group_by(strat) %>%
  filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

#### Daily synchrony ER and Water temp Q1

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q1 <- df_wtemp %>%
  dplyr::filter(Temp_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q1 <- ER_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q1 <- Sncf(x=ER_wide_day_q1$Lon,
                   y=ER_wide_day_q1$Lat,
                   z=ER_wide_day_q1[,4:ncol(ER_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q1)


ER_corrgram_summary.table <- summary(RR_df.corr_temp_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.temp.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q1")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.temp.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.temp.Q1$cbar <- Regional.synch_x



# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.temp.Q1 <- as.data.frame(t(RR_df.corr_temp_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.temp.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = ER.corr.bootValues.temp.Q1$`0.025`,
                                             ymaxGPP = ER.corr.bootValues.temp.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q1")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


#### Daily synchrony temp Q2

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q2 <- df_wtemp %>%
    dplyr::filter(Temp_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q2 <- ER_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q2 <- Sncf(x=ER_wide_day_q2$Lon,
                   y=ER_wide_day_q2$Lat,
                   z=ER_wide_day_q2[,4:ncol(ER_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q2)


ER_corrgram_summary.table <- summary(RR_df.corr_temp_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.temp.q2 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q2")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.temp.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.temp.q2$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.temp.q2 <- as.data.frame(t(RR_df.corr_temp_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.temp.q2 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = ER.corr.bootValues.temp.q2$`0.025`,
                                             ymaxGPP = ER.corr.bootValues.temp.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q2")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony temp Q3

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q3 <- df_wtemp %>%
      dplyr::filter(Temp_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q3 <- ER_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_temp_q3 <- Sncf(x=ER_wide_day_q3$Lon,
                   y=ER_wide_day_q3$Lat,
                   z=ER_wide_day_q3[,4:ncol(ER_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_temp_q3)


ER_corrgram_summary.table <- summary(RR_df.corr_temp_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.temp.q3 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_temp_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "temp_Q3")
## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.temp.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.temp.q3$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.temp.q3 <- as.data.frame(t(RR_df.corr_temp_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.temp.q3 <- data.frame(x = matrix(unlist(RR_df.corr_temp_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = ER.corr.bootValues.temp.q3$`0.025`,
                                             ymaxGPP = ER.corr.bootValues.temp.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "temp_Q3")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_temp <- bind_rows(ER.corr.pred.temp.Q1, ER.corr.pred.temp.q2, ER.corr.pred.temp.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_temp   <- bind_rows(ER.corr.boot.temp.Q1, ER.corr.boot.temp.q2, ER.corr.boot.temp.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_temp %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


##### Calculate the percent difference between different flow quantiles


```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_temp <- pred_df_all_temp %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_temp <- pred_wide_temp %>%
  mutate(
    temp_Q1_diff = (abs(temp_Q1 - temp_Q1) / temp_Q1) * 100,
    temp_Q2_diff = (abs(temp_Q1 - temp_Q2) / temp_Q1) * 100,
    temp_Q3_diff = (abs(temp_Q1 - temp_Q3) / temp_Q1) * 100)

pred_diff_temp
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "temp_Q1" = "#00BFC4",   
  "temp_Q2" = "#7CAE00",
  "temp_Q3" = "#F8766D"
)
```

### Figure 4c (?): Spline Correlogram of GPP at different temp characteristics

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}
###
ER_synch_spline_temp <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_temp,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  geom_line(data = pred_df_all_temp,
            aes(x = x, y = y, color = strat),
            size = 1) +
       # Horizontal mean synchrony lines
   geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  labs(title = "ER across water temperature",
              y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
      geom_vline(xintercept = 1083, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Low", "Moderate", "High")) +
  theme_classic() +  ylim(-0.2, 0.75) +
  theme(legend.position = "bottom")

ER_synch_spline_temp
```


Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_temp %>%
  filter(x<cbar)%>%
  group_by(strat) %>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

#### B. Streamflow

###### Quantile of median temps at each site :

```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
river_metab_flow <- na.omit(river_LF_sum$m_flow)

# Get 1/3 cutoffs (33.3% and 66.6%)
cut_points <- quantile(river_metab_flow, probs = c(1/3, 2/3))

# Create group labels
river_LF_sum$Flow_group <- cut(
  river_LF_sum$m_flow,
  breaks = c(-Inf, cut_points[1], cut_points[2], Inf),
  labels = c("Q1", "Q2", "Q3"),
  right = TRUE
)


tab_river_LF_sum <- river_LF_sum%>%
  group_by(Flow_group) %>%
  summarise(X_min=round(min(m_flow, na.rm=T),2),
            X_max=round(max(m_flow, na.rm=T),2), 
            site_no =length(unique(Site_ID)))


tab_river_LF_sum

```


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

df_flow <- river_metab_dat1 %>%
  left_join(river_LF_sum%>%dplyr::select(Site_ID, Flow_group), by=c("Site_ID"))

```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Q1" = "#00BFC4",   
  "Q2" = "#7CAE00",
  "Q3" = "#F8766D"
)
```


```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_flow, aes(x = log(discharge), fill = factor(Flow_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colors,
    name = "temp category",
    labels = c("Q1", "Q2", "Q3")
  ) +
  theme_minimal() +
  labs(title = "Streamflow. values by quantile",
       x = "log(Streamflow)",
       y = "Count")


# Plot histogram
Flow_dist_plt <- ggplot(df_flow, aes(x = log(discharge), fill = Flow_group)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(
    values = strat_colors,
    name = "Streamflow group",
    labels = c("Low", "Moderate", "High")
  ) +
    #coord_flip() +  # Rotate 90 degrees
  theme_minimal() +
  labs(title = "Low, moderate, and\nhigh flow sites",
       x = "log(Flow)",
       y = "Density") +theme_classic() +
      theme(plot.title = element_text(size = 10))

```



```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

metab_units_area <- (expression(paste('Metabolism (g C', m^-2,' ',y^-1,')', sep = ' ')))

low_p <- '#6e2f9e'
high_p <- '#f4e6ff'

low_er <- '#381d01'
high_er <- '#fa9600'


WQ_qdis_plot <- df_wtemp%>%
  group_by(Site_ID) %>%
  summarise(
    GPP = mean(GPP_filled, na.rm=T),
    ER = mean(c(ER_filled*-1), na.rm=T),
    Qdis =mean(log(discharge), na.rm=T))


library(ggplot2)
library(ggnewscale)  

fig_combined3 <- ggplot(WQ_qdis_plot, aes(x = Qdis)) +
  # GPP points (upward triangle)
  geom_point(aes(y = GPP, fill = GPP, color = GPP), shape = 24, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "GPP", low = low_p, high = high_p) +
  scale_color_gradient(name = "GPP", low = low_p, high = high_p) +
  # Reset fill scale so ER can have its own
  new_scale_fill() +
    new_scale_color() +
  # ER points (downward triangle)
  geom_point(aes(y = ER, fill = ER, color = ER), shape = 25, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "ER", low = low_er, high = high_er) +
  scale_color_gradient(name = "ER", low = low_er, high = high_er) +
  scale_y_continuous(name = metab_units_area, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
    theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    panel.grid = element_blank()
  )


# ggsave(plot = fig_combined3, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/WQ_fig1e_25.png",sep=""),width=5.25,height=3.5,dpi=300)

```




GLM:

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

lm_model_flow <- lmer(log(GPP_filled) ~ discharge + (1|Flow_group), data = df_flow)
summary(lm_model_flow)
hist(residuals(lm_model_flow))
r.squaredGLMM(lm_model_flow)

lm_model_flown <- lm(log(GPP_filled) ~ discharge, data = df_flow)
summary(lm_model_flown)
hist(residuals(lm_model_flown))

AIC(lm_model_flow, lm_model_flown)

```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
lm_model_flow <- lmer(log(ER_filled*-1) ~ discharge + (1|Flow_group), data = df_flow)
summary(lm_model_flow)
hist(residuals(lm_model_flow))
r.squaredGLMM(lm_model_flow)

lm_model_flown <- lm(log(ER_filled*-1) ~ discharge, data = df_flow)
summary(lm_model_flown)
hist(residuals(lm_model_flown))

AIC(lm_model_flow, lm_model_flown)
```

#### Daily synchrony GPP and flow Q1

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q1 <- df_flow %>%
  dplyr::filter(Flow_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q1")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.flow.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.flow.Q1$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.Q1 <- as.data.frame(t(RR_df.corr_flow_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q1")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony GPP and flow Q2

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- df_flow %>%
    dplyr::filter(Flow_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q2")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.flow.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.flow.q2$cbar <- Regional.synch_x


# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.q2 <- as.data.frame(t(RR_df.corr_flow_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q2")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony GPP and flow Q3

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- df_flow %>%
      dplyr::filter(Flow_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_flow_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "flow_Q3")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.flow.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.flow.q3$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.flow.q3 <- as.data.frame(t(RR_df.corr_flow_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.flow.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.flow.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "flow_Q3")

```



```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_flow <- bind_rows(GPP.corr.pred.flow.Q1, GPP.corr.pred.flow.q2, GPP.corr.pred.flow.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_flow   <- bind_rows(GPP.corr.boot.flow.Q1, GPP.corr.boot.flow.q2, GPP.corr.boot.flow.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_flow <- pred_df_all_flow %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_flow <- pred_wide_flow %>%
  mutate(
    flow_Q1_diff = (abs(flow_Q1 - flow_Q1) / flow_Q1) * 100,
    flow_Q2_diff = (abs(flow_Q1 - flow_Q2) / flow_Q1) * 100,
    flow_Q3_diff = (abs(flow_Q1 - flow_Q3) / flow_Q1) * 100)

pred_diff_flow
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "flow_Q1" = "#00BFC4",   
  "flow_Q2" = "#7CAE00",
  "flow_Q3" = "#F8766D"
)
```


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_flow %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df


ERbar_chart_flow  <-ci_df_all_flow %>%
  filter(x_rounded_5<635)%>%
  group_by(strat) %>%
  summarise(
    ymin= mean(yminGPP), 
    ymax=mean(ymaxGPP))

coeff_chart_flow <- pred_df_all_flow %>%
  group_by(strat) %>%
  summarise(
    y= max(mean_synch))

coef_plot_flow_dat <- coeff_chart_flow %>%
  left_join(ERbar_chart_flow, by =c("strat" = "strat"))

```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_flow <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_flow,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  ylim(-0.2, 0.75) +
  geom_line(data = pred_df_all_flow,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Draw horizontal segments from start to x_intersect
  geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  labs(title = "GPP across flow",       
       y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
        geom_vline(xintercept = 807, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Low", "Moderate", "High")) +
  theme_classic() +  ylim(-0.2, 0.75) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_flow

```

Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_flow %>%
  group_by(strat) %>%
    filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

#### Daily synchrony ER and flow Q1

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q1 <- df_flow %>%
  dplyr::filter(Flow_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q1 <- ER_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q1 <- Sncf(x=ER_wide_day_q1$Lon,
                   y=ER_wide_day_q1$Lat,
                   z=ER_wide_day_q1[,4:ncol(ER_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q1)


ER_corrgram_summary.table <- summary(RR_df.corr_flow_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "flow_Q1")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.flow.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.flow.Q1$cbar <- Regional.synch_x


# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.flow.Q1 <- as.data.frame(t(RR_df.corr_flow_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.flow.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q1$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.flow.Q1$`0.025`,
                                             ymaxER = ER.corr.bootValues.flow.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "flow_Q1")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony ER and flow Q2

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q2 <- df_flow %>%
    dplyr::filter(Flow_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q2 <- ER_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q2 <- Sncf(x=ER_wide_day_q2$Lon,
                   y=ER_wide_day_q2$Lat,
                   z=ER_wide_day_q2[,4:ncol(ER_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q2)


ER_corrgram_summary.table <- summary(RR_df.corr_flow_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "flow_Q2")


## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.flow.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.flow.q2$cbar <- Regional.synch_x


# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.flow.q2 <- as.data.frame(t(RR_df.corr_flow_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.flow.q2 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q2$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.flow.q2$`0.025`,
                                             ymaxER = ER.corr.bootValues.flow.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "flow_Q2")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony ER and flow Q3

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q3 <- df_flow %>%
      dplyr::filter(Flow_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q3 <- ER_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_flow_q3 <- Sncf(x=ER_wide_day_q3$Lon,
                   y=ER_wide_day_q3$Lat,
                   z=ER_wide_day_q3[,4:ncol(ER_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_flow_q3)


ER_corrgram_summary.table <- summary(RR_df.corr_flow_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_flow_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "flow_Q3")

## Sample average line


## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.flow.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.flow.q3$cbar <- Regional.synch_x


# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.flow.q3 <- as.data.frame(t(RR_df.corr_flow_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.flow.q3 <- data.frame(x = matrix(unlist(RR_df.corr_flow_q3$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.flow.q3$`0.025`,
                                             ymaxER = ER.corr.bootValues.flow.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "flow_Q3")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_flow <- bind_rows(ER.corr.pred.flow.Q1, ER.corr.pred.flow.q2, ER.corr.pred.flow.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_flow   <- bind_rows(ER.corr.boot.flow.Q1, ER.corr.boot.flow.q2, ER.corr.boot.flow.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different flow quantiles

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_flow <- pred_df_all_flow %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_flow <- pred_wide_flow %>%
  mutate(
    flow_Q1_diff = (abs(flow_Q1 - flow_Q1) / flow_Q1) * 100,
    flow_Q2_diff = (abs(flow_Q1 - flow_Q2) / flow_Q1) * 100,
    flow_Q3_diff = (abs(flow_Q1 - flow_Q3) / flow_Q1) * 100)

pred_diff_flow
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "flow_Q1" = "#00BFC4",   
  "flow_Q2" = "#7CAE00",
  "flow_Q3" = "#F8766D"
)
```


##### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_flow %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df
```


```{r, warning=F, size = 'small', echo=T, include=T}

ER_synch_spline_flow <- ggplot() +
  geom_ribbon(data = ci_df_all_flow,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = strat),
              alpha = 0.1) +
  geom_line(data = pred_df_all_flow,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Draw horizontal segments from start to x_intersect
  geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  labs(title = "ER across flow",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
        geom_vline(xintercept = 1083, color="grey40", lty=3) +
    scale_color_manual(values = strat_colors,
                     labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,
                    labels = c("Low", "Moderate", "High")) +
  theme_classic() +  ylim(-0.2, 0.75) +
  theme_classic() +  ylim(-0.2, 0.75) +
  theme(legend.position = "bottom")


```



### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_flow %>%
  group_by(strat) %>%
      filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

#### C. Light at stream surface



```{r, size='large', echo=T, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
river_metab_light <- na.omit(river_LF_sum$m_PAR)

# Get 1/3 cutoffs (33.3% and 66.6%)
cut_points <- quantile(river_metab_light, probs = c(1/3, 2/3))

# Create group labels
river_LF_sum$Light_group <- cut(
  river_LF_sum$m_PAR,
  breaks = c(-Inf, cut_points[1], cut_points[2], Inf),
  labels = c("Q1", "Q2", "Q3"),
  right = TRUE
)


tab_river_LF_sum <- river_LF_sum%>%
  group_by(Light_group) %>%
  summarise(X_min=round(min(m_PAR, na.rm=T),2),
            X_max=round(max(m_PAR, na.rm=T),2), 
            site_no =length(unique(Site_ID)))

tab_river_LF_sum
```


```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
df_light <- river_metab_dat1 %>%
  left_join(river_LF_sum%>%dplyr::select(Site_ID, Light_group), by=c("Site_ID"))


# Site 2336313 PC1: 5.19449576
# Site 2336360 PC1: 2.53144230
# Site 2336526 PC1: 2.97663410
# Site 2336728 PC1: 1.01503347

# 1400500, 14206950, 14211010, 1649500, 2169000, 2203655, 2336313, 2336360, 2336526,8049500	

df_light <- df_light%>%
  mutate(Light_group = if_else(is.na(Light_group) & Site_ID == 1400500, 
                              "Q2", Light_group),
    Light_group = if_else(is.na(Light_group) & Site_ID == 2336313,
                              "Q2", Light_group),
         Light_group = if_else(is.na(Light_group) & Site_ID == 2336360,
                              "Q1", Light_group),
         Light_group = if_else(is.na(Light_group) & Site_ID == 2336526,
                              "Q2", Light_group),
             Light_group = if_else(is.na(Light_group) & Site_ID == 14206950,
                              "Q2", Light_group),
                Light_group = if_else(is.na(Light_group) & Site_ID == 14211010,
                              "Q3", Light_group),
                    Light_group = if_else(is.na(Light_group) & Site_ID == 1649500,
                              "Q3", Light_group),
                Light_group = if_else(is.na(Light_group) & Site_ID == 2169000,
                              "Q2", Light_group),
        Light_group = if_else(is.na(Light_group) & Site_ID == 2203655,
                              "Q1", Light_group),
            Light_group = if_else(is.na(Light_group) & Site_ID == 8049500,
                              "Q3", Light_group),
        Light_group = if_else(is.na(Light_group) & Site_ID == 8049500,
                              "Q3", Light_group),
         Light_group = if_else(is.na(Light_group) & Site_ID == 2336728,
                              "Q1", Light_group)) 


tab_river_LF_sum <- df_light%>%
  group_by(Light_group) %>%
  summarise(site_no =length(unique(Site_ID)))

tab_river_LF_sum

```



```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "Q1" = "#00BFC4",   
  "Q2" = "#7CAE00",
  "Q3" = "#F8766D"
)
```


```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}
ggplot(df_light, aes(x = daily_PAR, fill = factor(Light_group))) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  scale_fill_manual(
    values = strat_colors,
    name = "temp category",
    labels = c("Q1", "Q2", "Q3")
  ) +
  theme_minimal() +
  labs(title = "Light at stream surface values by quantile",
       x = "Light",
       y = "Count")



# Plot histogram
Light_dist_plt <- ggplot(df_light, aes(x = daily_PAR, fill = Light_group)) +
  geom_density(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(
    values = strat_colors,
    name = "Light group",
    labels = c("Low", "Moderate", "High")
  ) +
      #coord_flip() +  # Rotate 90 degrees
  theme_minimal() +
  labs(title = "Low, moderate, and\nhigh light sites",
       x = "Light as PAR",
       y = "Density") +theme_classic() +
    theme(plot.title = element_text(size = 10))

```




```{r, size='large', echo=F, include=T, warning=FALSE, cache=TRUE, fig.width=6, fig.height=4}

metab_units_area <- (expression(paste('Metabolism (g C', m^-2,' ',y^-1,')', sep = ' ')))

low_p <- '#6e2f9e'
high_p <- '#f4e6ff'

low_er <- '#381d01'
high_er <- '#fa9600'


WQ_light_plot <- df_light%>%
  group_by(Site_ID) %>%
  summarise(
    GPP = mean(GPP_filled, na.rm=T),
    ER = mean(c(ER_filled*-1), na.rm=T),
    PAR =mean(daily_PAR, na.rm=T))


library(ggplot2)
library(ggnewscale)  

fig_combined4 <- ggplot(WQ_light_plot, aes(x = PAR)) +
  # GPP points (upward triangle)
  geom_point(aes(y = GPP, fill = GPP, color = GPP), shape = 24, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "GPP", low = low_p, high = high_p) +
  scale_color_gradient(name = "GPP", low = low_p, high = high_p) +
  # Reset fill scale so ER can have its own
  new_scale_fill() +
    new_scale_color() +
  # ER points (downward triangle)
  geom_point(aes(y = ER, fill = ER, color = ER), shape = 25, size = 2, alpha=0.75) +
  scale_fill_gradient(name = "ER", low = low_er, high = high_er) +
  scale_color_gradient(name = "ER", low = low_er, high = high_er) +
  scale_y_continuous(name = metab_units_area, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
      theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    panel.grid = element_blank()
  )
     
# ggsave(plot = fig_combined4, filename = paste("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/WQ_fig1f_25.png",sep=""),width=5.25,height=3.5,dpi=300)

```


GLM:

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}

library(lme4)
library(lmerTest)
library(MuMIn)

lm_model_light <- lmer(log(GPP_filled) ~ daily_PAR + (1|Light_group), data = df_light)
summary(lm_model_light)
hist(residuals(lm_model_light))
r.squaredGLMM(lm_model_light)

lm_model_lightn <- lm(log(GPP_filled) ~ daily_PAR, data = df_light)
summary(lm_model_lightn)
hist(residuals(lm_model_lightn))

AIC(lm_model_light, lm_model_lightn)

```

```{r, size='large', echo=FALSE, include=T, warning=FALSE, fig.width=6, fig.height=4}
library(lme4)
library(lmerTest)
library(MuMIn)

lm_model_light <- lmer(log(ER_filled*-1) ~ daily_PAR + (1|Light_group), data = df_light)
summary(lm_model_light)
hist(residuals(lm_model_light))
r.squaredGLMM(lm_model_light)

lm_model_lightn <- lm(log(ER_filled*-1) ~ discharge, data = df_light)
summary(lm_model_lightn)
hist(residuals(lm_model_lightn))

AIC(lm_model_light, lm_model_lightn)
```

#### Daily synchrony GPP and light Q1

```{r, warning=F, size = 'small', echo=F, include=F}

# df_light <- df_light %>% dplyr::filter(!is.na(Light_group))


GPP_long_q1 <- df_light %>%
 dplyr::filter(Light_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q1 <- GPP_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q1 <- Sncf(x=GPP_wide_day_q1$Lon,
                   y=GPP_wide_day_q1$Lat,
                   z=GPP_wide_day_q1[,4:ncol(GPP_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q1)


GPP_corrgram_summary.table <- summary(RR_df.corr_light_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.light.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_light_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "light_Q1")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.light.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.light.Q1$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.light.Q1 <- as.data.frame(t(RR_df.corr_light_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.light.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_light_q1$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.light.Q1$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.light.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "light_Q1")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


#### Daily synchrony in GPP and light Q2

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q2 <- df_light %>%
   dplyr::filter(Light_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q2 <- GPP_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q2 <- Sncf(x=GPP_wide_day_q2$Lon,
                   y=GPP_wide_day_q2$Lat,
                   z=GPP_wide_day_q2[,4:ncol(GPP_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q2)


GPP_corrgram_summary.table <- summary(RR_df.corr_light_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.light.q2 <- data.frame(x = matrix(unlist(RR_df.corr_light_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "light_Q2")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.light.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.light.q2$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.light.q2 <- as.data.frame(t(RR_df.corr_light_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.light.q2 <- data.frame(x = matrix(unlist(RR_df.corr_light_q2$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.light.q2$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.light.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "light_Q2")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


#### Daily synchrony in GPP and light Q3

```{r, warning=F, size = 'small', echo=F, include=F}
GPP_long_q3 <- df_light %>%
   dplyr::filter(Light_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(GPP_filled = mean(GPP_filled, na.rm = TRUE), .groups = "drop")

GPP_wide_day_q3 <- GPP_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = GPP_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q3 <- Sncf(x=GPP_wide_day_q3$Lon,
                   y=GPP_wide_day_q3$Lat,
                   z=GPP_wide_day_q3[,4:ncol(GPP_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q3)


GPP_corrgram_summary.table <- summary(RR_df.corr_light_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
GPP.corr.pred.light.q3 <- data.frame(x = matrix(unlist(RR_df.corr_light_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "GPP",
                               strat = "light_Q3")

## Sample average line
Regional.synch <- GPP_corrgram_summary.table$Regional.synch
GPP.corr.pred.light.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
GPP.corr.pred.light.q3$cbar <- Regional.synch_x

# ...and a dataframe of the y values for the bootstrap prediction
GPP.corr.bootValues.light.q3 <- as.data.frame(t(RR_df.corr_light_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
GPP.corr.boot.light.q3 <- data.frame(x = matrix(unlist(RR_df.corr_light_q3$boot$boot.summary$predicted$x)),
                                             yminGPP = GPP.corr.bootValues.light.q3$`0.025`,
                                             ymaxGPP = GPP.corr.bootValues.light.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "GPP",
                               strat = "light_Q3")

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(GPP_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(GPP_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(GPP_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(GPP_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_light <- bind_rows(GPP.corr.pred.light.Q1, GPP.corr.pred.light.q2, GPP.corr.pred.light.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_light   <- bind_rows(GPP.corr.boot.light.Q1, GPP.corr.boot.light.q2, GPP.corr.boot.light.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different types

```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_light <- pred_df_all_light %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_light <- pred_wide_light %>%
  mutate(
    light_Q1_diff = (abs(light_Q1 - light_Q1) / light_Q1) * 100,
    light_Q2_diff = (abs(light_Q1 - light_Q2) / light_Q1) * 100,
    light_Q3_diff = (abs(light_Q1 - light_Q3) / light_Q1) * 100)

pred_diff_light
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "light_Q1" = "#00BFC4",   
  "light_Q2" = "#7CAE00",
  "light_Q3" = "#F8766D"
)
```


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_light %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df




ERbar_chart_light  <-ci_df_all_light %>%
  filter(x_rounded_5<635)%>%
  group_by(strat) %>%
  summarise(
    ymin= mean(yminGPP, na.rm=T), 
    ymax=mean(ymaxGPP, na.rm=T))

coeff_chart_light <- pred_df_all_light %>%
  group_by(strat) %>%
  summarise(
    y= max(mean_synch))

coef_plot_light_dat <- coeff_chart_light %>%
  left_join(ERbar_chart_light, by =c("strat" = "strat"))




```


```{r, echo=FALSE, include=F, warning=FALSE, fig.width=8, fig.height=6}

###
GPP_synch_spline_light <- ggplot() +
  # Ribbon for CI
  geom_ribbon(data = ci_df_all_light,
              aes(x = x, ymin = yminGPP, ymax = ymaxGPP, fill = strat),
              alpha = 0.1) +
  # Main spline lines
  ylim(-0.2, 0.75) +
  geom_line(data = pred_df_all_light,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Draw horizontal segments from start to x_intersect
  geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  labs(title = "GPP across light",       
       y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
          geom_vline(xintercept = 807, color="grey40", lty=3) +

  scale_color_manual(values = strat_colors,
                         labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,     
                    labels = c("Low", "Moderate", "High")) +
  theme_classic() +
  theme(legend.position = "bottom")

GPP_synch_spline_light

```

Catchments with underlying metamorphic rock are more similar than
igneous or sedimentary

### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_light %>%
  group_by(strat) %>%
        filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

#### Daily synchrony ER and light Q1

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q1 <- df_light %>%
  dplyr::filter(Light_group=="Q1") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q1 <- ER_long_q1 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q1 <- Sncf(x=ER_wide_day_q1$Lon,
                   y=ER_wide_day_q1$Lat,
                   z=ER_wide_day_q1[,4:ncol(ER_wide_day_q1)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q1)


ER_corrgram_summary.table <- summary(RR_df.corr_light_q1)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.light.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_light_q1$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q1$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "light_Q1")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.light.Q1$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.light.Q1$cbar <- Regional.synch_x


# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.light.Q1 <- as.data.frame(t(RR_df.corr_light_q1$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.light.Q1 <- data.frame(x = matrix(unlist(RR_df.corr_light_q1$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.light.Q1$`0.025`,
                                             ymaxER = ER.corr.bootValues.light.Q1$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "light_Q1") 

```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```


#### Daily synchrony in ER and light Q2

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q2 <- df_light %>%
    dplyr::filter(Light_group=="Q2") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q2 <- ER_long_q2 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q2 <- Sncf(x=ER_wide_day_q2$Lon,
                   y=ER_wide_day_q2$Lat,
                   z=ER_wide_day_q2[,4:ncol(ER_wide_day_q2)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q2)


ER_corrgram_summary.table <- summary(RR_df.corr_light_q2)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.light.q2 <- data.frame(x = matrix(unlist(RR_df.corr_light_q2$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q2$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "light_Q2")
## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.light.q2$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.light.q2$cbar <- Regional.synch_x

# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.light.q2 <- as.data.frame(t(RR_df.corr_light_q2$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.light.q2 <- data.frame(x = matrix(unlist(RR_df.corr_light_q2$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.light.q2$`0.025`,
                                             ymaxER = ER.corr.bootValues.light.q2$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "light_Q2")
```


```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

#### Daily synchrony in ER and light Q3

```{r, warning=F, size = 'small', echo=F, include=F}
ER_long_q3 <- df_light %>%
      dplyr::filter(Light_group=="Q3") %>%
  group_by(Site_ID, date, Lat, Lon) %>%
  summarise(ER_filled = mean(ER_filled, na.rm = TRUE), .groups = "drop")

ER_wide_day_q3 <- ER_long_q3 %>%
  pivot_wider(
    names_from = date,
    values_from = ER_filled)

# Now, calculate spatial Synchrony in the time series using the Sncf() function
RR_df.corr_light_q3 <- Sncf(x=ER_wide_day_q3$Lon,
                   y=ER_wide_day_q3$Lat,
                   z=ER_wide_day_q3[,4:ncol(ER_wide_day_q3)],
                   resamp = 1000, 
                   latlon = T,
                   na.rm = T) # allows for NA given each stream has 280 days per year


plot(RR_df.corr_light_q3)


ER_corrgram_summary.table <- summary(RR_df.corr_light_q3)

# Plot using ggplot for export
# create a dataframe of the predicted values
ER.corr.pred.light.q3 <- data.frame(x = matrix(unlist(RR_df.corr_light_q3$real$predicted$x)),
                                             y = matrix(unlist(RR_df.corr_light_q3$real$predicted$y)),
                               time_scale = "daily",
                               parameter = "ER",
                               strat = "light_Q3")

## Sample average line
Regional.synch <- ER_corrgram_summary.table$Regional.synch
ER.corr.pred.light.q3$mean_synch <- round(Regional.synch,2)

Regional.synch_x <-  round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0)
ER.corr.pred.light.q3$cbar <- Regional.synch_x


# ..and a dataframe of the y values for the bootstrap prediction
ER.corr.bootValues.light.q3 <- as.data.frame(t(RR_df.corr_light_q3$boot$boot.summary$predicted$y))

# merge x and y values (min and max) for the bootstrap confidence intervals into a dataframe
ER.corr.boot.light.q3 <- data.frame(x = matrix(unlist(RR_df.corr_light_q3$boot$boot.summary$predicted$x)),
                                             yminER = ER.corr.bootValues.light.q3$`0.025`,
                                             ymaxER = ER.corr.bootValues.light.q3$`0.975`,
                               time_scale = "daily",
                                  parameter = "ER",
                               strat = "light_Q3")

```



```{r, warning=F, size = 'small', echo=F, include=T, cache=TRUE}
cat(paste0(
  "$y.intercept: ", round(ER_corrgram_summary.table$estimates["intercepts", "y"], 2), "\n",
  "$Regional.synch: ", round(ER_corrgram_summary.table$Regional.synch[1], 2), "\n",  
  "$cbar: ", round(ER_corrgram_summary.table$estimates["intercepts", "cbar"], 0), " km\n",
  "$x.intercept: ", round(ER_corrgram_summary.table$quantiles["0.5", "x"], 0), " km\n"
))
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=12, fig.height=8}

pred_df_all_light <- bind_rows(ER.corr.pred.light.Q1, ER.corr.pred.light.q2, ER.corr.pred.light.q3)%>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%dplyr::filter(x<half_dist_km)


ci_df_all_light   <- bind_rows(ER.corr.boot.light.Q1, ER.corr.boot.light.q2, ER.corr.boot.light.q3 ) %>%
  mutate(x_rounded_5 = round(x / 5) * 5)%>%
  group_by(time_scale, parameter, strat, x_rounded_5)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")%>%dplyr::filter(x<half_dist_km)

```

##### Calculate the percent difference between different light quantiles


```{r, echo=F, include=T, warning=FALSE, fig.width=12, fig.height=8}

# Spread the data so each LC Q becomes a column
pred_wide_light <- pred_df_all_light %>%
  dplyr::select(x_rounded_5, y, strat) %>%
  pivot_wider(names_from = strat, values_from = y)

# Calculate percent differences from daily for each other time scale
# Formula: ((other - daily) / daily) * 100
pred_diff_light <- pred_wide_light %>%
  mutate(
    light_Q1_diff = (abs(light_Q1 - light_Q1) / light_Q1) * 100,
    light_Q2_diff = (abs(light_Q1 - light_Q2) / light_Q1) * 100,
    light_Q3_diff = (abs(light_Q1 - light_Q3) / light_Q1) * 100)

pred_diff_light
```

```{r, echo=FALSE, include=F, warning=FALSE, fig.width=6, fig.height=3}
# Set a color-blind friendly palette
strat_colors <- c(
  "light_Q1" = "#00BFC4",   
  "light_Q2" = "#7CAE00",
  "light_Q3" = "#F8766D"
)
```


#### Look when synchrony dips below the sample average/ regional synch value for each timescale:

This distances where parameters are no longer regionally synchronous

```{r, warning=F, size = 'small', echo=T, include=T}

intersection_df <- pred_df_all_light %>%
  group_by(strat) %>%
  summarise(
    y= mean(mean_synch),
    x = mean(cbar))

intersection_df


```


### Figure 4c (?): Spline Correlogram of GPP at different light characteristics

```{r, warning=F, size = 'small', echo=T, include=T}

ER_synch_spline_light <- ggplot() +
  geom_ribbon(data = ci_df_all_light,
              aes(x = x, ymin = yminER, ymax = ymaxER, fill = strat),
              alpha = 0.1) +
  geom_line(data = pred_df_all_light,
            aes(x = x, y = y, color = strat),
            size = 1) +
  # Draw horizontal segments from start to x_intersect
  geom_segment(data = intersection_df,
               aes(x = 0, xend = x, 
                   y = y, yend = y,
                   color = strat),
               linetype = "dashed") +
  # Draw vertical segments from bottom to y_intersect
  geom_segment(data = intersection_df,
               aes(x = x, xend = x, 
                   y = -0.2, yend = y,
                   color = strat),
               linetype = "dashed") +
  labs(title = "ER across light",
       y = "Correlation",
       x = "Distance (km)",
       fill = "Quantiles",
       color = "Quantiles") +
   geom_vline(xintercept = 1083, color="grey40", lty=3) +
  scale_color_manual(values = strat_colors,
                         labels = c("Low", "Moderate", "High")) +
  scale_fill_manual(values = strat_colors,     
                    labels = c("Low", "Moderate", "High")) +
  theme_classic() +  ylim(-0.2, 0.75) +
  theme(legend.position = "bottom")


```


### Quantify synchrony decay with distance:

Where b parameter represents the decay rate: higher b = faster loss of
synchrony with distance.

```{r, warning=F, size = 'small', echo=F, include=T}

fit_models <- pred_df_all_light %>%
  group_by(strat) %>%
          filter(x<cbar)%>%
  nest() %>%
  mutate(model = map(data, ~ nls(y ~ a * exp(-b * x), data = ., start = list(a = 1, b = 0.001))),
         params = map(model, tidy)) %>%
  unnest(params) %>%
  mutate(across(c(estimate, std.error, statistic, p.value), ~ round(.x, 5), .names = "rounded_{.col}"))

fit_models_tab <- as.data.frame(fit_models)%>%
  dplyr::select(strat, term,rounded_estimate: rounded_p.value)

fit_models_tab
```

```{r, warning=F, size = 'small', echo=F,include=TRUE, fig.width=10, fig.height=12}
# Arrange plots with shared legend at the bottom
library(ggpubr)

spline_grid2 <- ggarrange(
GPP_synch_spline_flow,
ER_synch_spline_flow,
Flow_dist_plt,
GPP_synch_spline_light,
ER_synch_spline_light,
Light_dist_plt,
GPP_synch_spline_temp,
ER_synch_spline_temp,
Temp_dist_plt,
  ncol = 3, nrow = 3,
  common.legend = T, 
  widths = c(1, 1, 0.65),
  legend = "bottom"
)

spline_grid2
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/River_Recovery_Dynamics_Analysis/figures/Finescale_spline_correlogram.png", plot = spline_grid2, width = 8.25, height = 7.1, units = "in")

```



```{r}

coef_plot <- rbind(daily_dat,coef_plot_light_dat, coef_plot_temp_dat, coef_plot_flow_dat)


library(ggplot2)

ggplot(coef_plot, aes(y = strat, x = y)) +
  geom_point(size = 3, color = "steelblue") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2, color = "steelblue") +
  theme_minimal(base_size = 14) +
  coord_flip() +  # optional: flip axes for readability
  labs(
    title = "Coefficient Plot",
    x = "Stratification",
    y = "Estimate (with CI)"
  )


```
